### 2024-11-26 paper/paper.qmd
update research topic and ideas to paper

---
title: "Analyzing Toronto Fire Incident Severity"
subtitle: "Study Thorugh Financial and Casualty Loss to Optimize Resource Allocation for Toronto Fire Services"
author: 
  - Maggie Zhang
thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
date: today
date-format: long
abstract: "This study investigates different severity levels of Toronto fire incidents based on financial and casualty losses to support efficient resource allocation for Toronto Fire Service (TFS). The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters. The study identifies high-priority response criteria, including incidents originating from XXXX regions, occurring within defined time frames: XXXX, and involving XXXX as the ignition sources. These insights, derived from the modeling analysis, can guide TFS to enhance response times and prioritize tasks based on common high-risk indicators. This work aims to inform decision-making and optimize fire response operations to improve community safety in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

This study investigates different severity levels of Toronto fire incidents based on financial and casualty losses to support efficient resource allocation for Toronto Fire Service (TFS). The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters.
The research identifies specific high-priority response criteria that can improve TFS operations. By examining the data, it has been found that incidents originating from XXX regions, occurring within defined time frames: XXX, and involving XXXX ignition sources exhibit higher risk profiles. This enables TFS to develop strategic response protocols that prioritize these types of incidents. The modeling analysis conducted in this study provides valuable insights into which factors contribute most significantly to incident severity, equipping TFS with the knowledge to enhance response times and allocate resources more effectively.

The insights from this work aim to inform strategic decision-making processes and optimize fire response operations. By understanding and addressing the indicators of high-risk incidents, TFS can better safeguard the community and improve public safety throughout Toronto. This research is an essential step toward developing data-driven approaches that enhance both the efficiency of fire service operations and the overall resilience of the city.

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-12-01 scripts/02-download_data.R
download data drom web, and store it in repository

@@ -1,26 +1,26 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Data Toronto (https://open.toronto.ca/dataset/fire-incidents/)
+# Author: Maggie Zhang
+# Date: 29 November 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Plan data ans simulated
+# Any other information needed? NO
 
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+#### TFS Data ####
+
+raw_tfs_fire <- read_csv(file = "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/64a26694-01dc-4ec3-aa87-ad8509604f50/resource/1e824947-d73b-4f48-9bac-7f7f3731a6b9/download/Fire%20Incidents%20Data.csv",
+                            show_col_types = FALSE)
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# Save the data to the repository
+write_csv(raw_tfs_fire, "/data/raw_fire_data.csv") 
 
          

### 2024-12-01 scripts/02-download_data.R
Save download file

@@ -21,6 +21,6 @@ raw_tfs_fire <- read_csv(file = "https://ckan0.cf.opendata.inter.prod-toronto.ca
 
 #### Save data ####
 # Save the data to the repository
-write_csv(raw_tfs_fire, "/data/raw_fire_data.csv") 
+write_csv(raw_tfs_fire, "data/01-raw_data/raw_fire_data.csv") 
 
          

### 2024-12-01 scripts/03-clean_data.R
Data cleaning such as remove NAs according to context, select columns to study, change date time etc.

@@ -1,17 +1,21 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data by Open Data Toronto
+# Author: Maggie Zhang
+# Date: 30 November 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Have the data downloaded
+# Any other information needed? No
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
+
+str(raw_data)
+
 
 cleaned_data <-
   raw_data |>

### 2024-12-01 scripts/03-clean_data.R
Create a column calculate tfs response time

@@ -10,39 +10,87 @@
 #### Workspace setup ####
 library(tidyverse)
 library(dplyr)
+library(lubridate)
 
-#### Clean data ####
+#### Load data ####
 raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
 
-str(raw_data)
+### Keep Columns in interest ###
+columns_to_keep <- c("_id", "Area_of_Origin", "Building_Status", "Business_Impact", 
+                     "Civilian_Casualties", "Estimated_Dollar_Loss", "Exposures", 
+                     "Final_Incident_Type", "Fire_Alarm_System_Operation", 
+                     "Ignition_Source", "Initial_CAD_Event_Type", 
+                     "Material_First_Ignited", "Method_Of_Fire_Control", 
+                     "Number_of_responding_apparatus", "Number_of_responding_personnel", 
+                     "Possible_Cause", "Smoke_Spread", "Sprinkler_System_Presence", 
+                     "TFS_Alarm_Time", "TFS_Arrival_Time", "TFS_Firefighter_Casualties")
+
+clean_data <- raw_data[, columns_to_keep]
+# Change columns name to lower case and ensure no space
+colnames(clean_data) <- tolower(gsub(" ", "_", colnames(clean_data)))
+
+### Preview data ###
+colnames(clean_data)
+head(clean_data, 6)
+summary(clean_data)   
+
+### Ensure Datetime Consistency ###
+
+# Convert timestamps to POSIXct format (YYYY-MM-DD HH:MM:SS)
+clean_data$tfs_alarm_time <- ymd_hms(clean_data$tfs_alarm_time)  
+clean_data$tfs_arrival_time <- ymd_hms(clean_data$tfs_arrival_time) 
+clean_data <- clean_data[!is.na(clean_data$tfs_alarm_time), ]
+clean_data <- clean_data[!is.na(clean_data$tfs_arrival_time), ]
 
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
[…]

### 2024-12-01 scripts/00-simulate_data.R
update simulation script code

@@ -1,52 +1,115 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates Data
+# Author: Maggie Zhang 
+# Date: 19 September 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# Pre-requisites: First sketch the idea of the dataset
+# Any other information needed? Not much
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
+set.seed(1009633096)
 
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+
+clean_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
+
+#number of rows
+n_simulated <- nrow(clean_data)
+
+## Numerical Values ##
+# civilian_casualties, use Poisson distribution for count data
+simulated_civilian_casualties <- rpois(n_simulated, 
+                                       mean(clean_data$civilian_casualties,
+                                            na.rm = TRUE))  
+# Estimated lost, use Normal distribution for continuous data
+simulated_estimated_dollar_loss <- rnorm(n_simulated, 
+                                         mean(clean_data$estimated_dollar_loss, 
+                                              na.rm = TRUE), 
+                                         sd(clean_data$estimated_dollar_loss, 
+                                            na.rm = TRUE))
+# Number of responsing apparatus, use random number that range from 1 to 50.
+simulated_number_of_responding_apparatus <- sample(1:50, n_simulated, 
+                                                   replace = TRUE)  
+
+
+## Categorical Data ##
+# Fire alarm system Operation, distributed roughly by it's percentage in original dataset
+fire_alarm_categories <- unique(clean_data$fire_alarm_system_operation)
+category_probs <- table(clean_data$fire_alarm_system_operation) / nrow(clean_data)
+simulated_fire_alarm_system_operation <- sample(fire_alarm_categories, 
[…]

### 2024-12-01 scripts/00-simulate_data.R
make sure no negtive numeric variables

@@ -23,13 +23,29 @@ n_simulated <- nrow(clean_data)
 # civilian_casualties, use Poisson distribution for count data
 simulated_civilian_casualties <- rpois(n_simulated, 
                                        mean(clean_data$civilian_casualties,
-                                            na.rm = TRUE))  
-# Estimated lost, use Normal distribution for continuous data
-simulated_estimated_dollar_loss <- rnorm(n_simulated, 
-                                         mean(clean_data$estimated_dollar_loss, 
-                                              na.rm = TRUE), 
-                                         sd(clean_data$estimated_dollar_loss, 
-                                            na.rm = TRUE))
+                                            na.rm = TRUE)) 
+
+# Estimated lost, use Log-normal distribution, with adjustment to mean and sd.
+clean_data_no_na <- clean_data$estimated_dollar_loss[
+  !is.na(clean_data$estimated_dollar_loss) & 
+    clean_data$estimated_dollar_loss > 0]
+log_mean <- log(mean(clean_data_no_na))
+log_sd <- sd(log(clean_data_no_na))
+simulated_estimated_dollar_loss <- rlnorm(
+  n_simulated, meanlog = log_mean, sdlog = log_sd)
+# Adjust the distribution to get a more realistic mean and quantiles
+target_mean <- mean(clean_data$estimated_dollar_loss, na.rm = TRUE)
+simulated_estimated_dollar_loss <- simulated_estimated_dollar_loss * 
+  (target_mean / mean(simulated_estimated_dollar_loss))
+# Apply truncation based on the original data's max value
+max_loss <- max(clean_data$estimated_dollar_loss, na.rm = TRUE)
+simulated_estimated_dollar_loss <- 
+  pmin(simulated_estimated_dollar_loss, max_loss)
+# Ensure the same proportion of zero values as in the original data
+zero_percentage <- mean(clean_data$estimated_dollar_loss == 0, na.rm = TRUE)
+simulated_estimated_dollar_loss[
+  sample(1:n_simulated, size = round(zero_percentage * n_simulated))] <- 0
+
 # Number of responsing apparatus, use random number that range from 1 to 50.
 simulated_number_of_responding_apparatus <- sample(1:50, n_simulated, 
                                                    replace = TRUE)  

### 2024-12-01 scripts/01-test_simulated_data.R
Test for simulated data, including check for NA, NUll, negtive, data-time, etc

@@ -1,89 +1,222 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated TFS dataset 
+# Author: Maggie Zhang
+# Date: 30 November 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+  # - 03-clean_data.R must have been run
 
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
+# Test if the simulated data was successfully loaded
+if (exists("simulated_data")) {
+  message("Test Passed: The simulated dataset was successfully loaded.")
 } else {
-  stop("Test Failed: The dataset could not be loaded.")
+  stop("Test Failed: The simulated dataset could not be loaded.")
 }
 
 
-#### Test data ####
+### Test Simulated Data - Structure ###
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the simulated dataset has same number of rows as actual analysis data
+if (nrow(simulated_data) == nrow(analysis_data)) {
+  message("Test Passed: The simulated dataset has ", 
+          nrow(analysis_data), " rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  message("Test Failed: The simulated dataset does not have ", 
+          nrow(analysis_data), 
+          " rows. It has ", nrow(simulated_data), " rows instead.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
[…]

### 2024-12-01 scripts/01-test_simulated_data.R
Add one test to simulated data to test time input

@@ -204,7 +204,7 @@ if (any(simulated_data$response_time < 0, na.rm = TRUE)) {
 }
 
 
-### Test Simulated Data - Date-Time Variables (Format)###
+### Test Simulated Data - Date-Time Variables (Format and Consistency)###
 
 # Check if 'tfs_alarm_time' follows the correct format
 if (any(is.na(as.POSIXct(simulated_data$tfs_alarm_time, format = "%Y-%m-%d %H:%M:%S", tz = "UTC")))) {
@@ -220,3 +220,14 @@ if (any(is.na(as.POSIXct(simulated_data$tfs_arrival_time, format = "%Y-%m-%d %H:
   message("Test Passed: Date format is valid in the 'tfs_arrival_time' column.")
 }
 
+# Check if all arrival times occur after alarm times
+invalid_time_sequence <- 
+  sum(simulated_data$tfs_arrival_time < simulated_data$tfs_alarm_time, 
+      na.rm = TRUE)
+
+if (invalid_time_sequence == 0) {
+  message("Test Passed: All arrival times occur after alarm times.")
+} else {
+  message("Test Failed: There are ", invalid_time_sequence, 
+          " cases where arrival time is before alarm time.")
+}

### 2024-12-01 scripts/01-test_simulated_data.R
add two more test to test range of time variables

@@ -231,3 +231,34 @@ if (invalid_time_sequence == 0) {
   message("Test Failed: There are ", invalid_time_sequence, 
           " cases where arrival time is before alarm time.")
 }
+
+
+# Check alarm time is in a realistic range
+min_alarm_time <- min(simulated_data$tfs_alarm_time, na.rm = TRUE)
+max_alarm_time <- max(simulated_data$tfs_alarm_time, na.rm = TRUE)
+expected_alarm_start <- as.POSIXct('2011-01-01')
+expected_alarm_end <- as.POSIXct('2024-11-30')
+
+if (min_alarm_time >= expected_alarm_start & max_alarm_time <= expected_alarm_end) {
+  message("Test Passed: Alarm times are within the realistic range.")
+} else {
+  message("Test Failed: Alarm times are outside the realistic range. ",
+          "Minimum: ", min_alarm_time, 
+          ", Maximum: ", max_alarm_time)
+}
+
+# Check arrival time is in a realistic range
+min_arrive_time <- min(simulated_data$tfs_arrival_time, na.rm = TRUE)
+max_arrive_time <- max(simulated_data$tfs_arrival_time, na.rm = TRUE)
+expected_arrive_start <- as.POSIXct('2011-01-01')
+expected_arrive_end <- as.POSIXct('2024-11-30')
+
+if (min_arrive_time >= expected_arrive_start & 
+    max_arrive_time <= expected_arrive_end) {
+  message(
+    "Test Passed: Arrival times are within the realistic range.")
+} else {
+  message("Test Failed: Arrival times are outside the realistic range. ",
+          "Minimum: ", min_arrive_time, 
+          ", Maximum: ", max_arrive_time)
+}

### 2024-12-01 scripts/01-test_simulated_data.R
update tests for zero

@@ -262,3 +262,19 @@ if (min_arrive_time >= expected_arrive_start &
           "Minimum: ", min_arrive_time, 
           ", Maximum: ", max_arrive_time)
 }
+
+
+### Check for Unreasonable zeros ###
+# Check for zero values in 'response_time'
+if (any(simulated_data$response_time == 0, na.rm = TRUE)) {
+  stop("Test Failed: Zero values found in the 'response_time' column.")
+} else {
+  message("Test Passed: No zero values in the 'response_time' column.")
+}
+
+# Check for zero values in 'id'
+if (any(simulated_data$id == 0, na.rm = TRUE)) {
+  stop("Test Failed: Zero values found in the 'id' column.")
+} else {
+  message("Test Passed: No zero values in the 'id' column.")
+}

### 2024-12-01 scripts/04-test_analysis_data.R

@@ -1,69 +1,269 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the structure and validity of the analysis(cleaned) TFS dataset 
+# Author: Maggie Zhang
+# Date: 30 November 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
+# - 03-clean_data.R must have been run
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(testthat)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
+raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
 
 
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
+# Test if the Analysis data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The analysis dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The analysis dataset could not be loaded.")
+}
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
+### Test Analysis Data - Repetitiveness ###
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
+# Check proportion of rows retained, consider representative if at least 60% o original retianed.
+retention_rate <- (nrow(analysis_data) / nrow(raw_data)) * 100
+if (retention_rate > 60) {
+  message("Test Passed: The proportion of rows retained is ", 
+          retention_rate, "%. Consider representitive.")
+} else {
[…]

### 2024-12-01 scripts/00-simulate_data.R
edit simulation to better reflect actual data

@@ -101,19 +101,13 @@ simulated_data <- data.frame(
   fire_alarm_system_operation = simulated_fire_alarm_system_operation,
   ignition_source = sample(unique(clean_data$ignition_source), 
                            n_simulated, replace = TRUE),
-  initial_cad_event_type = sample(unique(clean_data$initial_cad_event_type), 
-                                  n_simulated, replace = TRUE),
   material_first_ignited = sample(unique(clean_data$material_first_ignited), 
-                                  n_simulated, replace = TRUE),
-  method_of_fire_control = sample(unique(clean_data$method_of_fire_control), 
                                   n_simulated, replace = TRUE),
   number_of_responding_apparatus = simulated_number_of_responding_apparatus,
   number_of_responding_personnel = sample(10:50, 
                                           n_simulated, replace = TRUE),
   possible_cause = sample(unique(clean_data$possible_cause), 
                           n_simulated, replace = TRUE),
-  smoke_spread = sample(unique(clean_data$smoke_spread), 
-                        n_simulated, replace = TRUE),
   sprinkler_system_presence = sample(
     unique(clean_data$sprinkler_system_presence), 
     n_simulated, replace = TRUE),

### 2024-12-01 scripts/03-clean_data.R
more column selection

@@ -19,10 +19,10 @@ raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
 columns_to_keep <- c("_id", "Area_of_Origin", "Building_Status", "Business_Impact", 
                      "Civilian_Casualties", "Estimated_Dollar_Loss", "Exposures", 
                      "Final_Incident_Type", "Fire_Alarm_System_Operation", 
-                     "Ignition_Source", "Initial_CAD_Event_Type", 
-                     "Material_First_Ignited", "Method_Of_Fire_Control", 
+                     "Ignition_Source", 
+                     "Material_First_Ignited", 
                      "Number_of_responding_apparatus", "Number_of_responding_personnel", 
-                     "Possible_Cause", "Smoke_Spread", "Sprinkler_System_Presence", 
+                     "Possible_Cause", "Sprinkler_System_Presence", 
                      "TFS_Alarm_Time", "TFS_Arrival_Time", "TFS_Firefighter_Casualties")
 
 clean_data <- raw_data[, columns_to_keep]
@@ -34,6 +34,10 @@ colnames(clean_data)
 head(clean_data, 6)
 summary(clean_data)   
 
+# Since I want to study possible sources leading to different severity of fire, no info about sources should be dropped
+clean_data <- clean_data %>%
+  filter(ignition_source != "999 - Undetermined")
+
 ### Ensure Datetime Consistency ###
 
 # Convert timestamps to POSIXct format (YYYY-MM-DD HH:MM:SS)
@@ -41,6 +45,7 @@ clean_data$tfs_alarm_time <- ymd_hms(clean_data$tfs_alarm_time)
 clean_data$tfs_arrival_time <- ymd_hms(clean_data$tfs_arrival_time) 
 clean_data <- clean_data[!is.na(clean_data$tfs_alarm_time), ]
 clean_data <- clean_data[!is.na(clean_data$tfs_arrival_time), ]
+
 
 
 ### Missing Value ###
@@ -51,19 +56,18 @@ missing_info <- data.frame(Column = names(missing_percentage),
                            Missing_Percentage = missing_percentage)
 print(missing_info)
 
-## Identify columns with more than 30% missing values and remove the columns
-cols_to_remove <- missing_info$Column[missing_info$Missing_Percentage > 30]
+## Identify columns with more than 20% missing values and remove the columns
+cols_to_remove <- missing_info$Column[missing_info$Missing_Percentage > 20]
 cols_to_remove
 
[…]

### 2024-12-01 scripts/01-test_simulated_data.R
simulation edit, align with new dataset

@@ -78,12 +78,12 @@ if (any(null_cells)) {
 ### Test Simulated Data - Categorical Variables (Valid Categories)###
 
 # Check if area_of_origin columns contains only valid categories as in actual data
-valid_area_categories <- unique(analysis_data$fire_alarm_system_operation)
+valid_area_categories <- unique(analysis_data$area_of_origin)
 
-if (all(simulated_data$fire_alarm_system_operation %in% valid_area_categories)) {
-  message("Test Passed: The 'fire_alarm_system_operation' column contains only valid categories as actual dataset.")
+if (all(simulated_data$area_of_origin %in% valid_area_categories)) {
+  message("Test Passed: The 'area_of_origin' column contains only valid categories as actual dataset.")
 } else {
-  stop("Test Failed: The 'fire_alarm_system_operation' column contains invalid categories.")
+  stop("Test Failed: The 'area_of_origin' column contains invalid categories.")
 }
 
 # Check if fire_alarm_system_operation contains only valid categories as in actual data
@@ -110,14 +110,6 @@ if (all(simulated_data$ignition_source %in% valid_ignition_source_categories)) {
   stop("Test Failed: The 'ignition_source' column contains invalid categories.")
 }
 
-# Initial CAD event type
-valid_initial_cad_event_categories <- unique(analysis_data$initial_cad_event_type)
-if (all(simulated_data$initial_cad_event_type %in% valid_initial_cad_event_categories)) {
-  message("Test Passed: The 'initial_cad_event_type' column contains only valid categories.")
-} else {
-  stop("Test Failed: The 'initial_cad_event_type' column contains invalid categories.")
-}
-
 # Material first ignited
 valid_material_first_ignited_categories <- unique(analysis_data$material_first_ignited)
 if (all(simulated_data$material_first_ignited %in% valid_material_first_ignited_categories)) {
@@ -126,28 +118,12 @@ if (all(simulated_data$material_first_ignited %in% valid_material_first_ignited_
   stop("Test Failed: The 'material_first_ignited' column contains invalid categories.")
 }
 
-# Method of fire control
-valid_method_of_fire_control_categories <- unique(analysis_data$method_of_fire_control)
-if (all(simulated_data$method_of_fire_control %in% valid_method_of_fire_control_categories)) {
-  message("Test Passed: The 'method_of_fire_control' column contains only valid categories.")
-} else {
[…]

### 2024-12-01 scripts/04-test_analysis_data.R
update test accordingly with new dataset

@@ -26,13 +26,13 @@ if (exists("analysis_data")) {
 
 ### Test Analysis Data - Repetitiveness ###
 
-# Check proportion of rows retained, consider representative if at least 60% o original retianed.
+# Check proportion of rows retained, consider representative if at least 40% o original retianed.
 retention_rate <- (nrow(analysis_data) / nrow(raw_data)) * 100
-if (retention_rate > 60) {
+if (retention_rate > 40) {
   message("Test Passed: The proportion of rows retained is ", 
           retention_rate, "%. Consider representitive.")
 } else {
-  message("Test Failed: Less than 60% of rows retained, but got ", 
+  message("Test Failed: Less than 40% of rows retained, but got ", 
           retention_rate, "% instead.")
 }
 
@@ -68,12 +68,12 @@ if (any(null_cells)) {
 ### Test Analysis Data - Categorical Variables (Valid Categories)###
 
 # Check if area_of_origin columns contains only valid categories as in actual data
-valid_area_categories <- unique(analysis_data$fire_alarm_system_operation)
+valid_area_categories <- unique(analysis_data$area_of_origin)
 
-if (all(analysis_data$fire_alarm_system_operation %in% valid_area_categories)) {
-  message("Test Passed: The 'fire_alarm_system_operation' column contains only valid categories as actual dataset.")
+if (all(analysis_data$area_of_origin %in% valid_area_categories)) {
+  message("Test Passed: The 'area_of_origin' column contains only valid categories as actual dataset.")
 } else {
-  stop("Test Failed: The 'fire_alarm_system_operation' column contains invalid categories.")
+  stop("Test Failed: The 'area_of_origin' column contains invalid categories.")
 }
 
 # Check if fire_alarm_system_operation contains only valid categories as in actual data
@@ -100,13 +100,6 @@ if (all(analysis_data$ignition_source %in% valid_ignition_source_categories)) {
   stop("Test Failed: The 'ignition_source' column contains invalid categories.")
 }
 
-# Initial CAD event type
-valid_initial_cad_event_categories <- unique(analysis_data$initial_cad_event_type)
-if (all(analysis_data$initial_cad_event_type %in% valid_initial_cad_event_categories)) {
-  message("Test Passed: The 'initial_cad_event_type' column contains only valid categories.")
-} else {
-  stop("Test Failed: The 'initial_cad_event_type' column contains invalid categories.")
-}
 
[…]

### 2024-12-01 scripts/05-exploratory_data_analysis.R

@@ -1,37 +1,171 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Explore data of variables in interest and plot them
+# Author: Maggie Zhang 
+# Date: 30 November 2024
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(lubridate)
+library(knitr)
+
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+fire_data <- read.csv("data/02-analysis_data/tfs_analysis_data")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+### Data Overview ###
+# Structure
+head(fire_data)
+str(fire_data)
+
+# Missing Value
+missing_data <- colSums(is.na(fire_data))
+print(missing_data)
 
 
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "models/first_model.rds"
-)
+### Plot/Summary distribution of all Variables ###
+## 1. Civilian Casualties (Numerical)
+#Overall Table
+civilian_casualties_df <- as.data.frame(table(fire_data$civilian_casualties))
+colnames(civilian_casualties_df) <- c("Number_of_Casualties", "Frequency")
+civilian_casualties_df
+# Summary table
+summary(fire_data$civilian_casualties)
+
+### 2. Estimated Dollar Loss (Numerical)
+#Overall Table
+est_loss_df <- as.data.frame(table(fire_data$estimated_dollar_loss))
[…]

### 2024-12-01 scripts/03-clean_data.R
update  data for better cleaning

@@ -19,8 +19,7 @@ raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
 columns_to_keep <- c("_id", "Area_of_Origin", "Building_Status", "Business_Impact", 
                      "Civilian_Casualties", "Estimated_Dollar_Loss", "Exposures", 
                      "Final_Incident_Type", "Fire_Alarm_System_Operation", 
-                     "Ignition_Source", 
-                     "Material_First_Ignited", 
+                     "Ignition_Source",
                      "Number_of_responding_apparatus", "Number_of_responding_personnel", 
                      "Possible_Cause", "Sprinkler_System_Presence", 
                      "TFS_Alarm_Time", "TFS_Arrival_Time", "TFS_Firefighter_Casualties")

### 2024-12-01 scripts/00-simulate_data.R
Change simulation according to new dataset

@@ -101,8 +101,6 @@ simulated_data <- data.frame(
   fire_alarm_system_operation = simulated_fire_alarm_system_operation,
   ignition_source = sample(unique(clean_data$ignition_source), 
                            n_simulated, replace = TRUE),
-  material_first_ignited = sample(unique(clean_data$material_first_ignited), 
-                                  n_simulated, replace = TRUE),
   number_of_responding_apparatus = simulated_number_of_responding_apparatus,
   number_of_responding_personnel = sample(10:50, 
                                           n_simulated, replace = TRUE),

### 2024-12-01 scripts/01-test_simulated_data.R
write new test for simulated and analysis dataset

@@ -110,14 +110,6 @@ if (all(simulated_data$ignition_source %in% valid_ignition_source_categories)) {
   stop("Test Failed: The 'ignition_source' column contains invalid categories.")
 }
 
-# Material first ignited
-valid_material_first_ignited_categories <- unique(analysis_data$material_first_ignited)
-if (all(simulated_data$material_first_ignited %in% valid_material_first_ignited_categories)) {
-  message("Test Passed: The 'material_first_ignited' column contains only valid categories.")
-} else {
-  stop("Test Failed: The 'material_first_ignited' column contains invalid categories.")
-}
-
 # Possible cause
 valid_possible_cause_categories <- unique(analysis_data$possible_cause)
 if (all(simulated_data$possible_cause %in% valid_possible_cause_categories)) {

### 2024-12-01 scripts/04-test_analysis_data.R

@@ -101,15 +101,6 @@ if (all(analysis_data$ignition_source %in% valid_ignition_source_categories)) {
 }
 
 
-# Material first ignited
-valid_material_first_ignited_categories <- unique(analysis_data$material_first_ignited)
-if (all(analysis_data$material_first_ignited %in% valid_material_first_ignited_categories)) {
-  message("Test Passed: The 'material_first_ignited' column contains only valid categories.")
-} else {
-  stop("Test Failed: The 'material_first_ignited' column contains invalid categories.")
-}
-
-
 # Possible cause
 valid_possible_cause_categories <- unique(analysis_data$possible_cause)
 if (all(analysis_data$possible_cause %in% valid_possible_cause_categories)) {

### 2024-12-01 scripts/05-exploratory_data_analysis.R
EDA, data overview and plot of every variables in dataset. Preapare for later paper data section

@@ -19,7 +19,7 @@ library(knitr)
 #### Read data ####
 fire_data <- read.csv("data/02-analysis_data/tfs_analysis_data")
 
-### Data Overview ###
+#### Data Overview ####
 # Structure
 head(fire_data)
 str(fire_data)
@@ -29,24 +29,189 @@ missing_data <- colSums(is.na(fire_data))
 print(missing_data)
 
 
-### Plot/Summary distribution of all Variables ###
-## 1. Civilian Casualties (Numerical)
+#### Plot/Summary distribution of all Variables ####
+## 1. Area of Origin (Categorical) ##
+all_area <- unique(fire_data$area_of_origin)
+# List of Area of Origin
+area_of_origin <- c("22 - Sleeping Area or Bedroom (inc. patients room, dormitory, etc)", 
+                    "55 - Mechanical/Electrical Services Room", 
+                    "28 - Office", 
+                    "24 - Cooking Area or Kitchen", 
+                    "81 - Engine Area", 
+                    "25 - Washroom or Bathroom (toilet, restroom/locker room)", 
+                    "73 - Parking Area, Parking Lot", 
+                    "44 - Trash, Rubbish Storage (inc garbage chute room, garbage/industr", 
+                    "50 - Basement/cellar (not partitioned)", 
+                    "64 - Porch or Balcony", 
+                    "78 - Attached Deck", 
+                    "86 - Passenger Area", 
+                    "52 - HVAC Equipment Room (furnace room, water heater closet, boiler)", 
+                    "21 - Living Area (e.g. living, TV, recreation, etc)", 
+                    "61 - Exterior Wall", 
+                    "62 - Roof", 
+                    "12 - Hallway, Corridor", 
+                    "83 - Electrical Systems", 
+                    "59 - Utility Shaft (eg. electrical wiring/phone, etc.)", 
+                    "79 - Other Outside Area", 
+                    "75 - Trash, rubbish area (outside)", 
+                    "89 - Other Vehicle Area", 
+                    "31 - Process Manufacturing (inc manf, prod assembly, repair)", 
+                    "53 - Chimney/Flue Pipe", 
+                    "26 - Sauna", 
+                    "99 - Undetermined (formerly 98)", 
+                    "87 - Trunk/Cargo Area", 
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R
understand variables possibilty related to severity. Correlation study

@@ -510,4 +510,192 @@ ggplot(fire_data, aes(x = response_time)) +
 summary(fire_data$response_time)
 
 
+#### Understand Potential Indicators of Severity ####
+## Flag High-Severity Incident##
+# Define thresholds (example: $30000 loss or 5+ casualties)
+fire_data$high_severity <- ifelse(fire_data$estimated_dollar_loss > 50000 | fire_data$total_casualties >= 5, "High", "Low")
+# Count of High vs. Low severity
+table(fire_data$high_severity)
+
+# Bar chart of high vs. low severity
+ggplot(fire_data, aes(x = high_severity, fill = high_severity)) +
+  geom_bar() +
+  labs(
+    title = "Count of High vs. Low Severity Incidents",
+    x = "Severity Level",
+    y = "Count"
+  ) +
+  scale_fill_manual(values = c("High" = "darkblue", "Low" = "darkgreen")) +
+  theme_minimal()
+
+## Financial loss vs. Casualties ##
+fire_data$total_casualties <- fire_data$civilian_casualties + fire_data$tfs_firefighter_casualties
+
+#heatmap
+fire_data <- fire_data %>%
+  mutate(
+    loss_group = cut(
+      estimated_dollar_loss,
+      breaks = c(0, 1000, 10000, 50000, 500000, Inf),
+      labels = c("<1k", "1k-10k", "10k-50k", "50k-500k", ">500k"),
+      right = FALSE
+    ),
+    casualty_group = cut(
+      total_casualties,
+      breaks = c(0, 1, 2, 5, Inf),
+      labels = c("0", "1", "2-5", ">5"),
+      right = FALSE
+    )
+  )
+
+severity_table <- fire_data %>%
+  group_by(loss_group, casualty_group) %>%
+  summarize(count = n())
+
+ggplot(severity_table, aes(x = loss_group, y = casualty_group, fill = count)) +
+  geom_tile(color = "white") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(
+    title = "Heatmap of Financial Loss and Casualty Groups",
+    x = "Financial Loss Group",
+    y = "Casualty Group",
+    fill = "Incident Count"
+  ) +
+  theme_minimal()
+
+# scatter
+ggplot(fire_data, aes(x = as.factor(total_casualties), y = estimated_dollar_loss)) +
+  geom_boxplot(fill = "skyblue", outlier.color = "red") +
+  labs(
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R
more exploartion on variables contributed to classification of severe

@@ -647,6 +647,22 @@ fire_data %>%
        x = "Month", y = "Average Response Time (minutes)") +
   theme_minimal()
 
+
+##number of responding apparatus and the Financial dollar loss ##
+fire_data$Log_Dollar_Loss <- log(fire_data$estimated_dollar_loss + 1)
+
+# Scatter plot between number of responding apparatus and final dollar loss
+ggplot(fire_data, aes(y = number_of_responding_apparatus, x = Log_Dollar_Loss)) +
+  geom_point(color = 'blue', alpha = 0.5) +
+  labs(title = 'Number of Responding Apparatus vs Log of Final Dollar Loss',
+       x = 'Number of Responding Apparatus',
+       y = 'Log of Final Dollar Loss') +
+  theme_minimal()
+
+cor(fire_data$Log_Dollar_Loss, fire_data$number_of_responding_apparatus, use = "complete.obs")
+
+
+
 #### Response Time Understanding####
 ## Outliers ##
 # Calculate IQR for response_time
@@ -697,5 +713,30 @@ ggplot(fire_data_clean_loss, aes(x = estimated_dollar_loss)) +
   scale_x_continuous(labels = scales::comma) +  # Format axis labels with commas
   theme_minimal()
 
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+####
 
 

### 2024-12-03 scripts/06-model_data.R
Initial model for predicting serverity

@@ -11,21 +11,94 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(rpart)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+fire_data_model <- read_csv("data/02-analysis_data/tfs_analysis_data")
+
+#### Quantify Severity ####
+
+# Load libraries
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(fitdistrplus)
+
+# Step 1: Choose a column to quantify fire severity, (number_of_responding_apparatus)
+fire_data_model <- fire_data_model %>%
+  mutate(severity_bin = cut(number_of_responding_apparatus,
+                            breaks = c(0, 10, 20, 30, 40, 50, Inf),
+                            labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
+                            right = TRUE))
+
+# Step 2: Create a frequency distribution of severity bins
+severity_dist <- fire_data_model %>%
+  group_by(severity_bin) %>%
+  summarise(frequency = n()) %>%
+  ungroup()
+
+# Log-log plot
+ggplot(severity_dist, aes(x = as.numeric(severity_bin), y = frequency)) +
+  geom_point() +
+  scale_x_continuous(breaks = 1:length(levels(fire_data$severity_bin)), 
+                     labels = levels(fire_data$severity_bin)) +
+  scale_y_log10() +
+  labs(x = "Severity Bins (Apparatus Count)", 
+       y = "Frequency (Log Scale)", 
+       title = "Log-Log Plot of Fire Severity Distribution") +
+  theme_minimal()
+
+# Step 3: Fit a Pareto distribution to the severity data
+# Fit Pareto distribution to the column 'number_of_responding_apparatus' (or 'number_of_responding_personnel')
+fit_pareto <- fitdist(fire_data_model$number_of_responding_apparatus, "pareto")
+
+# Summary of Pareto fit
+summary(fit_pareto)
+
+# Step 4: Examine excesses above a threshold (e.g., severity > 20)
+threshold <- 20  # Chosen threshold for extreme events
+extreme_events <- fire_data %>%
+  filter(number_of_responding_apparatus > threshold) %>%
+  summarise(excess = mean(number_of_responding_apparatus - threshold))
+
+# Plot excesses above threshold
[…]

### 2024-12-03 scripts/06-define_severity.R
Defining severity

@@ -1,0 +1,110 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(rpart)
+
+#### Read data ####
+fire_data_model <- read_csv("data/02-analysis_data/tfs_analysis_data")
+
+#### Quantify Severity ####
+
+# Load libraries
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(fitdistrplus)
+
+# Step 1: Choose a column to quantify fire severity, (number_of_responding_apparatus)
+fire_data_model <- fire_data_model %>%
+  mutate(severity_bin = cut(number_of_responding_apparatus,
+                            breaks = c(0, 10, 20, 30, 40, 50, Inf),
+                            labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
+                            right = TRUE))
+
+# Step 2: Create a frequency distribution of severity bins
+severity_dist <- fire_data_model %>%
+  group_by(severity_bin) %>%
+  summarise(frequency = n()) %>%
+  ungroup()
+
+# Log-log plot
+ggplot(severity_dist, aes(x = as.numeric(severity_bin), y = frequency)) +
+  geom_point() +
+  scale_x_continuous(breaks = 1:length(levels(fire_data$severity_bin)), 
+                     labels = levels(fire_data$severity_bin)) +
+  scale_y_log10() +
+  labs(x = "Severity Bins (Apparatus Count)", 
+       y = "Frequency (Log Scale)", 
+       title = "Log-Log Plot of Fire Severity Distribution") +
+  theme_minimal()
+
+# Step 3: Fit a Pareto distribution to the severity data
+# Fit Pareto distribution to the column 'number_of_responding_apparatus' (or 'number_of_responding_personnel')
+fit_pareto <- fitdist(fire_data_model$number_of_responding_apparatus, "pareto")
+
+# Summary of Pareto fit
+summary(fit_pareto)
+
+# Step 4: Examine excesses above a threshold (e.g., severity > 20)
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,110 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(rpart)
-
-#### Read data ####
-fire_data_model <- read_csv("data/02-analysis_data/tfs_analysis_data")
-
-#### Quantify Severity ####
-
-# Load libraries
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(fitdistrplus)
-
-# Step 1: Choose a column to quantify fire severity, (number_of_responding_apparatus)
-fire_data_model <- fire_data_model %>%
-  mutate(severity_bin = cut(number_of_responding_apparatus,
-                            breaks = c(0, 10, 20, 30, 40, 50, Inf),
-                            labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
-                            right = TRUE))
-
-# Step 2: Create a frequency distribution of severity bins
-severity_dist <- fire_data_model %>%
-  group_by(severity_bin) %>%
-  summarise(frequency = n()) %>%
-  ungroup()
-
-# Log-log plot
-ggplot(severity_dist, aes(x = as.numeric(severity_bin), y = frequency)) +
-  geom_point() +
-  scale_x_continuous(breaks = 1:length(levels(fire_data$severity_bin)), 
-                     labels = levels(fire_data$severity_bin)) +
-  scale_y_log10() +
-  labs(x = "Severity Bins (Apparatus Count)", 
-       y = "Frequency (Log Scale)", 
-       title = "Log-Log Plot of Fire Severity Distribution") +
-  theme_minimal()
-
-# Step 3: Fit a Pareto distribution to the severity data
-# Fit Pareto distribution to the column 'number_of_responding_apparatus' (or 'number_of_responding_personnel')
-fit_pareto <- fitdist(fire_data_model$number_of_responding_apparatus, "pareto")
-
-# Summary of Pareto fit
-summary(fit_pareto)
-
-# Step 4: Examine excesses above a threshold (e.g., severity > 20)
[…]

### 2024-12-03 scripts/06-define_severity.R
define severity

@@ -12,99 +12,76 @@
 library(tidyverse)
 library(rstanarm)
 library(rpart)
-
-#### Read data ####
-fire_data_model <- read_csv("data/02-analysis_data/tfs_analysis_data")
-
-#### Quantify Severity ####
-
-# Load libraries
 library(ggplot2)
 library(dplyr)
 library(tidyr)
 library(fitdistrplus)
 
-# Step 1: Choose a column to quantify fire severity, (number_of_responding_apparatus)
+#### Read data ####
+fire_data_model <- read_csv("data/02-analysis_data/updated_tfs_analysis_data")
+
+#### Quantify Severity ####
+
+### Apparatus
 fire_data_model <- fire_data_model %>%
   mutate(severity_bin = cut(number_of_responding_apparatus,
                             breaks = c(0, 10, 20, 30, 40, 50, Inf),
                             labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
                             right = TRUE))
-
-# Step 2: Create a frequency distribution of severity bins
 severity_dist <- fire_data_model %>%
   group_by(severity_bin) %>%
   summarise(frequency = n()) %>%
   ungroup()
-
-# Log-log plot
-ggplot(severity_dist, aes(x = as.numeric(severity_bin), y = frequency)) +
-  geom_point() +
-  scale_x_continuous(breaks = 1:length(levels(fire_data$severity_bin)), 
-                     labels = levels(fire_data$severity_bin)) +
-  scale_y_log10() +
-  labs(x = "Severity Bins (Apparatus Count)", 
-       y = "Frequency (Log Scale)", 
-       title = "Log-Log Plot of Fire Severity Distribution") +
-  theme_minimal()
-
-# Step 3: Fit a Pareto distribution to the severity data
-# Fit Pareto distribution to the column 'number_of_responding_apparatus' (or 'number_of_responding_personnel')
-fit_pareto <- fitdist(fire_data_model$number_of_responding_apparatus, "pareto")
-
-# Summary of Pareto fit
-summary(fit_pareto)
-
-# Step 4: Examine excesses above a threshold (e.g., severity > 20)
-threshold <- 20  # Chosen threshold for extreme events
-extreme_events <- fire_data %>%
-  filter(number_of_responding_apparatus > threshold) %>%
-  summarise(excess = mean(number_of_responding_apparatus - threshold))
-
[…]

### 2024-12-03 scripts/06-define_severity.R
analysis differnt factor's relation to severity

@@ -49,10 +49,9 @@ fire_data_model <- fire_data_model %>%
   mutate(
     # Define Casualty Severity
     casualty_severity = case_when(
-      total_casualties == 0 | total_casualties == 1 ~ "Low", 
-      total_casualties >= 2 & total_casualties <= 4 ~ "Medium",
-      total_casualties >= 5 ~ "High",
-      TRUE ~ "Low"
+      total_casualties == 0 ~ "Low", 
+      total_casualties >= 4 ~ "High",
+      TRUE ~ "Medium"
     ),
     
     # Define Final Severity based on Apparatus, Dollar Loss, and Casualties
@@ -64,15 +63,15 @@ fire_data_model <- fire_data_model %>%
       
       # Apparatus severity classifications based on bins
       (casualty_severity == "Low" 
-       | severity_bin %in% c("[1,10]", "[11,20]") 
+       | severity_bin %in% c("[1,10]") 
        | log_estimated_dollar_loss <= loss_quantiles[1]) ~ "Low",
       
       (casualty_severity == "Medium" 
-       | severity_bin %in% c("[21,30]", "[31,40]") 
+       | severity_bin %in% c("[11,20]", "[21,30]") 
        | log_estimated_dollar_loss <= loss_quantiles[2]) ~ "Medium",
       
       (casualty_severity == "High" 
-       | severity_bin %in% c("[41,50]", ">50") 
+       | severity_bin %in% c("[31,40]", "[41,50]", ">50") 
        | log_estimated_dollar_loss > loss_quantiles[2]) ~ "High",
       
       TRUE ~ "Low" 
@@ -84,4 +83,112 @@ fire_data_model <- fire_data_model %>%
 write_csv(fire_data_model, "data/02-analysis_data/severity_tfs_analysis_data")
 
 
+#### Understand what is related to Severity ####
+
+## Factor of Ignition Source ##
+# Convert Severity and ignition_source_grouped to factors
+fire_data_model <- fire_data_model %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    ignition_source_grouped = factor(ignition_source_grouped)  # Convert ignition_source_grouped to factor
+  )
+
+# Calculate the top 5 ignition sources for each Severity group
+top_ignition_sources <- fire_data_model %>%
+  group_by(Severity, ignition_source_grouped) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
[…]

### 2024-12-03 scripts/08-current-system.R
understand current auto fire solving system

@@ -1,0 +1,135 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(rpart)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(fitdistrplus)
+
+#### Read data ####
+fire_data_current <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+
+#### fire_alarm_system_operation ####
+
+# Convert Severity and alarm_fire_data_filtered to factors
+fire_data_current <- fire_data_current %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    fire_alarm_system_operation = factor(fire_alarm_system_operation) 
+  )
+
+# Calculate the top 5 Status of fire Alarm for each Severity group
+top_fire_alarm_system_operation <- fire_data_current %>%
+  group_by(Severity, fire_alarm_system_operation) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
+# Filter the dataset to include only the top status of fire alarm for each severity
+alarm_fire_data_filtered <- fire_data_current %>%
+  semi_join(top_fire_alarm_system_operation, by = c("Severity", "fire_alarm_system_operation"))
+
+# Create a stacked bar plot with percentages
+ggplot(alarm_fire_data_filtered, aes(x = Severity, fill = fire_alarm_system_operation)) +
+  geom_bar(position = "fill", stat = "count") + 
+  labs(
+    title = "Proportional Distribution of fire_alarm_system_operation by Severity",
+    x = "Severity",
+    y = "Proportion",
+    fill = "fire_alarm_system_operation"
+  ) +
+  scale_y_continuous(labels = scales::percent) +  # Format y-axis as percentages
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 0, hjust = 0.5),  # X-axis label alignment
+    strip.text = element_text(size = 12)  # Adjust facet label size
+  ) 
+
[…]

### 2024-12-03 scripts/07-model_severity.R
original tried xgboost, really bad :(

@@ -10,10 +10,78 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(dplyr)
+library(caret)
+library(xgboost)
 
-#### Load data ####
-# [...ADD CODE HERE...]
+#### Read data ####
+fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+
+#### Prepare Data ####
+fire_data_model$area_of_origin_grouped <- as.factor(fire_data_model$area_of_origin_grouped)
+fire_data_model$ignition_source_grouped <- as.factor(fire_data_model$ignition_source_grouped)
+fire_data_model$month <- as.factor(fire_data_model$month)
+
+# Encode the target variable (Severity) as a binary outcome for severity (high = 1, others = 0)
+fire_data_model <- fire_data_model %>%
+  mutate(Severity_high = ifelse(Severity == "High", 1, 0))
+fire_data_model <- fire_data_model[, c(
+  "area_of_origin_grouped", 
+  "ignition_source_grouped", 
+  "month",
+  "Severity_high"
+)]
 
 
+#Split into test and trannig set
+set.seed(123)
+trainIndex <- createDataPartition(fire_data_model$Severity_high, p = 0.8, list = FALSE)
+train_data <- fire_data_model[trainIndex, ]
+test_data <- fire_data_model[-trainIndex, ]
+
+train_matrix <- model.matrix(Severity_high ~ . - 1, data = train_data)
+test_matrix <- model.matrix(Severity_high ~ . - 1, data = test_data)
+
+### Model (XGBoost) Traning####
+
+xgb_model <- xgboost(data = train_matrix, label = train_data$Severity_high, 
+                     objective = "binary:logistic", 
+                     nrounds = 100, 
+                     max_depth = 6, 
+                     eta = 0.3, 
+                     subsample = 0.8, 
+                     colsample_bytree = 0.8, 
+                     eval_metric = "logloss")
+
+# View model summary
+print(xgb_model)
+
+#### Model Evalaution ####
+predictions <- predict(xgb_model, test_matrix)
+predictions_class <- factor(predictions_class, levels = levels(factor(test_data$Severity_high)))
+conf_matrix <- confusionMatrix(predictions_class, factor(test_data$Severity_high))
+print(conf_matrix)
+
+
+# ROC Curve
+library(ROCR)
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R
more on EDA, study correaltion

@@ -470,7 +470,7 @@ incident_table <- incident_count_by_year_month %>%
   pivot_wider(names_from = month, values_from = count, values_fill = 0)
 
 # Print the table
-view(incident_table)
+# View(incident_table)
 
 
 ## 13. tfs_arrival_time ##
@@ -508,6 +508,10 @@ ggplot(fire_data, aes(x = response_time)) +
   xlim(0, 15) 
 #Summary Table
 summary(fire_data$response_time)
+
+#### Save new analysis data ####
+fire_data$total_casualties <- fire_data$civilian_casualties + fire_data$tfs_firefighter_casualties
+write_csv(fire_data, "data/02-analysis_data/updated_tfs_analysis_data")
 
 
 #### Understand Potential Indicators of Severity ####
@@ -716,27 +720,5 @@ ggplot(fire_data_clean_loss, aes(x = estimated_dollar_loss)) +
 
 
 
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-####
 
 

### 2024-12-03 scripts/07-model_severity.R
modelling with logistic regression

@@ -18,19 +18,38 @@ library(xgboost)
 fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
 
 #### Prepare Data ####
-fire_data_model$area_of_origin_grouped <- as.factor(fire_data_model$area_of_origin_grouped)
-fire_data_model$ignition_source_grouped <- as.factor(fire_data_model$ignition_source_grouped)
-fire_data_model$month <- as.factor(fire_data_model$month)
 
-# Encode the target variable (Severity) as a binary outcome for severity (high = 1, others = 0)
 fire_data_model <- fire_data_model %>%
-  mutate(Severity_high = ifelse(Severity == "High", 1, 0))
+  mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
+
 fire_data_model <- fire_data_model[, c(
   "area_of_origin_grouped", 
   "ignition_source_grouped", 
-  "month",
-  "Severity_high"
+  "Severity_high",
+  "month"
 )]
+fire_data_model$month <- as.factor(fire_data_model$month)
+fire_data_model_origin_encoded <- model.matrix(~ area_of_origin_grouped - 1, data = fire_data_model)
+fire_data_model_source_encoded <- model.matrix(~ ignition_source_grouped - 1, data = fire_data_model)
+
+fire_data_model <- cbind(fire_data_model, fire_data_model_source_encoded, fire_data_model_origin_encoded)
+
+fire_data_model <- fire_data_model %>%
+  rename_all(~ gsub(" ", "_", .)) %>%
+  rename_all(~ gsub("-", "_", .)) %>%
+  rename_all(~ gsub(",", "_", .)) %>%
+  rename_all(~ gsub(":", "_", .))
+
+
+library(randomForest)
+# Fit random forest model
+rf_model <- randomForest(Severity_high ~ ., data = fire_data_model, importance = TRUE)
+
+# Print the model summary
+print(rf_model)
+
+# View feature importance
+importance(rf_model)
 
 
 #Split into test and trannig set
@@ -39,43 +58,21 @@ trainIndex <- createDataPartition(fire_data_model$Severity_high, p = 0.8, list =
 train_data <- fire_data_model[trainIndex, ]
 test_data <- fire_data_model[-trainIndex, ]
 
-train_matrix <- model.matrix(Severity_high ~ . - 1, data = train_data)
[…]

### 2024-12-04 scripts/07-model_severity.R
tried random forest, not working :(

@@ -19,60 +19,41 @@ fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
 
 #### Prepare Data ####
 
-fire_data_model <- fire_data_model %>%
-  mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
-
 fire_data_model <- fire_data_model[, c(
   "area_of_origin_grouped", 
   "ignition_source_grouped", 
-  "Severity_high",
+  "Severity",
   "month"
 )]
-fire_data_model$month <- as.factor(fire_data_model$month)
-fire_data_model_origin_encoded <- model.matrix(~ area_of_origin_grouped - 1, data = fire_data_model)
-fire_data_model_source_encoded <- model.matrix(~ ignition_source_grouped - 1, data = fire_data_model)
 
-fire_data_model <- cbind(fire_data_model, fire_data_model_source_encoded, fire_data_model_origin_encoded)
-
-fire_data_model <- fire_data_model %>%
-  rename_all(~ gsub(" ", "_", .)) %>%
-  rename_all(~ gsub("-", "_", .)) %>%
-  rename_all(~ gsub(",", "_", .)) %>%
-  rename_all(~ gsub(":", "_", .))
-
-
-library(randomForest)
-# Fit random forest model
-rf_model <- randomForest(Severity_high ~ ., data = fire_data_model, importance = TRUE)
-
-# Print the model summary
-print(rf_model)
-
-# View feature importance
-importance(rf_model)
-
-
-#Split into test and trannig set
-set.seed(123)
-trainIndex <- createDataPartition(fire_data_model$Severity_high, p = 0.8, list = FALSE)
+trainIndex <- createDataPartition(fire_data_model$Severity, p = 0.8, list = FALSE)
 train_data <- fire_data_model[trainIndex, ]
 test_data <- fire_data_model[-trainIndex, ]
 
-### Model (Logistic) Traning####
-table(train_data$Severity_high)
-weights <- ifelse(train_data$Severity_high == 1, 10, 1)
+train_data$Severity <- factor(train_data$Severity, levels = c("Low", "Medium", "High"))
 
-logit_model <- glm(Severity_high ~ ., data = train_data, family = binomial, weights = weights)
-summary(logit_model)
-
-## Drop paramter based on p-vlaue ##
-model_updated <- glm(
-  Severity_high ~ 
-    
[…]

### 2024-12-04 scripts/07-model_severity.R
Update with logtsic model

@@ -19,47 +19,61 @@ fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
 
 #### Prepare Data ####
 
+fire_data_model <- fire_data_model %>%
+  mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
+
 fire_data_model <- fire_data_model[, c(
   "area_of_origin_grouped", 
   "ignition_source_grouped", 
-  "Severity",
-  "month"
+  "Severity_high"
 )]
 
-trainIndex <- createDataPartition(fire_data_model$Severity, p = 0.8, list = FALSE)
+fire_data_model$ignition_source_grouped <- as.factor(fire_data_model$ignition_source_grouped)
+fire_data_model$area_of_origin_grouped <- as.factor(fire_data_model$area_of_origin_grouped)
+
+fire_data_model$ignition_source_grouped <- gsub(" ", "_", fire_data_model$ignition_source_grouped)
+fire_data_model$area_of_origin_grouped <- gsub(",", "_", fire_data_model$area_of_origin_grouped)
+fire_data_model$area_of_origin_grouped <- gsub(" ", "_", fire_data_model$area_of_origin_grouped)
+fire_data_model$area_of_origin_grouped <- gsub("__", "_", fire_data_model$area_of_origin_grouped)
+
+
+trainIndex <- createDataPartition(fire_data_model$Severity_high, p = 0.8, list = FALSE)
 train_data <- fire_data_model[trainIndex, ]
 test_data <- fire_data_model[-trainIndex, ]
 
-train_data$Severity <- factor(train_data$Severity, levels = c("Low", "Medium", "High"))
+### Model (Logistic) Traning####
+logistic_model <- glm(Severity_high ~ area_of_origin_grouped + ignition_source_grouped, data = train_data, family = binomial)
+summary(logistic_model)
 
-tune_grid <- expand.grid(
-  mtry = c(2, 3, 4)# Number of variables randomly selected at each split
-)
 
-train_control <- trainControl(method = "cv", number = 5)
-# Calculate weight
-freq <- table(train_data$Severity)
-class_weights <- 1 / freq
-class_weights
+## Drop paramter based on p-vlaue ##
+# Check for correaltion
+library(car)
+vif(logistic_model)
 
-tuned_rf_model <- train(Severity ~ area_of_origin_grouped + ignition_source_grouped + month, 
-                        data = train_data, 
-                        method = "rf", 
-                        trControl = train_control, 
[…]

### 2024-12-04 scripts/06-define_severity.R
script update

@@ -82,6 +82,8 @@ fire_data_model <- fire_data_model %>%
 #### Save Data ####
 write_csv(fire_data_model, "data/02-analysis_data/severity_tfs_analysis_data")
 
+fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+
 
 #### Understand what is related to Severity ####
 

### 2024-12-04 paper/paper.qmd
Data and Model

@@
 date-format: long
-abstract: "This study investigates different severity levels of Toronto fire incidents based on financial and casualty losses to support efficient resource allocation for Toronto Fire Service (TFS).
+abstract: "This study investigates different severity levels of Toronto fire incidents to support efficient resource allocation for Toronto Fire Service (TFS).
 The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters.
@@
 The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters.
+
 The research identifies specific high-priority response criteria that can improve TFS operations.
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+We use the statistical programming language R [@citeR]to investigate data about fire incidents[@TFSfireedata], downloaded from the Open Data Toronto Portal: https://open.toronto.ca/dataset/fire-incidents/.
+The dataset was last updated on December 31, 2023.
+It includes fire incidents following the Ontario Fire Marshal's (OFM) definition, with descriptions according to OFM standards.
+The dataset also provides detailed information on Toronto Fire Services' (TFS) responses to each fire incident.
+The data is mostly complete, though some incidents are still under investigation and categorized as "no loss outdoor fire." The dataset includes information on the origin of fires, impacts on civilians and buildings, as well as geographical data.
+Additionally, it contains details on TFS response times and actions taken.
+Overall, the dataset is comprehensive, informative, and provides valuable insights for my later study.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The dataset is clearly structured, with the first column as ‘\_id,’ which is a unique identifier for each record in the database.
+There are 43 columns in total, containing information ranging from Incident Type to Infrastructure.
+These columns can be mainly divided into four categories based on my understanding:
 
-## Outcome variables
+1.
+**Incident Type and Cause**: This category contains information directly related to each fire incident, including descriptive details about the incident, its nature, and origin.
+It covers aspects such as where the fire started, the materials involved, and potential causes.
+These variables are crucial for predicting fire behavior, and include columns like 'area of origin,' 'level of origin,' and 'smoke spread.'
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+2.
+**Impact on Humans and the Economy**: These variables describe the direct outcomes of fire incidents, including financial and environmental impacts, as well as effects on human life.
+This category is key for defining the severity of fires.
+Variables include: firefighter casualties, civilian casualties, building status after the fire, and the number of displaced people.
+
+3.
+**Response Evaluation**: This category records TFS's reactions to each reported fire incident.
+It provides valuable insight into the effectiveness of the current system and helps when predicting fire severity.
+Variables include: 'fire control time,' 'incident location,' 'number of responding personnel,' and 'TFS arrival times.'
+
+4.
+**Safety Infrastructure**: This category includes data related to the status of existing infrastructure during the fire incident.
+It helps us understand the current system and supports decision-making for future improvements.
+Variables include: presence of fire alarm systems, sprinkler systems, and their operational status.
+
+This structured approach allows us to analyze and understand various aspects of fire incidents in Toronto.
+All variables in the dataset have the potential to provide valuable insights and perspectives for various analyses, contributing to fire prevention and safety improvements.
+However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
+A more detailed explanation will be provided in the Appendix.
+
+I have selected 17 variables to begin my exploration of the dataset:
+
+1.  **\_id**: A unique identifier for each record, essential for distinguishing between incidents and for tracking and referencing.
+
+2.  **Area_of_Origin**: Represents the fire’s origin location, useful for analyzing geographical patterns and fire causes.
+
+3.  **Building_Status**: Indicates whether the building was occupied, under construction, or vacant, providing insights into fire severity and response needs.
+
+4.  **Business_Impact**: Classifies the economic impact of the fire on businesses, helping to understand disruptions and guide recovery efforts.
+
+5.  **Civilian_Casualties**: The count of civilian casualties, reflecting the human impact and assessing fire safety and evacuation effectiveness.
+
+6.  **Estimated_Dollar_Loss**: The monetary loss due to fire damage, key for assessing the economic burden and informing prevention efforts.
+
+7.  **Exposures**: The number of secondary fires caused by the primary incident, useful for evaluating the fire's spread and containment success.
+
+8.  **Final_Incident_Type**: The final classification of the fire (e.g., accidental, arson), critical for identifying causes and informing preventive strategies.
+
+9.  **Fire_Alarm_System_Operation**: Indicates whether the fire alarm system functioned properly, evaluating the system’s effectiveness in alerting and activating suppression.
+
+10. **Ignition_Source**: The source of ignition (e.g., electrical faults, arson), crucial for identifying fire causes and guiding safety measures.
+
+11. **Number_of_Responding_Appliances**: The number of fire trucks and equipment dispatched, reflecting the scale of the incident and resource allocation.
+
+12. **Number_of_Responding_Personnel**: The number of fire personnel involved, providing insight into the severity of the fire and the required response intensity.
+
+13. **Possible_Cause**: The potential cause of the fire, useful for identifying trends and informing fire prevention and safety campaigns.
+
+14. **Sprinkler_System_Presence**: Indicates if a sprinkler system was present, highlighting its impact on fire severity and potential casualties.
+
+15. **TFS_Alarm_Time**: The time when the fire service was notified, key for analyzing response time and minimizing damage.
+
+16. **TFS_Arrival_Time**: The time when the first fire service unit arrived, crucial for evaluating response speed and fire control effectiveness.
+
+17. **TFS_Firefighter_Casualties**: The number of firefighter casualties, useful for assessing the safety of firefighting operations and improving risk management.
+
+By focusing on these 17 variables, I aim to uncover meaningful patterns and insights that can support fire safety improvements across the city.
+
+## Data Cleaning
+
+After loading the raw data, I selected the columns of interest as explained in the previous section.
+These columns were then renamed to lowercase, and spaces were replaced with underscores to ensure consistency.
+After reviewing the data, I kept in mind that this study focuses on the relationship between ignition sources and fire severity.
+As a result, I removed rows with "Undetermined" ignition sources.
+
+Since there are time-related variables in the dataset, I converted them to the POSIX format.
+Additionally, I noticed the presence of several missing (NA) and null values.
+Given that the dataset contains numerous categorical variables, I considered how to handle missing data differently for these types.
+I first examined whether any columns contained too little information to provide valuable insights into fire severity or its predictors.
+After calculating, I decided to remove columns with more than 20% missing data, carefully considering their relevance to the study.
+
+Variables removed:
+
+1.
+**Building_Status**: This column provides information on the condition of the building, but it is not directly related to the core objectives of the study, which focus on fire severity, casualties, and financial losses.
+With more than 20% missing data, it was deemed unnecessary.
+
+2.
+**Business_Impact**: While this column tracks the impact of the fire on businesses, it overlaps with the **Estimated_Dollar_Loss** column, which provides more comprehensive financial data.
+Due to the high missing rate, this column was removed.
+
+3.
+**Exposures**: This column captures the number of exposure fires but does not directly contribute to understanding fire severity or its impact on casualties and losses.
+With more than 20% missing data, it was excluded from the analysis.
+
+Kept variables: The remaining variables are essential for the completeness of the analysis. For categorical data entries, missing values are replaced with “unreported.”
+
+-   **Fire_Alarm_System_Operation**: This variable is crucial because the study aims to understand fire infrastructure.
+
+-   **Sprinkler_System_Presence**: This is also an important variable for understanding fire safety. Given that there are fewer variables under "Safety Infrastructure," it is essential to retain them.
+
+To maintain data quality, rows with more than 30% missing data are removed.
+After ensuring that there are no missing values in the dataset, a new variable, Response_Time, is created to calculate the time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time.
+This new variable will help analyze fire response efficiency and is a key indicator for the study.
 
@@
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
-Talk way more about it.
+## Data Explore
 
-## Predictor variables
+**Understanding Variables of Interest**:\
+To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics.
+The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
+Below are a few important variables considered:
 
-Add graphs, tables and text.
+1.  **Area of Origin**:\
+    This is a categorical variable that indicates the location within a building or structure where a fire started.
+Initially, this variable had over 100 categories, so the goal was to categorize these locations into broader groups for easier interpretation and analysis.
+I grouped the areas into categories such as "Residential and Living Spaces," "Cooking and Dining Areas," and "Storage and Utility Areas," among others.
+These groups were then visualized to better understand the most common fire origins within the dataset (see Table 1: \[code\]).
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+2.  **Civilian Casualties**:\
+    This is a numerical variable that provides insight into the severity of the incidents.
+Basic descriptive statistics (e.g., mean, median, and range) were calculated using the `summary()` function.
+Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see \[code\]).
+
+3.  **Final Incident Type**:\
+    This is a categorical variable that classifies the type of incident (e.g., fire, explosion).
+A bar plot was created to visualize the distribution of different incident types.
+The frequency of each incident type is also displayed in the table (see \[code\]).
+
+4.  **Ignition Source**:\
+    This variable provides information about the cause of the fire.
+Categories of ignition sources, such as "Incandescent Lamp" or "Vehicle - Mechanical," were listed.
+The distribution of ignition sources was visualized with a bar plot (see \[code\]).
+
+**Understanding Potential Indicators of Severity and Their Replashioships**:\
+To explore the relationship between financial losses and casualties, incidents were grouped into financial loss and casualty categories.
+A heatmap below illustrates the interaction between these two indicators.
+
+.
+
+```{r}
+fire_data <- fire_data %>%
+  mutate(
+    loss_group = cut(
+      estimated_dollar_loss,
+      breaks = c(0, 1000, 10000, 50000, 500000, Inf),
+      labels = c("<1k", "1k-10k", "10k-50k", "50k-500k", ">500k"),
+      right = FALSE
+    ),
+    casualty_group = cut(
+      total_casualties,
+      breaks = c(0, 1, 2, 5, Inf),
+      labels = c("0", "1", "2-5", ">5"),
+      right = FALSE
+    )
+  )
+
+# Summarize data and plot heatmap
+severity_table <- fire_data %>%
+  group_by(loss_group, casualty_group) %>%
+  summarize(count = n(), .groups = "drop")
+
+ggplot(severity_table, aes(x = loss_group, y = casualty_group, fill = count)) +
+  geom_tile(color = "white") +
+  scale_fill_gradient(low = "lightblue", high = "darkblue") +
+  labs(
+    title = "Heatmap of Financial Loss and Casualty Groups",
+    x = "Financial Loss Group",
+    y = "Casualty Group",
+    fill = "Incident Count"
+  ) +
+  theme_minimal()
+
+```
+
+The scatter plot below showcases the relationship between the number of responding apparatus and the logarithmic scale of financial losses.
+A Pearson correlation coefficient was calculated to assess the strength of this relationship.
+
+```{r}
+fire_data$Log_Dollar_Loss <- log(fire_data$estimated_dollar_loss + 1)
+
+# Scatter plot and correlation
+ggplot(fire_data, aes(x = Log_Dollar_Loss, y = number_of_responding_apparatus)) +
+  geom_point(color = 'blue', alpha = 0.5) +
+  labs(
+    title = 'Number of Responding Apparatus vs Log of Financial Loss',
+    x = 'Log of Financial Loss',
+    y = 'Number of Responding Apparatus'
+  ) +
+  theme_minimal()
+
+# Calculate correlation
+cor(fire_data$Log_Dollar_Loss, fire_data$number_of_responding_apparatus, use = "complete.obs")
+```
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+## Define Severity
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The severity of fire incidents was categorized based on three key factors: the number of responding fire apparatus, the estimated dollar loss, and the total number of casualties.
+First, the total number of responding apparatus from TFS was calculated and divided into bins after plotting and observing its distribution.
+Each incident was assigned to a bin based on how many apparatus responded, helping categorize incidents in terms of the scale of response.
 
-## Model set-up
+Next, the estimated dollar loss for each fire incident was log-transformed.
+This was necessary because the data was extremely right-skewed, with highly influential outliers.
+The transformation compressed the scale of the data and managed extreme values better.
+The transformed data was stored in a new variable called `log_estimated_dollar_loss`.
+Quantiles (25th, 50th, 75th percentiles) were used to divide the dollar loss into categories to make severity thresholds more meaningful.
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+For casualty severity, the number of casualties, including civilians and firefighters, was categorized after observing its distribution. The categories were defined as:
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+-   Low: No casualties
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+-   Medium: 1 to 3 casualties
 
-### Model justification
+-   High: 4 or more casualties
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+Another key factor was the incident type.
+If the type of incident was determined to be an explosion, it was automatically classified as High Severity regardless of other factors.
+The final severity of fire incidents was determined based on the following logic:
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+-   If the incident type was an explosion, the severity was **High**.
+
+-   If casualty severity was "Low," apparatus used was in the lower bins (1-10 apparatus), and dollar loss was below the 25th percentile, the severity was **Low**.
+
+-   If casualty severity was "Medium," apparatus was in the mid-range (11-30 apparatus), or dollar loss was below the 50th percentile, the severity was **Medium**.
+
+-   If casualty severity was "High," apparatus response was in the higher bins (31-50 apparatus), or dollar loss was above the 50th percentile, the severity was **High**.
+
+The analysis shows that most incidents fall under the Low Severity category, which aligns with the reality of everyday fire service responses.
+
+```{r}
+
+```
+
+### Set Up Model For Predicting Severities
+
+In this analysis, I aimed to predict the likelihood of high-severity fire incidents using a logistic regression model.
+The dataset includes several predictor variables that show potential connections to fire incident severities, as identified during prior data visualization.
+Among these variables, `area_of_origin_grouped` and `ignition_source_grouped` are categorical variables representing the area where the fire started and the source of ignition, respectively.
+The target variable for the model is the binary classification of fire severity, focusing on predicting high-severity incidents.
+
+To prepare the target variable, the `severity` variable calculated earlier was transformed into a binary variable called `Severity_high`, where incidents classified as "High Severity" were coded as `1`.
+This binary transformation aligns with the analysis goal of predicting high-severity incidents.
+Since both predictor variables are categorical, they were converted into factors to ensure compatibility with logistic regression.
+Additionally, redundant spaces and commas in the variable names were cleaned to standardize the data and improve model interpretability.
+
+To evaluate model performance, the dataset was split into a training set (80%) and a testing set (20%) using the `createDataPartition()` function from the `caret` package in R.
+
+### Model Explnation
+
+A logistic regression model was chosen because the outcome variable, Severity_high, is binary (0/1), and logistic regression is well-suited for binary classification tasks, especially when working with categorical predictor variables.
+The logistic regression model used the following predictors.
+The logistic regression model was initially fitted to the training dataset using the formula: ***Severity_high \~ area_of_origin_grouped + ignition_source_grouped.*** This initial model can be expressed as:
+
+$$
+\text{logit}(P) = \ln\left(\frac{P}{1-P}\right) = \beta_0 + \sum_{j=1}^{n_{\text{area}}} \beta_{j,\text{area}} \cdot X_{j,\text{area}} + \sum_{k=1}^{n_{\text{ignition}}} \beta_{k,\text{ignition}} \cdot X_{k,\text{ignition}}
+$$
+
+Where: $P$ is the probability of the event ($P(\text{Severity_high}=1)$).
+$\text{logit}(P)$ is the log-odds of the probability.
+$\beta_0$ is the intercept.
+$\beta_1, \beta_2, \dots, \beta_k$ are the coefficients of the predictor variables $X_1, X_2, \dots, X_k$.
+
+### Model Justification
+
+The initial logistic regression model was evaluated for the significance of the predictors using the p-values from the model summary.
+Since a high p-value suggests that the variable does not significantly contribute to the model in prediction, it was removed to simplify the model and improve interpretability.
+This decision follows the principle outlined by Occam's Razor \[\@blumer1987occam\], which proposes that people prefer simplicity.
+
+To check for multicollinearity among the predictors, the Variance Inflation Factor (VIF) was calculated.
+According to standard practice, if any predictors had a high VIF (greater than 5), they were considered for removal to avoid multicollinearity, which could distort the model's estimates.
+
+Based on the results, the variable *ignition_source_grouped* was removed from the model due to its non-significance (high p-value).
+The updated model retained only *area_of_origin_grouped*, which showed a significant relationship with the likelihood of high-severity fires.
+Model is:
+
+$$
+\text{logit}(P) = \ln\left(\frac{P}{1-P}\right) = \beta_0 + \sum_{j=1}^{n_{\text{area}}} \beta_{j,\text{area}} \cdot X_{j,\text{area}}
+$$
+
+The model's performance was also assessed using the Akaike Information Criterion (AIC), a measure of the model's goodness of fit.
+A lower AIC indicates a better-fitting model, and removing the insignificant variable *ignition_source_grouped* improved the AIC, providing more evidence that the updated, simpler model was a better fit for the data.
+
+### **Model Evaluation**
+
+The updated logistic regression model was evaluated using the receiver operating characteristic (ROC) curve, with predictions made on the test dataset.
+The curve was plotted to assess the model’s ability to discriminate between high and low-severity fire incidents.
+The area under the curve (AUC) is a key metric for model performance; higher AUC values indicate better model performance \[\@metz1978basic\].
+The simpler model, after removing insignificant predictors, combined with its performance on the test set, provides a better understanding of the relationship between fire characteristics and severity.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+### **Logistic Regression Model Results**
+
+The logistic regression model was fitted to predict the probability of high severity fire incidents based on the categorical variables **area of origin**.
+The final model retained only significant predictors based on p-values and collinearity diagnostics.
+Below are the key findings:
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-12-04 paper/paper.qmd
Some extra edits

@@
 #| message: false
+library(tidyverse)
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(lubridate)
+library(knitr)
 
-library(tidyverse)
-library(palmerpenguins)
+fire_data <- read.csv(here::here("~/TFS_Toronto_Fire_Incident/data/02-analysis_data/tfs_analysis_data"))
+
 ```
@@
 This new variable will help analyze fire response efficiency and is a key indicator for the study.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+(@fig-area-origin-distribution).
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-area-origin-distribution
+#| fig-cap: Distribution of the Area of Origin for fire incidents, categorized into various groups such as Residential, Office, Mechanical, and Industrial areas.
+The plot highlights the frequency of each area category, helping to understand the distribution of fire origins.
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+all_area <- unique(fire_data$area_of_origin)
+# List of Area of Origin
+area_of_origin <- c("22 - Sleeping Area or Bedroom (inc. patients room, dormitory, etc)", 
+#| echo: false
+
+all_area <- unique(fire_data$area_of_origin)
+# List of Area of Origin
+area_of_origin <- c("22 - Sleeping Area or Bedroom (inc. patients room, dormitory, etc)", 
+                    "55 - Mechanical/Electrical Services Room", 
+                    "28 - Office", 
+                    "24 - Cooking Area or Kitchen", 
+                    "81 - Engine Area", 
+                    "25 - Washroom or Bathroom (toilet, restroom/locker room)", 
+                    "73 - Parking Area, Parking Lot", 
+                    "44 - Trash, Rubbish Storage (inc garbage chute room, garbage/industr", 
+                    "50 - Basement/cellar (not partitioned)", 
+                    "64 - Porch or Balcony", 
+                    "78 - Attached Deck", 
+                    "86 - Passenger Area", 
+                    "52 - HVAC Equipment Room (furnace room, water heater closet, boiler)", 
+                    "21 - Living Area (e.g. living, TV, recreation, etc)", 
+                    "61 - Exterior Wall", 
+                    "62 - Roof", 
+                    "12 - Hallway, Corridor", 
+                    "83 - Electrical Systems", 
+                    "59 - Utility Shaft (eg. electrical wiring/phone, etc.)", 
+                    "79 - Other Outside Area", 
+                    "75 - Trash, rubbish area (outside)", 
+                    "89 - Other Vehicle Area", 
+                    "31 - Process Manufacturing (inc manf, prod assembly, repair)", 
+                    "53 - Chimney/Flue Pipe", 
+                    "26 - Sauna", 
+                    "99 - Undetermined (formerly 98)", 
+                    "87 - Trunk/Cargo Area", 
+                    "11 - Lobby, Entranceway", 
+                    "97 - Other - unclassified", 
+                    "67 - Concealed Floor Area", 
+                    "58 - Ducting - Exhaust (inc cooking, fumes, etc.)", 
+                    "29 - Electronic Equipment", 
+                    "51 - Elevator (includes shaft)", 
+                    "46 - Product Storage (inc products or materials awaiting manuf, assembly)", 
+                    "33 - Laboratory", 
+                    "84 - Fuel Systems (eg. fuel tank, etc.)", 
+                    "27 - Laundry Area", 
+                    "69 - Attic Area", 
+                    "45 - Supply Storage Room (inc maintenance/office/document storage, et", 
+                    "30 - Sales, Showroom Area", 
+                    "42 - Garage", 
+                    "66 - Concealed Ceiling Area", 
+                    "49 - Other Storage Area", 
+                    "13 - Stairway, Escalator", 
+                    "39 - Other Functional Area", 
+                    "71 - Open Area (inc lawn, field, farmyard, park, playing field, pier)", 
+                    "93 - Residential/Business: Other business area", 
+                    "72 - Court, Patio, Terrace", 
+                    "74 - Storage Area (outside)", 
+                    "56 - Conveyor Shaft or Chute (inc dumbwaiter, laundry chute, garbage", 
+                    "23 - Dining or Beverage Area (inc mess, canteen, lunchroom, cafeteria", 
+                    "92 - Residential/Business: Restaurant area", 
+                    "63 - Awning or Canopy", 
+                    "68 - Concealed Wall Area", 
+                    "41 - Closet (eg. clothes, broom, linen closet, etc.)", 
+                    "82 - Running Gear (inc wheels and braking systems, transmission system", 
+                    "60 - Other Building Services/Support Facilities", 
+                    "85 - Operator/Control Area", 
+                    "990 - Under Investigation", 
+                    "43 - Locker (apartment storage)", 
+                    "57 - Ducting - Heating, Air Conditioning", 
+                    "70 - Other Structural Area", 
+                    "35 - Performance Area (inc stage, rink, boxing ring, gym floor, altar)", 
+                    "65 - Crawl Space (includes sub-structure)", 
+                    "91 - Multiple Areas of Origin", 
+                    "18 - Covered Court, Atrium, mall concourse", 
+                    "36 - Backstage, dressing room", 
+                    "32 - Assembly Area (inc school room, spectator area, church, etc)", 
+                    "19 - Other Means of Egress", 
+                    "47 - Shipping/Receiving/Loading Platform", 
+                    "34 - Operating Room, Treatment or Examination Area", 
+                    "76 - Fuel Dispensing Area (outside)", 
+                    "48 - Records storage area (inc vaults)", 
+                    "54 - Incinerator Room")
+
+# Create a function to categorize area of origin
+categorize_area_of_origin <- function(area) {
+  # Residential and Living Spaces
+  if (grepl("Bedroom|Living Area|Dormitory|Hallway|Lobby|Basement|Closet|Sauna", area)) {
+    return("Residential and Living Spaces")
+  } 
+  # Cooking and Dining Areas
+  else if (grepl("Kitchen|Dining|Canteen|Lunchroom|Mess", area)) {
+    return("Cooking and Dining Areas")
+  } 
+  # Office and Workspace
+  else if (grepl("Office|Workspace|Lobby|Showroom|Laboratory|Assembly Area", area)) {
+    return("Office and Workspace")
+  } 
+  # Mechanical, HVAC, and Electrical Areas
+  else if (grepl("Mechanical|Electrical|HVAC|Boiler|Furnace", area)) {
+    return("Mechanical, HVAC, and Electrical Areas")
+  } 
+  # Storage and Utility Areas
+  else if (grepl("Storage|Concealed|Shaft|Utility|Warehouse|Garage", area)) {
+    return("Storage and Utility Areas")
+  } 
+  # Outdoor and Exterior Areas
+  else if (grepl("Parking|Exterior|Deck|Porch|Lawn|Field|Pier|Terrace", area)) {
+    return("Outdoor and Exterior Areas")
+  } 
+  # Vehicle and Transport Areas
+  else if (grepl("Vehicle|Trunk|Cargo|Engine|Garage", area)) {
+    return("Vehicle and Transport Areas")
+  } 
+  # Specialized Rooms
+  else if (grepl("Bathroom|Laundry|Shower|Incinerator|Sauna|Elevator", area)) {
+    return("Specialized Rooms")
+  } 
+  # Industrial and Manufacturing Areas
+  else if (grepl("Manufacturing|Assembly|Fuel Systems|Conveyor|Chute", area)) {
+    return("Industrial and Manufacturing Areas")
+  }
+
+  # Public and Performance Spaces
+  else if (grepl("Stage|Rink|Gym|Performance|Concourse|Court", area)) {
+    return("Public and Performance Spaces")
+  } 
+  # Miscellaneous and Unclassified
+  else {
+    return("Miscellaneous and Unclassified")
+  }
+}
+# Apply the categorization function to each area of origin
+fire_data$area_of_origin_grouped <- sapply(fire_data$area_of_origin, 
+                                           categorize_area_of_origin)
+# Plot
+ggplot(fire_data, aes(x = area_of_origin_grouped)) +
+  geom_bar(fill = "red") +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  ggtitle("Area of Origin Distribution") +
+  xlab("Area of Origin") +
+  ylab("Frequency") +
+  geom_text(stat = "count", aes(label = ..count..), vjust = 0.5)
+
 ```

### 2024-12-04 scripts/02-download_data.R
preamble update

@@ -4,7 +4,7 @@
 # Date: 29 November 2024
 # Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Plan data ans simulated
+# Pre-requisites: Plan data and simulated
 # Any other information needed? NO
 
 

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Explore data of variables in interest and plot them
-# Author: Maggie Zhang 
+# Purpose: Explore data of variables of interest and visualize their distributions
+# Author: Maggie Zhang
 # Date: 30 November 2024
 # Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: R version 4.0 or higher, installed packages: dplyr, ggplot2, tidyr, lubridate, knitr
+# Any other information needed: This script analyzes fire incident data from the TFS dataset. It categorizes and visualizes variables such as Area of Origin, Civilian Casualties, Estimated Dollar Loss, Final Incident Type, Fire Alarm System Operation, and others.
 
 
 #### Workspace setup ####

### 2024-12-04 scripts/06-define_severity.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: The purpose of this script is to analyze fire incident data, specifically quantifying severity based on multiple factors such as the number of responding apparatus, estimated dollar loss, and total casualties. The script also explores the relationships between severity and ignition sources, as well as the area of origin, through various visualizations.
+# Author: Maggie Zhang
+# Date: 02 December 2023
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Required libraries include 'tidyverse', 'rstanarm', 'rpart', 'ggplot2', 'dplyr', 'tidyr', and 'fitdistrplus'. Ensure these packages are installed.
+# Any other information needed? The dataset used here is fire incident data, and the focus is on classifying the severity of incidents based on multiple features, followed by visualizing relationships between severity and other factors.
 
 
 #### Workspace setup ####

### 2024-12-04 scripts/07-model_severity.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Replicated graphs and models for analyzing severity of fire incidents based on various factors, including ignition source and area of origin.
+# Author: Maggie Zhang
+# Date: 02 December 2023
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: tidyverse, dplyr, caret, xgboost, car, pROC
+# Any other information needed? The analysis involves logistic regression modeling to predict severity based on categorical variables, followed by model evaluation using ROC curves.
 
 
 #### Workspace setup ####

### 2024-12-04 scripts/08-current-system.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Analyze and visualize the relationship between fire incident severity and factors such as fire alarm system operation and sprinkler system presence. Includes creating stacked bar plots and separate plots to explore the proportional distribution and counts by severity.
+# Author: Maggie Zhang
+# Date: 02 December 2023
+# Contact: maggiey.zhang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: tidyverse, rstanarm, rpart, ggplot2, dplyr, tidyr, fitdistrplus
+# Any other information needed? The analysis involves creating visualizations to explore the distribution of fire alarm and sprinkler system statuses by severity.
 
 
 #### Workspace setup ####

### 2024-12-04 paper/paper.qmd
paper regenrate

@@
 ---
-title: "Analyzing Toronto Fire Incident Severity"
-subtitle: "Study Thorugh Financial and Casualty Loss to Optimize Resource Allocation for Toronto Fire Services"
+title: "Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response"
+subtitle: "Identifying Key Factors for Fire Severity and Evaluating the Effectiveness of Public Fire Response Systems"
 author: 
@@
 thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
-date: today
+date: "December 3, 2024"
 date-format: long
-abstract: "This study investigates different severity levels of Toronto fire incidents to support efficient resource allocation for Toronto Fire Service (TFS).
-The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters.
-The study identifies high-priority response criteria, including incidents originating from XXXX regions, occurring within defined time frames: XXXX, and involving XXXX as the ignition sources.
-These insights, derived from the modeling analysis, can guide TFS to enhance response times and prioritize tasks based on common high-risk indicators.
-This work aims to inform decision-making and optimize fire response operations to improve community safety in Toronto."
+abstract: "This study investigates fire severity levels in Toronto to improve resource allocation for Toronto Fire Services (TFS).
+The analysis identifies that higher-severity incidents are strongly correlated with incidents originating from mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires.
+The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment, are more likely to be severe.
+The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are successful in low and medium-severity fires, while sprinkler systems are often only partially operational.
+Results can guide TFS to enhance response times, prioritize resources, and optimize fire response operations, ultimately improving community safety in Toronto."
 format: pdf
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
-fire_data <- read.csv(here::here("~/TFS_Toronto_Fire_Incident/data/02-analysis_data/tfs_analysis_data"))
+fire_data <- read.csv(file = here::here("data/02-analysis_data/tfs_analysis_data"))
 
@@
 
-This study investigates different severity levels of Toronto fire incidents based on financial and casualty losses to support efficient resource allocation for Toronto Fire Service (TFS).
-The analysis reveals that higher-severity incidents correlate with greater financial losses and more significant human casualties, impacting both civilians and firefighters.
+The study aims to estimate the severity of fires in the Toronto region.
+A fire incident, as defined by the Ontario Fire Marshal (OFM), refers to any event involving fire, explosion, or other situations suspected to lead to significant outcomes, such as injury or death [@ontario_fire_marshal].
 
-The research identifies specific high-priority response criteria that can improve TFS operations.
-By examining the data, it has been found that incidents originating from XXX regions, occurring within defined time frames: XXX, and involving XXXX ignition sources exhibit higher risk profiles.
-This enables TFS to develop strategic response protocols that prioritize these types of incidents.
-The modeling analysis conducted in this study provides valuable insights into which factors contribute most significantly to incident severity, equipping TFS with the knowledge to enhance response times and allocate resources more effectively.
+The severity of a fire can be assessed through various factors, including the intensity of the flames, the size of the affected area, the duration of the fire, its impact on human activities, and its potential threat to animal habitats.
+There are numerous standards and considerations when discussing fire severity, but when it comes to prediction, it is unrealistic to consider all factors.
+Therefore, defining severity becomes the first task before any further analysis.
+In this paper, the severity of a fire is defined based on three aspects: casualty and financial loss, the type of incident, and the number of apparatuses dispatched by Toronto Fire Services (TFS) in response to the fire.
+Further explanation of how severity is defined will be provided later in the paper.
 
-The insights from this work aim to inform strategic decision-making processes and optimize fire response operations.
-By understanding and addressing the indicators of high-risk incidents, TFS can better safeguard the community and improve public safety throughout Toronto.
-This research is an essential step toward developing data-driven approaches that enhance both the efficiency of fire service operations and the overall resilience of the city.
+This research also identifies specific high-priority response criteria that can enhance TFS operations.
+By examining the data, it has been found that the area of origin of a fire is a significant predictor of fire severity.
+Fires originating from mechanical, HVAC, and electrical areas are highly correlated with severe incidents, whereas fires originating from cooking and heating areas are more commonly associated with low-severity incidents.
 
-Estimand paragraph
+Additionally, beyond investigating potential predictors of severe fire incidents, this study also explores the effectiveness of current public fire response mechanisms, such as fire alarms and sprinkler systems, in the Toronto region.
+It was discovered that sprinkler systems, across all levels of fire severity, were largely only partially operational.
+On the other hand, fire alarm systems were successfully activated in low and medium-severity fire incidents.
 
-Results paragraph
+The Office of the Fire Marshal (OFM) plays a crucial role in overseeing fire safety regulations in Ontario, ensuring fire protection, prevention, and public safety across the province.
+They are responsible for setting safety standards and improving public safety for all municipalities.
+Under the leadership of the OFM, municipal fire departments, such as Toronto Fire Services (TFS), allocate resources and develop plans to protect shared communities [@ontario_fire_marshal].
+Therefore, by analyzing data within their standards, identifying factors that contribute to high-risk fire incidents, and evaluating current operational mechanisms, TFS can gain valuable insights to develop strategic response protocols that prioritize these incidents.
+This knowledge will help both the public and fire departments enhance response times, allocate resources more effectively, and reduce the occurrence of lower-severity incidents.
 
-Why it matters paragraph
+The remainder of this paper is structured as follows.
+[@sec-data] describes the data used in the analysis, including the sources and key variables and their visualizations.[@sec-model] outlines the models applied to estimate fire severity and identify key factors.
+[@sec-results] presents the results of the analysis.
+[@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements.
+Finally, [@sec-appendixa] includes supplementary information and additional data details.
+[@sec-appendix-B] provided insight into Fire Incidence, about their sampling method and potential bias.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+In this paper, I used several R packages for data analysis and visualization.
+The `tidyverse`[@wickham2023tidyverse], `dplyr`[@wickham2023dplyr], and `tidyr`[@wickham2023tidyr] were used for data manipulation.
+`ggplot2`[@wickham2016ggplot2] facilitated data visualization, while `lubridate`[@grolemund2011lubridate] was used for date-time handling.
+`knitr`[@xie2023knitr] helped integrate code into reports.
+For modeling, I used `rstanarm`[@gelman2020rstanarm] and `modelsummary`[@fox2021modelsummary] for summaries, `caret`[@kuhn2023caret] for model training, and `xgboost`[@chen2016xgboost] for boosting.
+Additional tools like `rpart`[@therneau2015rpart], `fitdistrplus`[@delignette-muller2023fitdistrplus], and `pROC`[@robin2023pROC] supported partitioning, distribution fitting, and model evaluation.
 
@@
 
-We use the statistical programming language R [@citeR]to investigate data about fire incidents[@TFSfireedata], downloaded from the Open Data Toronto Portal: https://open.toronto.ca/dataset/fire-incidents/.
+We use the statistical programming language R [@citeR]to investigate data about fire incidents[@TFSfireedata], downloaded from the Open Data Toronto Portal: https://open.toronto.ca/dataset/fire-incidents/ .
 The dataset was last updated on December 31, 2023.
@@
 The dataset also provides detailed information on Toronto Fire Services' (TFS) responses to each fire incident.
-The data is mostly complete, though some incidents are still under investigation and categorized as "no loss outdoor fire." The dataset includes information on the origin of fires, impacts on civilians and buildings, as well as geographical data.
+The data is mostly complete, though some incidents are still under investigation and categorized as "no loss outdoor fire." The dataset includes information on the origin of fires, impacts on civilians and buildings, as well as geographical data (see @tbl-preview-data).
 Additionally, it contains details on TFS response times and actions taken.
 Overall, the dataset is comprehensive, informative, and provides valuable insights for my later study.
+
+```{r}
+#| label: tbl-preview-data
+#| tbl-cap: Preview of the Fire Incident data (https://open.toronto.ca/dataset/fire-incidents/) before cleaning, displaying the first six rows with columns decided first in interested for study.
+#| echo: false
+#| warning: false
+
+library(tidyverse)
+library(dplyr)
+library(lubridate)
+
+#### Load original data ####
+raw_data <- read_csv(file = here::here("data/01-raw_data/raw_fire_data.csv"))
+
+columns_to_keep <- c("_id", "Area_of_Origin", "Building_Status", "Business_Impact", 
+                     "Civilian_Casualties", "Estimated_Dollar_Loss", "Exposures", 
+                     "Final_Incident_Type", "Fire_Alarm_System_Operation", 
+                     "Ignition_Source",
+                     "Number_of_responding_apparatus", "Number_of_responding_personnel", 
+                     "Possible_Cause", "Sprinkler_System_Presence", 
+                     "TFS_Alarm_Time", "TFS_Arrival_Time", "TFS_Firefighter_Casualties")
+
+clean_data <- raw_data[, columns_to_keep]
+# Change columns name to lower case and ensure no space
+colnames(clean_data) <- tolower(gsub(" ", "_", colnames(clean_data)))
+
+### Preview data ###
+kable(head(clean_data, 6))
+```
 
@@
 However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
-A more detailed explanation will be provided in the Appendix.
+A more detailed explanation will be provided in the @sec-appendixa.
 
@@
 To maintain data quality, rows with more than 30% missing data are removed.
-After ensuring that there are no missing values in the dataset, a new variable, Response_Time, is created to calculate the time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time.
+After ensuring that there are no missing values in the dataset, a new variable, Response_Time, is created to calculate the time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time (see @fig-response-time-distribution).
 This new variable will help analyze fire response efficiency and is a key indicator for the study.
-(@fig-area-origin-distribution).
+
+```{r}
+#| label: fig-response-time-distribution
+#| fig-cap: Distribution of response times for fire incidents in Toronto, showing the frequency of various response times (in minutes).
+This visualization highlights the efficiency of the Toronto Fire Service's (TFS) response.
+It's almost normally distributed with a mean of around 5, and the spread of not too far.
+#| echo: false
+#| warning: false
+library(ggplot2)
+fire_data <- read.csv(here::here("data/02-analysis_data/tfs_analysis_data"))
+## 15. Response Time (Numerical) ##
+#Plot
+ggplot(fire_data, aes(x = response_time)) +
+  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
+  theme_minimal() +
+  ggtitle("Response Time Distribution") +
+  xlab("Response Time (minutes)") +
+  ylab("Frequency") +
+  stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 3) +
+  xlim(0, 15)
+
+```
+
+## Data Explore
+
+**Understanding Variables of Interest**:\
+To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics.
+The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
+Can check @sec-appendixa for visualization/summary of the distribution of all variables.
+Below are a few important variables considered:
+
+1.  **Area of Origin**:\
+    This is a categorical variable that indicates the location within a building or structure where a fire started.
+Initially, this variable had over 100 categories, so the goal was to categorize these locations into broader groups for easier interpretation and analysis.
+I grouped the areas into categories such as "Residential and Living Spaces," "Cooking and Dining Areas," and "Storage and Utility Areas," among others.
+These groups were then visualized to better understand the most common fire origins within the dataset (see @fig-area-origin-distribution).
+Among all categories, Cooking and Dinning area stands out to be the area which most fire originate from.
+Living space and Transportation area also has a great amount to incidence occured.
 
@@
 #| echo: false
-
-all_area <- unique(fire_data$area_of_origin)
-# List of Area of Origin
-area_of_origin <- c("22 - Sleeping Area or Bedroom (inc. patients room, dormitory, etc)", 
-#| echo: false
+#| warning: false
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  ggtitle("Area of Origin Distribution") +
   xlab("Area of Origin") +
@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-## Data Explore
-
-**Understanding Variables of Interest**:\
-To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics.
-The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
-Below are a few important variables considered:
-
-1.  **Area of Origin**:\
-    This is a categorical variable that indicates the location within a building or structure where a fire started.
-Initially, this variable had over 100 categories, so the goal was to categorize these locations into broader groups for easier interpretation and analysis.
-I grouped the areas into categories such as "Residential and Living Spaces," "Cooking and Dining Areas," and "Storage and Utility Areas," among others.
-These groups were then visualized to better understand the most common fire origins within the dataset (see Table 1: \[code\]).
-
 2.  **Civilian Casualties**:\
@@
 Basic descriptive statistics (e.g., mean, median, and range) were calculated using the `summary()` function.
-Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see \[code\]).
+Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see @tbl-civilian-casualties-summary).
+We can see that most of the time, there are no casualties.
+This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
+
+```{r}
+#| label: tbl-civilian-casualties-summary
+#| tbl-cap: Summary table showing the number of civilian casualties in fire incidents, including the frequency of each casualty count.
+This table summarizes the distribution of civilian casualties and provides an overview of the impact of fire incidents on civilians.
+#| echo: false
+#| warning: false
+library(knitr)
+
+#Overall Table
+civilian_casualties_df <- as.data.frame(table(fire_data$civilian_casualties))
+colnames(civilian_casualties_df) <- c("Number_of_Casualties", "Frequency")
+kable(civilian_casualties_df)
+
+```
 
@@
     This is a categorical variable that classifies the type of incident (e.g., fire, explosion).
-A bar plot was created to visualize the distribution of different incident types.
-The frequency of each incident type is also displayed in the table (see \[code\]).
+A bar plot was created to visualize the distribution of different incident types.(see @fig-final-incident-type-distribution).
+It's clear that there are more fire than explotion.
+
+```{r}
+#| label: fig-final-incident-type-distribution
+#| fig-cap: Bar plot illustrating the distribution of final incident types across fire incidents.
+The plot presents the frequency of each incident type, with the count displayed on top of the bars.
+We can see that there is much more fire than explotion.
+#| echo: false
+
+ggplot(fire_data, aes(x = final_incident_type)) +
+  geom_bar(fill = "purple") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1), 
+    plot.margin = margin(10, 10, 10, 10)  # Add space around the plot
+  ) +
+  scale_x_discrete(labels = function(x) sapply(x, function(y) paste(head(strsplit(y, " ")[[1]], 3), collapse = " "))) +
+  xlab("Final Incident Type") +
+  ylab("Frequency") +
+  geom_text(stat = "count", aes(label = ..count..), vjust = 0.5)
+
+```
 
@@
 Categories of ignition sources, such as "Incandescent Lamp" or "Vehicle - Mechanical," were listed.
-The distribution of ignition sources was visualized with a bar plot (see \[code\]).
+The distribution of ignition sources was visualized with a bar plot (see @fig-grouped-ignition-source-distribution).
+We can tell that fire-related and open flame is one of the biggest source of fire incidence; cooking and heating equipment are second place as sources, which align with @fig-area-origin-distribution result, where Ketchen is the place most fire incidence starts off.
+
+```{r}
+#| label: fig-grouped-ignition-source-distribution
+#| fig-cap: Bar plot illustrating the distribution of grouped ignition sources for fire incidents.
+The plot categorizes ignition sources into meaningful groups such as Lighting and Electrical Equipment, Cooking and Heating Equipment, and Mechanical Equipment, among others.
+The frequency of each group is displayed.
+Fire-related and Open Flame is the bigget ingnition source.
+#| echo: false
+#| warning: false
+all_sources <- unique(fire_data$ignition_source)
+ignition_sources <- c("51 - Incandescent Lamp - Light Bulb, Spotlight", 
+                      "23 - Distribution Equipment (includes panel boards, fuses, circuit br)", 
+                      "41 - Other Heating Equipment", "11 - Stove, Range-top burner", 
+                      "82 - Vehicle - Mechanical", "24 - Circuit Wiring - Copper", 
+                      "98 - Other", "81 - Vehicle - Electrical", "49 - Other Appliances", 
+                      "71 - Smoker's Articles (eg. cigarettes, cigars, pipes already ignited)", 
+                      "13 - Microwave", "77 - Matches or Lighters (unable to distinguish)", 
+                      "74 - Salamander", "12 - Oven", "47 - Refrigerator, Freezer (includes vending machine)", 
+                      "73 - Blow Torch, Bunsen Burner", "72 - Cutting/Welding Equipment", 
+                      "21 - Transformer", "55 - Candle", "29 - Extension Cord, Temporary Wiring", 
+                      "14 - Open Fired Barbeque - Fixed or Portable", "79 - Other Open Flame Tools/Smokers' Articles", 
+                      "84 - Other Mechanical", "36 - Fireplace - Masonry", "83 - Other Electrical", 
+                      "20 - Service/Utility Lines (includes power/hydro transmission lines)", 
+                      "92 - Open Fire (eg. camp fire, rubbish fire, etc.)", "17 - Wood burning stove", 
+                      "62 - Heat Treatment Equipment (eg. furnace, oven, kiln, quench tanks, etc.)", 
+                      "19 - Other Cooking Items (eg Toaster, Kettle, elec frying pan)", 
+                      "93 - Hot Ashes, Embers, Spark", "52 - Florescent Lamp (includes ballast)", 
+                      "64 - Chemical Processing Equipment (eg. reactors, distilling units, etc.)", 
+                      "43 - Clothes Dryer", "69 - Other Processing Equipment", 
+                      "88 - Multiple Ignition Source or Igniting Equipment (suspected arson)", 
+                      "75 - Matches (open flame)", "32 - Water Heater", 
+                      "26 - Terminations-Copper (incl receptacles, switches, lights)", "85 - Vehicle collision", 
+                      "28 - Cord, Cable for Appliance, Electrical Articles", "16 - Deep Fat Fryer", 
+                      "59 - Other Lighting Equipment", "30 - Other Electrical Distribution Item", 
+                      "76 - Lighters (open flame)", "33 - Space Heater - Fixed", "42 - Television, Radio, Stereo, Tape Recorder, etc.", 
+                      "39 - Chimney - Masonry", "34 - Space Heater - Portable", "44 - Iron, Pressing Machine", 
+                      "108 - Exposure, source other", "15 - Range Hood", "63 - Painting Equipment", 
+                      "107 - Exposure, source vehicle (outside structure)", 
+                      "104 - Exposure, source open fire (inc campfire, rubbish fire)", 
+                      "48 - Air Conditioner - Window or Room Unit", "96 - Chemical Reaction (eg. spontaneous combustion, etc.)", 
+                      "106 - Exposure, source grass, shrubs, trees", "35 - Fireplace - Factory Built", 
+                      "37 - Fireplace Insert", "94 - Static Electricity (spark)", "9990 - Under Investigation", 
+                      "31 - Central Heating/Cooling Unit", "56 - Halogen Lamp or light", 
+                      "91 - Fireworks", "46 - Electric Blanket, Heating Pad", 
+                      "105 - Exposure, source forest, trees, wildland", "100 - Outdoor fireplace/heater", 
+                      "103 - Exposure, source outside storage container, tank", 
+                      "38 - Chimney - Factory Built", "54 - Lamp (eg. coal, oil, naphtha, etc.)", 
+                      "102 - Exposure, source structure semi-detached or attached", "40 - Flue Pipe", 
+                      "27 - Terminations-Aluminum (incl receptables, switches, lights)", 
+                      "22 - Meter", "80 - Portable generator", "101 - Exposure, source structure detached", 
+                      "95 - Lightning", "97 - Rekindle", "45 - Washing Machine", 
+                      "25 - Circuit Wiring - Aluminum", "53 - Christmas Lights, Decorative Lighting", 
+                      "61 - Incinerator", "90 - Explosives")
+
+# Create a function to categorize sources
+categorize_ignition_sources <- function(source) {
+  # Lighting and Electrical Equipment
+  if (grepl("Lamp|Lighting|Electrical|Cable|Wire|Ballast|Transformer", source)) {
+    return("Lighting and Electrical Equipment")
+  } 
+  # Cooking and Heating Equipment
+  else if (grepl("Stove|Oven|Microwave|Grill|Fryer|Heater|Range", source)) {
+    return("Cooking and Heating Equipment")
+  } 
+  # Fire-related and Open Flame
+  else if (grepl("Fire|Matches|Candle|Flame|Torch|Burner|Smoker", source)) {
+    return("Fire-related and Open Flame")
+  } 
+  # Flammable Materials and Chemicals
+  else if (grepl("Chemical|Explosion|Flammable|Rekindle|Static|Spark", source)) {
+    return("Flammable Materials and Chemicals")
+  } 
+  # Mechanical Equipment
+  else if (grepl("Mechanical|Vehicle|Collision|Smokers|Blower", source)) {
+    return("Mechanical Equipment")
+  } 
+  # Electrical Distribution and Utility
+  else if (grepl("Distribution|Service|Utility|Panel|Circuit", source)) {
+    return("Electrical Distribution and Utility")
+  }
+
+  # Static Electricity and Sparks
+  else if (grepl("Spark|Static", source)) {
+    return("Static Electricity and Sparks")
+  } 
+  # Exposure to Fire and Environmental
+  else if (grepl("Exposure|Outside|Container|Fireplace|Campfire|Wildland", source)) {
+    return("Exposure to Fire and Environmental")
+  } 
+  # Processing Equipment (Industrial)
+  else if (grepl("Furnace|Kiln|Processing|Incinerator", source)) {
+    return("Processing Equipment")
+  } 
+  # Appliances
+  else if (grepl("Appliance|Refrigerator|Washing Machine|Dryer|Iron", source)) {
+    return("Appliances")
+  }
+
+  # Vehicle and Accidents
+  else if (grepl("Vehicle|Accident|Collision", source)) {
+    return("Vehicle and Accidents")
+  } 
+  # Multiple Ignition Sources (or suspected arson)
+  else if (grepl("Multiple|Arson", source)) {
+    return("Multiple Ignition Sources")
+  } 
+  # Cooling and Air Conditioning Equipment
+  else if (grepl("Cooling|Air Conditioner", source)) {
+    return("Cooling and Air Conditioning Equipment")
+  } 
+  # Other (for anything that doesn't fit into the above categories)
+  else {
+    return("Other")
+  }
+}
+# Apply the categorization function to each ignition source
+fire_data$ignition_source_grouped <- sapply(fire_data$ignition_source, categorize_ignition_sources)
+#Table
+#Plot
+# Plot of grouped ignition source distribution
+ggplot(fire_data, aes(x = ignition_source_grouped)) +
+  geom_bar(fill = "green") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  xlab("Ignition Source Group") +
+  ylab("Frequency") +
+  geom_text(stat = "count", aes(label = ..count..), vjust = 0.5)
+
+```
 
@@
 To explore the relationship between financial losses and casualties, incidents were grouped into financial loss and casualty categories.
-A heatmap below illustrates the interaction between these two indicators.
-
-.
+A heatmap below (@fig-financial-loss-vs-casualties-heatmap) illustrates the interaction between these two indicators.
 
 ```{r}
+#| label: fig-financial-loss-vs-casualties-heatmap
+#| fig-cap: Heatmap illustrating the relationship between financial loss and casualty groups in fire incidents.
+The plot categorizes incidents based on financial loss and total casualties, with the color intensity representing the frequency of incidents in each combination of loss and casualty group.
+This provides insights into the severity of fire incidents, highlighting areas with higher financial losses and casualties.
+We can see that there are greater anount in 1k -10k region.
+#| echo: false
+#| warning: false
+
+## Financial loss vs. Casualties ##
+fire_data$total_casualties <- fire_data$civilian_casualties + fire_data$tfs_firefighter_casualties
+
+#heatmap
 fire_data <- fire_data %>%
@@
 
-# Summarize data and plot heatmap
 severity_table <- fire_data %>%
   group_by(loss_group, casualty_group) %>%
-  summarize(count = n(), .groups = "drop")
+  summarize(count = n())
 
@@
   labs(
-    title = "Heatmap of Financial Loss and Casualty Groups",
     x = "Financial Loss Group",
@@
 
-The scatter plot below showcases the relationship between the number of responding apparatus and the logarithmic scale of financial losses.
-A Pearson correlation coefficient was calculated to assess the strength of this relationship.
+The scatter plot below showcases the relationship between the number of responding apparatuses and the logarithmic scale of financial losses (@fig-responding-apparatus-vs-log-dollar-loss).
+We can observe a slight trend between responding apparatus and financial loss A Pearson correlation coefficient was calculated to assess the strength of this relationship.
 
 ```{r}
+#| label: fig-responding-apparatus-vs-log-dollar-loss
+#| fig-cap: Scatter plot showing the relationship between the number of responding apparatus and the log-transformed financial loss in fire incidents.
+Each point represents a fire incident, with the x-axis displaying the log of the financial loss (with a 1 added for stability) and the y-axis showing the number of responding apparatus.
+There is a slight trend observed, indicating potential relationship.
+#| echo: false
+
 fire_data$Log_Dollar_Loss <- log(fire_data$estimated_dollar_loss + 1)
 
-# Scatter plot and correlation
-ggplot(fire_data, aes(x = Log_Dollar_Loss, y = number_of_responding_apparatus)) +
+# Scatter plot between number of responding apparatus and final dollar loss
+ggplot(fire_data, aes(y = number_of_responding_apparatus, x = Log_Dollar_Loss)) +
   geom_point(color = 'blue', alpha = 0.5) +
   labs(
-    title = 'Number of Responding Apparatus vs Log of Financial Loss',
-    x = 'Log of Financial Loss',
-    y = 'Number of Responding Apparatus'
-  ) +
+       x = 'Number of Responding Apparatus',
+       y = 'Log of Final Dollar Loss') +
   theme_minimal()
 
-# Calculate correlation
-cor(fire_data$Log_Dollar_Loss, fire_data$number_of_responding_apparatus, use = "complete.obs")
+# cor(fire_data$Log_Dollar_Loss, fire_data$number_of_responding_apparatus, use = "complete.obs")
+
 ```
 
-# Model
+# Model {#sec-model}
 
@@
 Quantiles (25th, 50th, 75th percentiles) were used to divide the dollar loss into categories to make severity thresholds more meaningful.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+fire_data_model <- read_csv( file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+
+```
 
@@
 
-The analysis shows that most incidents fall under the Low Severity category, which aligns with the reality of everyday fire service responses.
+The analysis shows that most incidents fall under the Low Severity category (see @tbl-severity-count), which aligns with the reality of everyday fire service responses.
 
 ```{r}
+#| label: tbl-severity-count
+#| tbl-cap: The table shows the distribution of fire incidents across Low, Medium, and High severity levels based on the number of responding apparatus, estimated dollar loss, and total casualties and incident type.
+Notice that there are musch more lower severe incidence.
+#| echo: false
+#| warning: false
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(rpart)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(fitdistrplus)
+library(knitr)
 
+#### Read data ####
+fire_data_model <- read_csv(file = here::here("data/02-analysis_data/updated_tfs_analysis_data"))
+
+#### Quantify Severity ####
+
+### Apparatus
+fire_data_model <- fire_data_model %>%
+  mutate(severity_bin = cut(number_of_responding_apparatus,
+                            breaks = c(0, 10, 20, 30, 40, 50, Inf),
+                            labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
+                            right = TRUE))
+severity_dist <- fire_data_model %>%
+  group_by(severity_bin) %>%
+  summarise(frequency = n()) %>%
+  ungroup()
+#severity_dist
+
+### Estimated_dollar_loss
+# Log transform: 
+fire_data_model$log_estimated_dollar_loss <- log(fire_data_model$estimated_dollar_loss + 1)
+# Recalculate quantiles on transformed data
+loss_quantiles <- quantile(fire_data_model$log_estimated_dollar_loss, probs = c(0.25, 0.5, 0.75))
+
+### Determining Severity
+fire_data_model <- fire_data_model %>%
+  mutate(
+    # Define Casualty Severity
+    casualty_severity = case_when(
+      total_casualties == 0 ~ "Low", 
+      total_casualties >= 4 ~ "High",
+      TRUE ~ "Medium"
+    ),
+
+    # Define Final Severity based on Apparatus, Dollar Loss, and Casualties
+    Severity = case_when(
+      # Explosion = High severity
+      final_incident_type == 
+        "02 - Explosion (including during Fire, excluding Codes 3 & 11-13)"
+      ~ "High",
+
+      # Apparatus severity classifications based on bins
+      (casualty_severity == "Low" 
+       | severity_bin %in% c("[1,10]") 
+       | log_estimated_dollar_loss <= loss_quantiles[1]) ~ "Low",
+
+      (casualty_severity == "Medium" 
+       | severity_bin %in% c("[11,20]", "[21,30]") 
+       | log_estimated_dollar_loss <= loss_quantiles[2]) ~ "Medium",
+
+      (casualty_severity == "High" 
+       | severity_bin %in% c("[31,40]", "[41,50]", ">50") 
+       | log_estimated_dollar_loss > loss_quantiles[2]) ~ "High",
+
+      TRUE ~ "Low" 
+    )
+  )
+severity_count_table <- fire_data_model %>%
+  count(Severity)
+
+# Display the table
+kable(severity_count_table)
 ```
 
-### Set Up Model For Predicting Severities
+## Set Up Model For Predicting Severities
 
@@
 
-To evaluate model performance, the dataset was split into a training set (80%) and a testing set (20%) using the `createDataPartition()` function from the `caret` package in R.
+To evaluate model performance, the dataset was split into a training set (80%) and a testing set (20%) using the `createDataPartition()` function from the `caret` package in R [@citeR].
 
-### Model Explnation
+## Model Explnation
 
@@
 
-Where: $P$ is the probability of the event ($P(\text{Severity_high}=1)$).
-$\text{logit}(P)$ is the log-odds of the probability.
-$\beta_0$ is the intercept.
+Where: $P$ is the probability of the event: $P(\text{Severity\_high}=1)$.\
+$\text{logit}(P)$ is the log-odds of the probability.\
+$\beta_0$ is the intercept.\
 $\beta_1, \beta_2, \dots, \beta_k$ are the coefficients of the predictor variables $X_1, X_2, \dots, X_k$.
 
-### Model Justification
+## Model Justification
 
@@
 Since a high p-value suggests that the variable does not significantly contribute to the model in prediction, it was removed to simplify the model and improve interpretability.
-This decision follows the principle outlined by Occam's Razor \[\@blumer1987occam\], which proposes that people prefer simplicity.
+This decision follows the principle outlined by Occam's Razor [@blumer1987occam], which proposes that people prefer simplicity.
 
 To check for multicollinearity among the predictors, the Variance Inflation Factor (VIF) was calculated.
-According to standard practice, if any predictors had a high VIF (greater than 5), they were considered for removal to avoid multicollinearity, which could distort the model's estimates.
-
-Based on the results, the variable *ignition_source_grouped* was removed from the model due to its non-significance (high p-value).
-The updated model retained only *area_of_origin_grouped*, which showed a significant relationship with the likelihood of high-severity fires.
+According to standard practice, if any predictors had a high VIF (greater than 5), they were considered for removal to avoid multicollinearity, which could distort the model's estimates [@alin2010multicollinearity].Based on the results, the variable `ignition_source_grouped` was removed from the model due to its non-significance (high p-value).
+The updated model retained only `area_of_origin_grouped`, which showed a significant relationship with the likelihood of high-severity fires.
 Model is:
@@
 The model's performance was also assessed using the Akaike Information Criterion (AIC), a measure of the model's goodness of fit.
-A lower AIC indicates a better-fitting model, and removing the insignificant variable *ignition_source_grouped* improved the AIC, providing more evidence that the updated, simpler model was a better fit for the data.
+A lower AIC indicates a better-fitting model, and removing the insignificant variable `ignition_source_grouped` improved the AIC, providing more evidence that the updated, simpler model was a better fit for the data.
 
-### **Model Evaluation**
+## Model Evaluation
 
@@
 The curve was plotted to assess the model’s ability to discriminate between high and low-severity fire incidents.
-The area under the curve (AUC) is a key metric for model performance; higher AUC values indicate better model performance \[\@metz1978basic\].
+The area under the curve (AUC) is a key metric for model performance; higher AUC values indicate better model performance [@metz1978basic].
 The simpler model, after removing insignificant predictors, combined with its performance on the test set, provides a better understanding of the relationship between fire characteristics and severity.
 
-# Results
+# Results {#sec-results}
 
-### **Logistic Regression Model Results**
+## Logistic Regression Model Results
 
 The logistic regression model was fitted to predict the probability of high severity fire incidents based on the categorical variables **area of origin**.
-The final model retained only significant predictors based on p-values and collinearity diagnostics.
-Below are the key findings:
+The final model keep only significant predictors based on p-values and col-linearity diagnostics.
+below is a summary ( see @fig-logisticresults-summary) of the model's estimates.
+The coefficients of each predictor.
+And Also included other information about model evaluation.
+It's noticeable that RMSE is not ideal, but this is already an improvement from the initial model.
 
@@
 #| message: false
+#| label: fig-logisticresults-summary
+#| fig-cap: "Table showing the summary of logistic regression model estimates for predicting fire severity.
+The table includes coefficients for each significant predictor, their standard errors, z-values, and associated p-values.
+Significant predictors are highlighted, with p-values less than 0.05 indicating statistical significance.
+The results provide insights into the relationship between 'Area of Origin' and the likelihood of high-severity fire incidents."
+
+library(rstanarm)
+library(modelsummary)
+library(tidyverse)
+library(dplyr)
+library(caret)
+library(xgboost)
+library(kableExtra)
+
+fire_data_model <- read_csv(file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+
+#### Prepare Data ####
+
+fire_data_model <- fire_data_model %>%
+  mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
+
+fire_data_model <- fire_data_model[, c(
+  "area_of_origin_grouped", 
+  "ignition_source_grouped", 
+  "Severity_high"
+)]
+
+fire_data_model$ignition_source_grouped <- as.factor(fire_data_model$ignition_source_grouped)
+fire_data_model$area_of_origin_grouped <- as.factor(fire_data_model$area_of_origin_grouped)
+
+fire_data_model$ignition_source_grouped <- gsub(" ", "_", fire_data_model$ignition_source_grouped)
+fire_data_model$area_of_origin_grouped <- gsub(",", "_", fire_data_model$area_of_origin_grouped)
+fire_data_model$area_of_origin_grouped <- gsub(" ", "_", fire_data_model$area_of_origin_grouped)
+fire_data_model$area_of_origin_grouped <- gsub("__", "_", fire_data_model$area_of_origin_grouped)
+
+trainIndex <- createDataPartition(fire_data_model$Severity_high, p = 0.8, list = FALSE)
+train_data <- fire_data_model[trainIndex, ]
+test_data <- fire_data_model[-trainIndex, ]
+
+logistic_model <-
+  readRDS(file = here::here("models/logtsic_model.rds"))
+
+# Get the summary directly from the stanreg model
+logistic_summary <- summary(logistic_model)
+
+# Access the coefficients table
+coefficients_table <- logistic_summary$coefficients
+
+# Print or work with the coefficients
+kable(coefficients_table)
+```
+
+The estimated coefficients for the significant predictor (`area_of_origin`) are visualized to interpret the relative influence of different categories below (@fig-logesiticresults).
+We can tell most categories are relative to the severity of fire incidence.
+For example, Fire from mechanical, HVAC, and electrical Areas are highly contributing to the occurrence of ore sever fire incidence.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-logesiticresults
+#| fig-cap: Bar plot of the logistic regression coefficients for the 'Area of Origin' variable, showing the effect of each location within a structure where the fire started on the severity of fire incidents.
+Positive coefficients suggest a higher likelihood of a high-severity fire in that area, while negative coefficients indicate a lower likelihood For example, the Electrical area is related to more severe fire incidence.
 
@@
 
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+logistic_model <-
+  readRDS(file = here::here("models/logtsic_model.rds"))
+
+library(ggplot2)
+coef_df <- data.frame(
+  Area_of_Origin = names(coef(logistic_model))[-1],
+  Coefficients = coef(logistic_model)[-1]
+)
+
+# Bar plot of coefficients
+ggplot(coef_df, aes(x = reorder(Area_of_Origin, Coefficients), y = Coefficients)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  coord_flip() +
+  labs(
+    x = "Area of Origin",
+    y = "Coefficient Value"
+  ) +
+  theme_minimal()
+
+```
+
+The ROC curve demonstrates the trade-off between the true positive rate (sensitivity) and false positive rate (1-specificity) for the logistic regression model.
+A higher AUC value signifies better predictive accuracy [@metz1978basic].
+In this case, the curve is above the diagonal line and towards the left corners (@fig-logesiticresults-roc), indicating it's a meaningful model for prediction.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-logesiticresults-roc
+#| fig-cap: ROC curve for logistic regression model showing the trade-off between the true positive rate and false positive rate. This roc graph indicate model favor more False Negative.
+
+library(pROC)
+predicted_probs <- predict(logistic_model, test_data, type = "response")
+roc_curve <- roc(test_data$Severity_high, predicted_probs)
+plot(roc_curve)
+```
+
+Therefore, `area_of_origin` is indeed a significant predictor of fire severity, better from it's initial model performance from comparison of AIC scores.
+
+## Fire infrastructure
+
+There is no doubt that infrastructure such as fire alarm and sprinkler systems play a crucial role in controlling fire severity.
+The presence of these systems will reduces the risk and impact of fires.
+However, the key question remains: how well do these fire safety infrastructures perform in real fire incidents?
+We come across result of question by looking at the result we get from analyzing the data set.
+In this analysis, I processed the data and visualized the distribution of fire alarm system operations and sprinkler system presence across different levels of fire severity (Low, Medium, High).
+These visualizations provide valuable insights into how the presence of these systems correlates with the severity of fire incidents.
+
+## Fire Alarm System Status
+
+The results shown from the exploration of the proportion of fire alarm system statuses by severity (@fig-firealarmsystem-proportion) reveal that the fire alarm system was successfully operated during many low to medium severity fire incidents.
+However, it did not perform as reliably in high severity cases.
+It is important to note that the y-axis represents the proportion, and considering that high-severity incidents are much less frequent, their proportion might not be as large in comparison to the number of incidents.
+Nevertheless, this comparison provides useful insights, as we never want the system to fail when a real, highly severe incident occurs.
+This allows for a clearer understanding of how alarm system functionality varies.
+It is evident that our fire alarm system needs improvement in its operation during actual fire events.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-firealarmsystem-proportion
+#| fig-cap: This shows a proportional distribution of fire alarm system operation status across different fire severity levels (Low, Medium, High).
+This stacked bar plot shows the relative frequencies of various fire alarm system operations within each severity group.
+We noctice that fire alarm was not always operated in high severity incidents, but not active in low severity fire.
+
+library(tidyverse)
+library(rstanarm)
+library(rpart)
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+
+fire_data_current <- read_csv( file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+
+#### fire_alarm_system_operation ####
+
+fire_data_current <- fire_data_current %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    fire_alarm_system_operation = factor(fire_alarm_system_operation) 
+  )
+
+top_fire_alarm_system_operation <- fire_data_current %>%
+  group_by(Severity, fire_alarm_system_operation) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
+# Filter the dataset to include only the Operation of alarm for each severity
+fire_operation_data_filtered <- fire_data_current %>%
+  semi_join(top_fire_alarm_system_operation, by = c("Severity", "fire_alarm_system_operation"))
+
+ggplot(fire_operation_data_filtered, aes(x = Severity, fill = fire_alarm_system_operation)) +
+  geom_bar(position = "fill", stat = "count") + 
+  labs(
+    x = "Severity",
+    y = "Proportion",
+    fill = "fire_alarm_system_operation"
+  ) +
+  scale_y_continuous(labels = scales::percent) + 
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 0, hjust = 0.5), 
+    strip.text = element_text(size = 12)
+  )  
+```
+
+As shown in (@fig-sprinklersystem), I have plotted the presence of the sprinkler system across different levels of fire severity.
+The operational status of the sprinkler system is represented by various colors corresponding to statuses such as "Full sprinkler system present," "Partial sprinkler system present,".
+The results reveal that in High severity incidents, the "Undetermined" status is more common, indicating there may be a lack of clarity or understanding regarding the sprinkler system's functionality.
+In Low severity incidents, the sprinkler system is most frequently reported as absent ("No sprinkler system").
+In fact, this absence is the dominant status across all three levels of severity.
+This suggests that improvements in sprinkler system availability and implementation should be a focus for the future, particularly in higher severity fire scenarios, where the system's functionality could play a critical role in preventing further negative impact.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-sprinklersystem
+#| fig-cap: This joint set of three plots illustrates the distribution of sprinkler system presence across different fire severity levels (Low, Medium, High).
+Different colors represent the various operational statuses of the sprinkler system.
+The results emphasize the varying presence and functionality of sprinkler systems, showing that, on average, the sprinkler system did not operate across all levels of fire severity.
+
+#### sprinkler_system_presence ####
+
+# Convert Severity and sprinkler_system_data_filtered to factors
+fire_data_current <- fire_data_current %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    sprinkler_system_presence = factor(sprinkler_system_presence)  # Convert ignition_source_grouped to factor
+  )
+
+# Calculate the top 5 Status of Sprinkler System for each Severity group
+top_sprinkler_system_presence <- fire_data_current %>%
+  group_by(Severity, sprinkler_system_presence) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
+# Filter the dataset to include only the top Status of Sprinkler System for each severity
+sprinkler_system_data_filtered <- fire_data_current %>%
+  semi_join(top_sprinkler_system_presence, by = c("Severity", "sprinkler_system_presence"))
+
+## Three Separate Plot
+ggplot(sprinkler_system_data_filtered, aes(x = sprinkler_system_presence, fill = sprinkler_system_presence)) +
+  geom_bar() +  
+  facet_wrap(~ Severity, scales = "free_y") + 
+  labs(
+    x = "sprinkler_system_presence",
+    y = "Count",
+    fill = "sprinkler_system_presence"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
+    strip.text = element_text(size = 12),  # Adjust facet label size
+    legend.position = "bottom"  # Place the legend at the bottom
+  )
+
+```
+
+# Discussion {#sec-discussion}
+
+## Trade-off between False Positives and False Negatives {#sec-first-point}
+
+From the ROC curve, we can tell that it minimizes false negatives more in focus.
+Putting back into our context of fire severity production, It would be less harmful to tolerate some false positives ( over-predicting the severity of fire); in order not to miss any positive cases (when real severe fire incidents came).
+
+In fact, the cost of missing high-severity fires is far greater than the cost of false positives.
+While false positives may lead to overreaction, failing to detect a severe fire could escalate into a major disaster, resulting in significant financial losses and potentially causing more casualties.
+Therefore, a model that is tuned to minimize false negatives is preferable, as it prioritizes catching severe fires—even at the expense of some false positives.
+However, I will continue to fit additional models to strike a better balance between specificity and sensitivity.
+
+## Government and Agency Actions for Improving Fire Safety Infrastructure {#sec-second-point}
+
+As the results from our analysis show, there is a significant need for improvement in both the facilitation and implementation of current fire safety systems.
+The fire alarm system, for example, demonstrated a good operational response in lower and medium severity incidents but failed to function effectively during high-severity situations.
+Similarly, sprinkler systems showed varying levels of performance, with a higher proportion of "Undetermined" and "Not applicable" statuses in high-severity incidents.
+These findings raise concerns, as high-severity fires, although less frequent, have a disproportionately severe impact.
+It is crucial for relevant agencies to ensure that these systems operate reliably during fire incidents, particularly in high-severity cases.
+Priority should be given to maintaining fire alarm systems and improving the implementation of sprinkler systems.
+These measures will be essential for mitigating the risks and impacts of severe fire incidents.
+
+## TFS Decision Making {#sec-third-point}
+
+The TFS (fire services) should prioritize decision-making before heading to the fire scene based on the predictor variables identified in the study, ensuring they reach the scene as quickly as possible without missing any high-severity incidents.
+It's also crucial to ensure the right amount of resources, such as firefighters, are allocated.
+By focusing on key variables like the area of origin, TFS can allocate resources more effectively.
+
+For example, the analysis of fire data reveals a clear pattern in the area of origin across different severity levels.
+Looking at the graph (@fig-areaoforigin), we can see that mechanical, HVAC, and electrical areas make up a significant portion of high-severity incidents, which aligns with the modeling results.
+
+Thus, when TFS get reported of a fire, if there is a possibility that the fire originated from these high-risk areas, they should be prepared to deploy additional resources and firefighters to these locations.
+This proactive approach would not only help in saving civilian lives but also protect the firefighters themselves, ensuring a safer response.
+Additionally, this strategy would allow for more efficient use of social resources, minimizing the impact of the fire while ensuring a well-coordinated respose.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-areaoforigin
+#| fig-caption: This plot shows the proportional distribution of top 5 different areas of origin across various fire severity levels (Low, Medium, High).
+The stacked bar plot illustrates the relative frequency of areas of origin, with the y-axis representing proportions.
+Mechanical, HVAC, and electrical areas are more common in high-severity incidents, suggesting these are higher-risk areas in terms of fire severity.
+
+fire_data_model <- read_csv(here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+
+#Convert Severity and area_of_origin_grouped to factors
+fire_data_model <- fire_data_model %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    area_of_origin_grouped = factor(area_of_origin_grouped)
+  )
+
+# Calculate the top 5 Area of Origins for each Severity group
+top_area_of_origin <- fire_data_model %>%
+  group_by(Severity, area_of_origin_grouped) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
+# Filter the dataset to include only the top Area of Origins for each severity
+area_data_filtered <- fire_data_model %>%
+  semi_join(top_area_of_origin, by = c("Severity", "area_of_origin_grouped"))
+
+# Create a stacked bar plot with percentages
+ggplot(area_data_filtered, aes(x = Severity, fill = area_of_origin_grouped)) +
+  geom_bar(position = "fill", stat = "count") +  # 'fill' makes the bars proportional
+  labs(
+    x = "Severity",
+    y = "Proportion",
+    fill = "Area of Origin"
+  ) +
+  scale_y_continuous(labels = scales::percent) +  # Format y-axis as percentages
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 0, hjust = 0.5),  # X-axis label alignment
+    strip.text = element_text(size = 12)  # Adjust facet label size
+  )
+
+```
+
+Using other factors, such as ignition sources (see @fig-ignitionsources), which can potentially be determined before reaching the fire scene, can significantly aid in fire reporting.
+For example, if it’s reported that the fire originated from electrical units or lightning equipment, there may be a higher likelihood of the incident being severe.
+Therefore, utilizing this information to make informed decisions will enhance the effectiveness of fire response efforts.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-ignitionsources
+#| fig-cap: This plot shows the top 5 ignition sources for each fire severity level (Low, Medium, High).
+The bar chart illustrates how different ignition sources contribute to fire severity, with separate plots for each level of severity.
+For example, fires originating from electrical units or lightning equipment may have a higher chance of being severe incidents.
+By understanding these patterns, fire reporting can be more effective, allowing for better decision-making and resource allocation.
+
+#### Understand what is related to Severity ####
+
+## Factor of Ignition Source ##
+# Convert Severity and ignition_source_grouped to factors
+fire_data_model <- fire_data_model %>%
+  mutate(
+    Severity = factor(Severity, levels = c("Low", "Medium", "High")),
+    ignition_source_grouped = factor(ignition_source_grouped)  # Convert ignition_source_grouped to factor
+  )
+
+# Calculate the top 5 ignition sources for each Severity group
+top_ignition_sources <- fire_data_model %>%
+  group_by(Severity, ignition_source_grouped) %>%
+  tally() %>%
+  group_by(Severity) %>%
+  top_n(5, n) %>%
+  ungroup()
+
+# Filter the dataset to include only the top ignition sources for each severity
+fire_data_filtered <- fire_data_model %>%
+  semi_join(top_ignition_sources, by = c("Severity", "ignition_source_grouped"))
+
+## Three Separate Plot
+ggplot(fire_data_filtered, aes(x = ignition_source_grouped, fill = ignition_source_grouped)) +
+  geom_bar() +  # Automatically assigns colors based on 'ignition_source_grouped'
+  facet_wrap(~ Severity, scales = "free_y") +  # Different y-axis for each facet
+  labs(
+    x = "Ignition Source Grouped",
+    y = "Count",
+    fill = "Ignition Source"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
+    strip.text = element_text(size = 12),  # Adjust facet label size
+    legend.position = "bottom"  # Place the legend at the bottom
+  )
+
+```
+
+\newpage
+
+\appendix
+
+# Appendix A {#sec-appendixa}
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+#### Workspace setup ####
+library(dplyr)
+library(ggplot2)
+library(tidyr)
+library(lubridate)
+library(knitr)
+
+#### Read data ####
+fire_data <- read.csv(here::here("data/02-analysis_data/tfs_analysis_data"))
+#### Data Overview ####
+# Structure
+head(fire_data)
+str(fire_data)
+
+# Missing Value
+missing_data <- colSums(is.na(fire_data))
+print(missing_data)
+
+```
+
+This appendix contains all plot/summary distribution of all variables in intrested choose in analysis data set.
+
+1.  Are of Origin (see @fig-area-origin-distribution)
+2.  Civilian Casualties (see @tbl-civilian-casualties-summary)
+3.  Estimated Dollar Loss (see @tbl-estimatedloss)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-estimatedloss
+#| tbl-cap: The table summarizes the distribution of estimated dollar losses from fire incidents.
+library(knitr)
+
+## 3. Estimated Dollar Loss (Numerical) ##
+# Overall Table
+est_loss_df <- as.data.frame(table(fire_data$estimated_dollar_loss))
+colnames(est_loss_df) <- c("Estimated of Loss", "Frequency")
+kable(summary(est_loss_df))
+# Summary table
+# Plot
+# (No plot code provided in the original snippet)
+
+```
+
+4.  Final Incidence Type (see @fig-final-incident-type-distribution)
+5.  Fire Alram Conditions (see @fig-firealarmsystem)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-firealarmsystem
+#| fig-cap: This bar chart shows the distribution of fire alarm system operation types. It highlights the frequency of each type of fire alarm system operation. Operated successfully most of the time.
+
+## 5. Fireground Conditions (Categorical) ##
+# Plot
+ggplot(fire_data, aes(x = fire_alarm_system_operation)) +
+  geom_bar(fill = "blue") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  xlab("Fire Alarm System Operation Type") +
+  ylab("Frequency") +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) 
+#Table
+#table(fire_data$fire_alarm_system_operation)
+
+```
+
+6.  Ignition Source (see @fig-grouped-ignition-source-distribution)
+7.  Number of Responding Apparatus (See @fig-responding-apparatus)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-responding-apparatus
+#| fig-cap: This histogram illustrates the distribution of the number of responding apparatus for fire incidents.
+The x-axis represents the number of apparatus dispatched, while the y-axis shows the frequency of incidents.
+The date is right skewed, closer to 1.
+## 7. Number of Responding Apparatus (Numerical) ##
+ggplot(fire_data, aes(x = number_of_responding_apparatus)) +
+  geom_histogram(bins = 30, fill = "cyan", color = "black") +
+  theme_minimal() +
+  xlab("Number of Responding Apparatus") +
+  ylab("Frequency") +
+  stat_bin(aes(label = ..count..), geom = "text", vjust = -0.5, size = 3)
+
+```
+
+8.  Number of Responding Personnel (See @fig-responding-personnel-hist)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-responding-personnel-hist
+#| fig-cap: This histogram shows the distribution of the number of responding personnel for fire incidents.
+The x-axis indicates the number of personnel dispatched, while the y-axis represents the frequency of such incidents.
+Similar dsitribution as @fig-responding-apparatus.
+# 8. Number of Responding Personnel (Numerical) ##
+#Plot
+ggplot(fire_data, aes(x = number_of_responding_personnel)) +
+  geom_histogram(bins = 30, fill = "yellow", color = "black") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
+  xlab("Number of Responding Personnel") +
+  ylab("Frequency") +
+  stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 3)
+
+```
+
+9.  Possible Cause (See @tbl-possible-cause-full)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-possible-cause-full
+#| tbl-cap: Comprehensive table showing all possible fire casues.
+## 9. Possible Cause ##
+#plot all that has frequency greater than 550
+filtered_data <- fire_data %>%
+  count(possible_cause) %>%
+  filter(n > 550)
+#Full Tabble
+cause_table <- table(fire_data$possible_cause)
+cause_table_df <- as.data.frame(cause_table) %>%
+  arrange(desc(Freq))
+kable(cause_table_df, col.names = c("Possible Cause", "Frequency"))
+```
+
+11. Sprinkler System Presence (@tbl-sprinkler-system)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: tbl-sprinkler-system
+#| tbl-cap: Table showing the frequency of sprinkler system presence across all fire incidents.
+
+## 11. Sprinkler System Presence (Categorical) ##
+#Plot
+kable(table(fire_data$sprinkler_system_presence), col.names = c("sprinkler_system_presence", "Frequncy"))
+
+```
+
+12. tfs_alarm_time (Amount of Fire Incidents by Year, see @fig-fire-incidents-by-year) (Fire Incidents by Month and Year, See @fig-year-month-incident-count)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-fire-incidents-by-year
+#| fig-caption: This bar chart shows the number of fire incidents by year, from 2011 to 2023.
+## 12. tfs_alarm_time ## 
+fire_data$year <- year(fire_data$tfs_alarm_time)
+fire_data$month <- month(fire_data$tfs_alarm_time)
+fire_data$month <- as.numeric(fire_data$month)
+
+# Plot Number of Incidents by Year
+ggplot(fire_data, aes(x = year)) +
+  geom_bar(fill = "lightblue", color = "black") +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) + 
+  labs(title = "Fire Incidents by Year", x = "Year", y = "Incident Count") +
+  theme_minimal() +
+  scale_x_continuous(breaks = seq(min(fire_data$year), max(fire_data$year), by = 1))
+theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# Table
+
+# Print the table
+# View(incident_table)
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
 #| warning: false
+#| message: false
+#| label: fig-year-month-incident-count
+#| tbl-cap: Monthly Incident Count by Year (Excluding 2022 and 2023 becasue from prvious plot, we can see they don't contribute much information)
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
+# from year plot, we can see that 2022 and 2023 does not provide much information so we can droped them.
+fire_data_filtered <- fire_data %>%
+  filter(!is.na(year) & !is.na(month) & month >= 1 & month <= 12)
+incident_count_by_year_month <- fire_data_filtered %>%
+  group_by(year, month) %>%
+  summarize(count = n(), .groups = 'drop')
+incident_count_by_year_month_ex <- incident_count_by_year_month %>%
+  filter(!(year %in% c(2022, 2023)))
+incident_count_by_year_month_ex$month <- factor(incident_count_by_year_month_ex$month, levels = 1:12)
+ggplot(incident_count_by_year_month_ex, aes(x = month, y = count)) +
+  geom_bar(stat = "identity", fill = "lightcoral", color = "black") +
+  labs( x = "Month", y = "Incident Count") +
+  scale_x_discrete(
+    breaks = 1:12,  # Show all months (1 to 12)
+    labels = month.name  # Use full month names
+  ) +
+  facet_wrap(~year, scales = "fixed") +  # Fix the y-axis scale for comparison
+  theme_minimal() +
+  theme(
+    strip.text.x = element_text(size = 10), 
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.5, size = 8), 
+    axis.title.x = element_text(size = 12), 
+    axis.title.y = element_text(size = 12), 
+    strip.background = element_rect(fill = "lightblue", color = "black"), 
+    panel.grid.major.x = element_blank(),
+    panel.grid.minor.x = element_blank()
   )
+
 ```
 
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+14. Firefighter Casualties (See @fig-firefighter-casulties)
 
@@
 #| eval: true
+#| warning: false
 #| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| label: fig-firefighter-casulties
+#| fig-cap: The Pie chart showing the distribution of firefighter casualties, rnaging from 0-3. With being the maximum amount of occurence.
 
-plot(first_model, "trace")
+## 14. Firefighter Casualties (Numerical) ##
+casualties_table <- table(fire_data$tfs_firefighter_casualties)
+casualties_df <- as.data.frame(casualties_table)
+# Plot the pie chart
+ggplot(casualties_df, aes(x = "", y = Freq, fill = factor(Var1))) +
+  geom_bar(stat = "identity", width = 1, color = "black") +
+  coord_polar(theta = "y") +
+  theme_void() +
+  ggtitle("Distribution of Firefighter Casualties") +
+  scale_fill_manual(values = c("pink", "lightblue", "lightgreen", "orange")) +
+  labs(fill = "Casualties")
+#table
+#table(fire_data$tfs_firefighter_casualties)
+```
 
-plot(first_model, "rhat")
-```
+15. Response Time (See @fig-response-time-distribution)
+
+# Appendix B {#sec-appendix-B}
+
+## Method Behind Data
+
+### Potential Bias
+
+The fire incidence dataset includes several categorical variables such as Ignition Source, Possible Cause, and Initial Types.
+During data collection, these variables, which have a wide range of categories, may introduce various types of bias.
+
+1.
+**Classification Bias**: Classification bias occurs when events are mis-classified into categories that do not accurately represent the event itself [@brogden1950theory].
+This leads to incorrect or misleading conclusions.
+For example, in the dataset, there is a variable called Possible_Cause, which has over 100 categories, including categories like "Undetermined" and "Suspected Vandalism." When categorizing causes into these categories, ambiguity arises.
+There may not be enough evidence to definitively classify a cause, and this uncertainty can result in bias.
+If we were to generalize these 100 causes into fewer categories for prediction purposes, one of the generalized categories could be biased, misrepresenting the actual distribution of causes.
+This is one of the reasons why I decided to remove this variable from the dataset, besides the fact that it contains many missing values (NAs).
+2.
+**Selective Reporting Bias**: Reporting bias occurs when data does not accurately reflect reality due to incomplete reporting [@saini2014selective].
+In our case, this is particularly relevant for civilian and firefighter casualties related to different fire incidents.
+It's possible that some fire incidents are not reported, especially smaller ones that don't require significant resources.
+On the other hand, large, tragic fire incidents are often over-reported or emphasized more than necessary.
+As a result, the reported number of casualties may be skewed.
+These discrepancies in the data can affect the definition of what is considered a "severe" fire.
+If we use these skewed measurements to make predictions or draw conclusions, the resulting insights may be biased.
+
+Due to these biases, the quality of the data depends heavily on how the original data collectors set up their categorization systems.
+It's crucial to ensure that these systems are designed to capture sufficient and accurate data.
+By acknowledging the existence of classification and reporting biases, we can avoid overly relying on conclusions and predictions that may be skewed by these factors.
+Fire incidents, by nature, have risks that should be carefully accounted for in the data collection process to ensure more reliable results.
+
+### **Privacy Protect**
+
+The fire incident dataset contains real, sensitive data collected from residents in Toronto. Therefore, it is essential to understand how the data is protected to ensure confidentiality.
+
+1.
+**Excluded Data:** As mentioned on the data portal, some data have been excluded from the dataset to protect the privacy of individuals.
+This might explain why there are fewer high-severity cases in the dataset.
+For more severe incidents, it is more likely that the information could be recognized by the public, as media coverage and news publications often spread the details.
+As a result, these incidents may be easier to trace back to specific individuals or locations.
+To protect privacy, such data have been removed.
+However, this exclusion creates a natural incompleteness in the dataset, as these important and representative data cases, especially important to predict high severity, are no longer available for analysis.
+
+2.
+**Aggregated Geographical Data**: Even though the study does not directly analyze fire incidents by district, geographical information remains important.
+To further protect privacy, the data collectors have aggregated exact locations into broader regions, such as large intersections or general areas.
+While this approach preserves privacy, it may introduce challenges for spatial analysis.
+Spatial analysis heavily relies on precise location data, so the aggregation of location information can limit the ability to perform detailed spatial studies and may affect the accuracy of geographical insights.
+
+### Sampling Method in Fire Incident Data
+
+The fire incident dataset comes from the Ontario Fire Marshal (OFM) and includes incidents to which the Toronto Fire Service (TFS) responded.
+This data collection is not random, as it is not based on surveys or experimental studies.
+Instead, it is observational data collected by a specific organization.
+This can be considered **event-based sampling**.
+
+Event-based sampling is a technique where data is collected when certain events occur, meeting predefined criteria [@sanchez2009application].
+This is different from systematically collecting data over a fixed time frame.
+In this case, data are recorded for every fire incident recognized by the OFM, with each fire incident treated as an event.
+Event-related data, such as the reason for the fire, time of occurrence, and the TFS's response actions, are recorded.
+
+This is a **non-random sampling method**, meaning that events are not randomly selected from the entire population[@brislin1971non].
+Here, the data is gathered based on the occurrence of specific events.
+As a result, the data may be skewed by the types of incidents that happen to be recorded.
+For example, areas with higher risks of fires, such as electrical factories according to our modelling, are more likely to be included in the dataset due to the higher frequency of fire incidents in these locations.
+Therefore, while the dataset provides valuable insights, it is important to acknowledge that the data are not randomly selected, which could introduce bias in the conclusions or predictions drawn from the dataset.
 

### 2024-12-07 scripts/00-simulate_data.R
store data into parquet, aand change code accordingly

@@ -9,12 +9,14 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+
 set.seed(1009633096)
 
 
 #### Simulate data ####
 
-clean_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
+clean_data <- read_parquet("data/02-analysis_data/tfs_analysis_data.parquet")
 
 #number of rows
 n_simulated <- nrow(clean_data)
@@ -120,4 +122,4 @@ simulated_data <- data.frame(
 
 
 #### Save data ####
-write_csv(simulated_data, "data/00-simulated_data/simulated_data.csv")
+write_parquet(simulated_data, "data/00-simulated_data/simulated_data.parquet")

### 2024-12-07 scripts/01-test_simulated_data.R

@@ -12,9 +12,10 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
-simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-analysis_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
+simulated_data <- read_parquet("data/00-simulated_data/simulated_data.parquet")
+analysis_data <- read_parquet("data/02-analysis_data/tfs_analysis_data.parquet")
 
 # Test if the simulated data was successfully loaded
 if (exists("simulated_data")) {

### 2024-12-07 scripts/03-clean_data.R

@@ -11,6 +11,7 @@
 library(tidyverse)
 library(dplyr)
 library(lubridate)
+library(arrow)
 
 #### Load data ####
 raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
@@ -96,4 +97,4 @@ clean_data$response_time <- as.numeric(difftime(clean_data$tfs_arrival_time,
 
 
 #### Save data ####
-write_csv(clean_data, "data/02-analysis_data/tfs_analysis_data")
+write_parquet(clean_data, "data/02-analysis_data/tfs_analysis_data.parquet")

### 2024-12-07 scripts/04-test_analysis_data.R

@@ -11,8 +11,9 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
-analysis_data <- read_csv("data/02-analysis_data/tfs_analysis_data")
+analysis_data <- read_parquet("data/02-analysis_data/tfs_analysis_data.parquet")
 raw_data <- read_csv("data/01-raw_data/raw_fire_data.csv")
 
 

### 2024-12-07 scripts/05-exploratory_data_analysis.R

@@ -14,10 +14,11 @@ library(ggplot2)
 library(tidyr)
 library(lubridate)
 library(knitr)
+library(arrow)
 
 
 #### Read data ####
-fire_data <- read.csv("data/02-analysis_data/tfs_analysis_data")
+fire_data <- read_parquet("data/02-analysis_data/tfs_analysis_data.parquet")
 
 #### Data Overview ####
 # Structure
@@ -179,7 +180,7 @@ civilian_casualties_df
 # Summary table
 summary(fire_data$civilian_casualties)
 #Plot
-ggplot(fire_data, aes(x = X_id, y = civilian_casualties)) +
+ggplot(fire_data, aes(x = `_id`, y = civilian_casualties)) +
   geom_point(color = "darkgreen", size = 2, shape = 16) +
   theme_minimal() +
   ggtitle("Scatterplot of Number of Responding Personnel vs. Incident ID") +
@@ -357,7 +358,7 @@ ggplot(fire_data, aes(x = number_of_responding_apparatus)) +
   ylab("Frequency") +
   stat_bin(aes(label = ..count..), geom = "text", vjust = -0.5, size = 3)
 #ScatterPlot
-ggplot(fire_data, aes(x = X_id, y = number_of_responding_apparatus)) +
+ggplot(fire_data, aes(x = `_id`, y = number_of_responding_apparatus)) +
   geom_point(color = "darkorange", size = 2, shape = 16) +
   theme_minimal() +
   ggtitle("Scatterplot of Number of Responding apparatus vs. Incident ID") +
@@ -378,7 +379,7 @@ ggplot(fire_data, aes(x = number_of_responding_personnel)) +
   ylab("Frequency") +
   stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 3)
 #ScatterPlot
-ggplot(fire_data, aes(x = X_id, y = number_of_responding_personnel)) +
+ggplot(fire_data, aes(x = `_id`, y = number_of_responding_personnel)) +
   geom_point(color = "orange", size = 2, shape = 16) +
   theme_minimal() +
   ggtitle("Scatterplot of Number of Responding Personnel vs. Incident ID") +
@@ -511,7 +512,7 @@ summary(fire_data$response_time)
 
 #### Save new analysis data ####
 fire_data$total_casualties <- fire_data$civilian_casualties + fire_data$tfs_firefighter_casualties
[…]

### 2024-12-07 scripts/06-define_severity.R

@@ -16,9 +16,10 @@ library(ggplot2)
 library(dplyr)
 library(tidyr)
 library(fitdistrplus)
+library(arrow)
 
 #### Read data ####
-fire_data_model <- read_csv("data/02-analysis_data/updated_tfs_analysis_data")
+fire_data_model <- read_parquet("data/02-analysis_data/updated_tfs_analysis_data.parquet")
 
 #### Quantify Severity ####
 
@@ -80,9 +81,9 @@ fire_data_model <- fire_data_model %>%
 
 
 #### Save Data ####
-write_csv(fire_data_model, "data/02-analysis_data/severity_tfs_analysis_data")
+write_parquet(fire_data_model, "data/02-analysis_data/severity_tfs_analysis_data.parquet")
 
-fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+fire_data_model <- read_parquet("data/02-analysis_data/severity_tfs_analysis_data.parquet")
 
 
 #### Understand what is related to Severity ####

### 2024-12-07 scripts/07-model_severity.R

@@ -13,9 +13,10 @@ library(tidyverse)
 library(dplyr)
 library(caret)
 library(xgboost)
+library(arrow)
 
 #### Read data ####
-fire_data_model <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+fire_data_model <- read_parquet("data/02-analysis_data/severity_tfs_analysis_data.parquet")
 
 #### Prepare Data ####
 

### 2024-12-07 scripts/08-current-system.R

@@ -16,9 +16,10 @@ library(ggplot2)
 library(dplyr)
 library(tidyr)
 library(fitdistrplus)
+library(arrow)
 
 #### Read data ####
-fire_data_current <- read_csv("data/02-analysis_data/severity_tfs_analysis_data")
+fire_data_current <- read_parquet("data/02-analysis_data/severity_tfs_analysis_data.parquet")
 
 #### fire_alarm_system_operation ####
 

### 2024-12-07 paper/paper.qmd
update parquet data in use and move bias/privacy of data into dicussion section

@@
 library(knitr)
-
-fire_data <- read.csv(file = here::here("data/02-analysis_data/tfs_analysis_data"))
+library(arrow)
+
+fire_data <- read_parquet(file = here::here("data/02-analysis_data/tfs_analysis_data.parquet"))
 
@@
 ## Data Cleaning
-
-After loading the raw data, I selected the columns of interest as explained in the previous section.
+The raw dataset was stored in its original csv format to ensure data integrity and reproducibility, and later analysis data after cleaning and processing is saved into parquet file using `write_parquet` function in `arrow` package [@arrow2024].
+
+After loading the original raw data, I selected the columns of interest as explained in the previous section.
 These columns were then renamed to lowercase, and spaces were replaced with underscores to ensure consistency.
@@
 library(ggplot2)
-fire_data <- read.csv(here::here("data/02-analysis_data/tfs_analysis_data"))
+library(arrow)
+fire_data <- read_parquet(here::here("data/02-analysis_data/tfs_analysis_data.parquet"))
 ## 15. Response Time (Numerical) ##
@@
   ylab("Frequency") +
-  stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 3) +
+  stat_bin(aes(label = ..count..), geom = "text", vjust = -0.5, size = 3) +
   xlim(0, 15)
@@
   ylab("Frequency") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = 0.5)
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5)
 
@@
 
-fire_data_model <- read_csv( file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+fire_data_model <- read_parquet( file = here::here("data/02-analysis_data/severity_tfs_analysis_data.parquet"))
 
@@
 #### Read data ####
-fire_data_model <- read_csv(file = here::here("data/02-analysis_data/updated_tfs_analysis_data"))
+fire_data_model <- read_parquet(file = here::here("data/02-analysis_data/updated_tfs_analysis_data.parquet"))
 
@@
 
-fire_data_model <- read_csv(file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+fire_data_model <- read_parquet(file = here::here("data/02-analysis_data/severity_tfs_analysis_data.parquet"))
 
@@
 
-## Fire Alarm System Status
+### Fire Alarm System Status
 
@@
 
-fire_data_current <- read_csv( file = here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+fire_data_current <- read_parquet( file = here::here("data/02-analysis_data/severity_tfs_analysis_data.parquet"))
 
@@
 ```
+
+### Sprinkler System Status
 
@@
 
-## Government and Agency Actions for Improving Fire Safety Infrastructure {#sec-second-point}
+## Limitations and Biases in the Data {#sec-second-point}
+### Potential Bias
+
+The fire incidence dataset includes several categorical variables such as Ignition Source, Possible Cause, and Initial Types.
+During data collection, these variables, which have a limited range of categories, may introduce various types of bias.
+
+1.
+**Classification Bias**: Classification bias occurs when events are mis-classified into categories that do not accurately represent the event itself [@brogden1950theory].
+This leads to incorrect or misleading conclusions.
+For example, in the dataset, there is a variable called Possible_Cause, which has over 100 categories, including categories like "Undetermined" and "Suspected Vandalism." When categorizing causes into these categories, ambiguity arises.
+There may not be enough evidence to definitively classify a cause, and this uncertainty can result in bias.
+If we were to generalize these 100 causes into fewer categories for prediction purposes, one of the generalized categories could be biased, misrepresenting the actual distribution of causes.
+This is one of the reasons why I decided to remove this variable from the dataset, besides the fact that it contains many missing values (NAs).
+2.
+**Selective Reporting Bias**: Reporting bias occurs when data does not accurately reflect reality due to incomplete reporting [@saini2014selective].
+In our case, this is particularly relevant for civilian and firefighter casualties related to different fire incidents.
+It's possible that some fire incidents are not reported, especially smaller ones that don't require significant resources.
+On the other hand, large, tragic fire incidents are often over-reported or emphasized more than necessary.
+As a result, the reported number of casualties may be skewed.
+These discrepancies in the data can affect the definition of what is considered a "severe" fire.
+If we use these skewed measurements to make predictions or draw conclusions, the resulting insights may be biased.
+
+Due to these biases, the quality of the data depends heavily on how the original data collectors set up their categorization systems.
+It's crucial to ensure that these systems are designed to capture sufficient and accurate data.
+By acknowledging the existence of classification and reporting biases, we can avoid overly relying on conclusions and predictions that may be skewed by these factors.
+Fire incidents, by nature, have risks that should be carefully accounted for in the data collection process to ensure more reliable results.
+
+### Sampling Method in the Data
+
+The fire incident dataset comes from the Ontario Fire Marshal (OFM) and includes incidents to which the Toronto Fire Service (TFS) responded.
+This data collection is not random, as it is not based on surveys or experimental studies.
+Instead, it is observational data collected by a specific organization.
+This can be considered **event-based sampling**.
+
+Event-based sampling is a technique where data is collected when certain events occur, meeting predefined criteria [@sanchez2009application].
+This is different from systematically collecting data over a fixed time frame.
+In this case, data are recorded for every fire incident recognized by the OFM, with each fire incident treated as an event.
+Event-related data, such as the reason for the fire, time of occurrence, and the TFS's response actions, are recorded.
+
+This is a **non-random sampling method**, meaning that events are not randomly selected from the entire population[@brislin1971non].
+Here, the data is gathered based on the occurrence of specific events.
+As a result, the data may be skewed by the types of incidents that happen to be recorded.
+For example, areas with higher risks of fires, such as electrical factories according to our modelling, are more likely to be included in the dataset due to the higher frequency of fire incidents in these locations.
+Therefore, while the dataset provides valuable insights, it is important to acknowledge that the data are not randomly selected, which could introduce bias in the conclusions or predictions drawn from the dataset.
+
+### Privacy Protection in the Data
+
+The fire incident dataset contains real, sensitive data collected from residents in Toronto. Therefore, it is essential to understand how the data is protected to ensure confidentiality.
+
+1.
+**Excluded Data:** As mentioned on the data portal, some data have been excluded from the dataset to protect the privacy of individuals.
+This might explain why there are fewer high-severity cases in the dataset.
+For more severe incidents, it is more likely that the information could be recognized by the public, as media coverage and news publications often spread the details.
+As a result, these incidents may be easier to trace back to specific individuals or locations.
+To protect privacy, such data have been removed.
+However, this exclusion creates a natural incompleteness in the dataset, as these important and representative data cases, especially important to predict high severity, are no longer available for analysis.
+
+2.
+**Aggregated Geographical Data**: Even though the study does not directly analyze fire incidents by district, geographical information remains important.
+To further protect privacy, the data collectors have aggregated exact locations into broader regions, such as large intersections or general areas.
+While this approach preserves privacy, it may introduce challenges for spatial analysis.
+Spatial analysis heavily relies on precise location data, so the aggregation of location information can limit the ability to perform detailed spatial studies and may affect the accuracy of geographical insights.
+
+## Government and Agency Actions for Improving Fire Safety Infrastructure {#sec-third-point}
 
@@
 
-## TFS Decision Making {#sec-third-point}
+## TFS Decision Making {#sec-fourth-point}
 
@@
 
-fire_data_model <- read_csv(here::here("data/02-analysis_data/severity_tfs_analysis_data"))
+fire_data_model <- read_parquet(here::here("data/02-analysis_data/severity_tfs_analysis_data.parquet"))
 
@@
 
-# Appendix A {#sec-appendixa}
+# Appendix {-}
+
+# Data and Variable Selection
 
@@
 #### Read data ####
-fire_data <- read.csv(here::here("data/02-analysis_data/tfs_analysis_data"))
+fire_data <- read_parquet(here::here("data/02-analysis_data/tfs_analysis_data.parquet"))
 #### Data Overview ####
@@
 
-This appendix contains all plot/summary distribution of all variables in intrested choose in analysis data set.
+This appendix contains all plot/summary distribution of all variables in interested choose in analysis data set.
 
@@
 
-# Appendix B {#sec-appendix-B}
-
-## Method Behind Data
-
-### Potential Bias
-
-The fire incidence dataset includes several categorical variables such as Ignition Source, Possible Cause, and Initial Types.
-During data collection, these variables, which have a wide range of categories, may introduce various types of bias.
-
-1.
-**Classification Bias**: Classification bias occurs when events are mis-classified into categories that do not accurately represent the event itself [@brogden1950theory].
-This leads to incorrect or misleading conclusions.
-For example, in the dataset, there is a variable called Possible_Cause, which has over 100 categories, including categories like "Undetermined" and "Suspected Vandalism." When categorizing causes into these categories, ambiguity arises.
-There may not be enough evidence to definitively classify a cause, and this uncertainty can result in bias.
-If we were to generalize these 100 causes into fewer categories for prediction purposes, one of the generalized categories could be biased, misrepresenting the actual distribution of causes.
-This is one of the reasons why I decided to remove this variable from the dataset, besides the fact that it contains many missing values (NAs).
-2.
-**Selective Reporting Bias**: Reporting bias occurs when data does not accurately reflect reality due to incomplete reporting [@saini2014selective].
-In our case, this is particularly relevant for civilian and firefighter casualties related to different fire incidents.
-It's possible that some fire incidents are not reported, especially smaller ones that don't require significant resources.
-On the other hand, large, tragic fire incidents are often over-reported or emphasized more than necessary.
-As a result, the reported number of casualties may be skewed.
-These discrepancies in the data can affect the definition of what is considered a "severe" fire.
-If we use these skewed measurements to make predictions or draw conclusions, the resulting insights may be biased.
-
-Due to these biases, the quality of the data depends heavily on how the original data collectors set up their categorization systems.
-It's crucial to ensure that these systems are designed to capture sufficient and accurate data.
-By acknowledging the existence of classification and reporting biases, we can avoid overly relying on conclusions and predictions that may be skewed by these factors.
-Fire incidents, by nature, have risks that should be carefully accounted for in the data collection process to ensure more reliable results.
-
-### **Privacy Protect**
-
-The fire incident dataset contains real, sensitive data collected from residents in Toronto. Therefore, it is essential to understand how the data is protected to ensure confidentiality.
-
-1.
-**Excluded Data:** As mentioned on the data portal, some data have been excluded from the dataset to protect the privacy of individuals.
-This might explain why there are fewer high-severity cases in the dataset.
-For more severe incidents, it is more likely that the information could be recognized by the public, as media coverage and news publications often spread the details.
-As a result, these incidents may be easier to trace back to specific individuals or locations.
-To protect privacy, such data have been removed.
-However, this exclusion creates a natural incompleteness in the dataset, as these important and representative data cases, especially important to predict high severity, are no longer available for analysis.
-
-2.
-**Aggregated Geographical Data**: Even though the study does not directly analyze fire incidents by district, geographical information remains important.
-To further protect privacy, the data collectors have aggregated exact locations into broader regions, such as large intersections or general areas.
-While this approach preserves privacy, it may introduce challenges for spatial analysis.
-Spatial analysis heavily relies on precise location data, so the aggregation of location information can limit the ability to perform detailed spatial studies and may affect the accuracy of geographical insights.
-
-### Sampling Method in Fire Incident Data
-
-The fire incident dataset comes from the Ontario Fire Marshal (OFM) and includes incidents to which the Toronto Fire Service (TFS) responded.
-This data collection is not random, as it is not based on surveys or experimental studies.
-Instead, it is observational data collected by a specific organization.
-This can be considered **event-based sampling**.
-
-Event-based sampling is a technique where data is collected when certain events occur, meeting predefined criteria [@sanchez2009application].
-This is different from systematically collecting data over a fixed time frame.
-In this case, data are recorded for every fire incident recognized by the OFM, with each fire incident treated as an event.
-Event-related data, such as the reason for the fire, time of occurrence, and the TFS's response actions, are recorded.
-
-This is a **non-random sampling method**, meaning that events are not randomly selected from the entire population[@brislin1971non].
-Here, the data is gathered based on the occurrence of specific events.
-As a result, the data may be skewed by the types of incidents that happen to be recorded.
-For example, areas with higher risks of fires, such as electrical factories according to our modelling, are more likely to be included in the dataset due to the higher frequency of fire incidents in these locations.
-Therefore, while the dataset provides valuable insights, it is important to acknowledge that the data are not randomly selected, which could introduce bias in the conclusions or predictions drawn from the dataset.
-
 \newpage

### 2024-12-07 paper/paper.qmd
%>% changed to |>

@@
 [@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements.
-Finally, [@sec-appendixa] includes supplementary information and additional data details.
-[@sec-appendix-B] provided insight into Fire Incidence, about their sampling method and potential bias.
+Finally, [@sec-appendix] includes supplementary information and additional data details.
 
@@
 However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
-A more detailed explanation will be provided in the @sec-appendixa.
+A more detailed explanation will be provided in the @sec-appendix.
 
@@
 The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
-Can check @sec-appendixa for visualization/summary of the distribution of all variables.
+Can check @sec-appendix for visualization/summary of the distribution of all variables.
 Below are a few important variables considered:
@@
 We can see that most of the time, there are no casualties.
-This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
+This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendix), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
 
@@
 #heatmap
-fire_data <- fire_data %>%
+fire_data <- fire_data |>
   mutate(
@@
 
-severity_table <- fire_data %>%
-  group_by(loss_group, casualty_group) %>%
+severity_table <- fire_data |>
+  group_by(loss_group, casualty_group) |>
   summarize(count = n())
@@
 ### Apparatus
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(severity_bin = cut(number_of_responding_apparatus,
@@
                             right = TRUE))
-severity_dist <- fire_data_model %>%
-  group_by(severity_bin) %>%
-  summarise(frequency = n()) %>%
+severity_dist <- fire_data_model |>
+  group_by(severity_bin) |>
+  summarise(frequency = n()) |>
   ungroup()
@@
 ### Determining Severity
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(
@@
   )
-severity_count_table <- fire_data_model %>%
+severity_count_table <- fire_data_model |>
   count(Severity)
@@
 
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
@@
 
-fire_data_current <- fire_data_current %>%
+fire_data_current <- fire_data_current |>
   mutate(
@@
 
-top_fire_alarm_system_operation <- fire_data_current %>%
-  group_by(Severity, fire_alarm_system_operation) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_fire_alarm_system_operation <- fire_data_current |>
+  group_by(Severity, fire_alarm_system_operation) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
@@
 # Filter the dataset to include only the Operation of alarm for each severity
-fire_operation_data_filtered <- fire_data_current %>%
+fire_operation_data_filtered <- fire_data_current |>
   semi_join(top_fire_alarm_system_operation, by = c("Severity", "fire_alarm_system_operation"))
@@
 # Convert Severity and sprinkler_system_data_filtered to factors
-fire_data_current <- fire_data_current %>%
+fire_data_current <- fire_data_current |>
   mutate(
@@
 # Calculate the top 5 Status of Sprinkler System for each Severity group
-top_sprinkler_system_presence <- fire_data_current %>%
-  group_by(Severity, sprinkler_system_presence) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_sprinkler_system_presence <- fire_data_current |>
+  group_by(Severity, sprinkler_system_presence) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
@@
 # Filter the dataset to include only the top Status of Sprinkler System for each severity
-sprinkler_system_data_filtered <- fire_data_current %>%
+sprinkler_system_data_filtered <- fire_data_current |>
   semi_join(top_sprinkler_system_presence, by = c("Severity", "sprinkler_system_presence"))
@@
 #Convert Severity and area_of_origin_grouped to factors
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(
@@
 # Calculate the top 5 Area of Origins for each Severity group
-top_area_of_origin <- fire_data_model %>%
-  group_by(Severity, area_of_origin_grouped) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_area_of_origin <- fire_data_model |>
+  group_by(Severity, area_of_origin_grouped) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
@@
 # Filter the dataset to include only the top Area of Origins for each severity
-area_data_filtered <- fire_data_model %>%
+area_data_filtered <- fire_data_model |>
   semi_join(top_area_of_origin, by = c("Severity", "area_of_origin_grouped"))
@@
 # Convert Severity and ignition_source_grouped to factors
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(
@@
 # Calculate the top 5 ignition sources for each Severity group
-top_ignition_sources <- fire_data_model %>%
-  group_by(Severity, ignition_source_grouped) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_ignition_sources <- fire_data_model |>
+  group_by(Severity, ignition_source_grouped) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
@@
 # Filter the dataset to include only the top ignition sources for each severity
-fire_data_filtered <- fire_data_model %>%
+fire_data_filtered <- fire_data_model |>
   semi_join(top_ignition_sources, by = c("Severity", "ignition_source_grouped"))
@@
 #plot all that has frequency greater than 550
-filtered_data <- fire_data %>%
-  count(possible_cause) %>%
+filtered_data <- fire_data |>
+  count(possible_cause) |>
   filter(n > 550)
@@
 cause_table <- table(fire_data$possible_cause)
-cause_table_df <- as.data.frame(cause_table) %>%
+cause_table_df <- as.data.frame(cause_table) |>
   arrange(desc(Freq))
@@
 # from year plot, we can see that 2022 and 2023 does not provide much information so we can droped them.
-fire_data_filtered <- fire_data %>%
+fire_data_filtered <- fire_data |>
   filter(!is.na(year) & !is.na(month) & month >= 1 & month <= 12)
-incident_count_by_year_month <- fire_data_filtered %>%
-  group_by(year, month) %>%
+incident_count_by_year_month <- fire_data_filtered |>
+  group_by(year, month) |>
   summarize(count = n(), .groups = 'drop')
-incident_count_by_year_month_ex <- incident_count_by_year_month %>%
+incident_count_by_year_month_ex <- incident_count_by_year_month |>
   filter(!(year %in% c(2022, 2023)))

### 2024-12-07 scripts/03-clean_data.R

@@ -35,7 +35,7 @@ head(clean_data, 6)
 summary(clean_data)   
 
 # Since I want to study possible sources leading to different severity of fire, no info about sources should be dropped
-clean_data <- clean_data %>%
+clean_data <- clean_data |>
   filter(ignition_source != "999 - Undetermined")
 
 ### Ensure Datetime Consistency ###

### 2024-12-07 scripts/05-exploratory_data_analysis.R

@@ -391,8 +391,8 @@ summary(fire_data$number_of_responding_personnel)
 
 ## 9. Possible Cause ##
 #plot all that hsa frequency greater than 550
-filtered_data <- fire_data %>%
-  count(possible_cause) %>%
+filtered_data <- fire_data |>
+  count(possible_cause) |>
   filter(n > 550)
 ggplot(filtered_data, aes(x = possible_cause, y = n)) +
   geom_bar(stat = "identity", fill = "blue") +
@@ -404,7 +404,7 @@ ggplot(filtered_data, aes(x = possible_cause, y = n)) +
   geom_text(aes(label = n), vjust = -0.5)
 #Full Tabble
 cause_table <- table(fire_data$possible_cause)
-cause_table_df <- as.data.frame(cause_table) %>%
+cause_table_df <- as.data.frame(cause_table) |>
   arrange(desc(Freq))
 kable(cause_table_df, col.names = c("Possible Cause", "Frequency"), 
              caption = "Frequency of Possible Causes (Descending Order)", 
@@ -439,12 +439,12 @@ ggplot(fire_data, aes(x = year)) +
 theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
 # from year plot, we can see that 2022 and 2023 does not provide much information so we can droped them.
-fire_data_filtered <- fire_data %>%
+fire_data_filtered <- fire_data |>
   filter(!is.na(year) & !is.na(month) & month >= 1 & month <= 12)
-incident_count_by_year_month <- fire_data_filtered %>%
-  group_by(year, month) %>%
+incident_count_by_year_month <- fire_data_filtered |>
+  group_by(year, month) |>
   summarize(count = n(), .groups = 'drop')
-incident_count_by_year_month_ex <- incident_count_by_year_month %>%
+incident_count_by_year_month_ex <- incident_count_by_year_month |>
   filter(!(year %in% c(2022, 2023)))
 incident_count_by_year_month_ex$month <- factor(incident_count_by_year_month_ex$month, levels = 1:12)
 ggplot(incident_count_by_year_month_ex, aes(x = month, y = count)) +
@@ -467,7 +467,7 @@ ggplot(incident_count_by_year_month_ex, aes(x = month, y = count)) +
   )
 
 # Table
-incident_table <- incident_count_by_year_month %>%
+incident_table <- incident_count_by_year_month |>
[…]

### 2024-12-07 scripts/06-define_severity.R

@@ -24,14 +24,14 @@ fire_data_model <- read_parquet("data/02-analysis_data/updated_tfs_analysis_data
 #### Quantify Severity ####
 
 ### Apparatus
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(severity_bin = cut(number_of_responding_apparatus,
                             breaks = c(0, 10, 20, 30, 40, 50, Inf),
                             labels = c("[1,10]", "[11,20]", "[21,30]", "[31,40]", "[41,50]", ">50"),
                             right = TRUE))
-severity_dist <- fire_data_model %>%
-  group_by(severity_bin) %>%
-  summarise(frequency = n()) %>%
+severity_dist <- fire_data_model |>
+  group_by(severity_bin) |>
+  summarise(frequency = n()) |>
   ungroup()
 #severity_dist
 
@@ -46,7 +46,7 @@ loss_quantiles <- quantile(fire_data_model$log_estimated_dollar_loss, probs = c(
 table(fire_data_model$total_casualties)
 
 ### Determining Severity
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(
     # Define Casualty Severity
     casualty_severity = case_when(
@@ -90,22 +90,22 @@ fire_data_model <- read_parquet("data/02-analysis_data/severity_tfs_analysis_dat
 
 ## Factor of Ignition Source ##
 # Convert Severity and ignition_source_grouped to factors
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(
     Severity = factor(Severity, levels = c("Low", "Medium", "High")),
     ignition_source_grouped = factor(ignition_source_grouped)  # Convert ignition_source_grouped to factor
   )
 
 # Calculate the top 5 ignition sources for each Severity group
-top_ignition_sources <- fire_data_model %>%
-  group_by(Severity, ignition_source_grouped) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_ignition_sources <- fire_data_model |>
+  group_by(Severity, ignition_source_grouped) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
 
 # Filter the dataset to include only the top ignition sources for each severity
-fire_data_filtered <- fire_data_model %>%
+fire_data_filtered <- fire_data_model |>
[…]

### 2024-12-07 scripts/07-model_severity.R

@@ -20,7 +20,7 @@ fire_data_model <- read_parquet("data/02-analysis_data/severity_tfs_analysis_dat
 
 #### Prepare Data ####
 
-fire_data_model <- fire_data_model %>%
+fire_data_model <- fire_data_model |>
   mutate(Severity_high = as.factor(ifelse(Severity == "High", 1, 0)))
 
 fire_data_model <- fire_data_model[, c(

### 2024-12-07 scripts/08-current-system.R

@@ -24,22 +24,22 @@ fire_data_current <- read_parquet("data/02-analysis_data/severity_tfs_analysis_d
 #### fire_alarm_system_operation ####
 
 # Convert Severity and alarm_fire_data_filtered to factors
-fire_data_current <- fire_data_current %>%
+fire_data_current <- fire_data_current |>
   mutate(
     Severity = factor(Severity, levels = c("Low", "Medium", "High")),
     fire_alarm_system_operation = factor(fire_alarm_system_operation) 
   )
 
 # Calculate the top 5 Status of fire Alarm for each Severity group
-top_fire_alarm_system_operation <- fire_data_current %>%
-  group_by(Severity, fire_alarm_system_operation) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_fire_alarm_system_operation <- fire_data_current |>
+  group_by(Severity, fire_alarm_system_operation) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
 
 # Filter the dataset to include only the top status of fire alarm for each severity
-alarm_fire_data_filtered <- fire_data_current %>%
+alarm_fire_data_filtered <- fire_data_current |>
   semi_join(top_fire_alarm_system_operation, by = c("Severity", "fire_alarm_system_operation"))
 
 # Create a stacked bar plot with percentages
@@ -81,22 +81,22 @@ ggplot(alarm_fire_data_filtered, aes(x = fire_alarm_system_operation, fill = fir
 #### sprinkler_system_presence ####
 
 # Convert Severity and sprinkler_system_data_filtered to factors
-fire_data_current <- fire_data_current %>%
+fire_data_current <- fire_data_current |>
   mutate(
     Severity = factor(Severity, levels = c("Low", "Medium", "High")),
     sprinkler_system_presence = factor(sprinkler_system_presence)  # Convert ignition_source_grouped to factor
   )
 
 # Calculate the top 5 Status of Sprinkler System for each Severity group
-top_sprinkler_system_presence <- fire_data_current %>%
-  group_by(Severity, sprinkler_system_presence) %>%
-  tally() %>%
-  group_by(Severity) %>%
-  top_n(5, n) %>%
+top_sprinkler_system_presence <- fire_data_current |>
+  group_by(Severity, sprinkler_system_presence) |>
+  tally() |>
+  group_by(Severity) |>
+  top_n(5, n) |>
   ungroup()
 
[…]

### 2024-12-08 paper/paper.qmd
Add information on varibles sellection and visualize selected variables

@@
 library(lubridate)
+library(kableExtra)
 
@@
 ### Preview data ###
-kable(head(clean_data, 6))
+head_table <- head(clean_data)
+kable(head(head_table)) |>
+  kable_styling(latex_options = c("scale_down"))
+
 ```
@@
 ## Data Cleaning
+
 The raw dataset was stored in its original csv format to ensure data integrity and reproducibility, and later analysis data after cleaning and processing is saved into parquet file using `write_parquet` function in `arrow` package [@arrow2024].
@@
 After calculating, I decided to remove columns with more than 20% missing data, carefully considering their relevance to the study.
-
-Variables removed:
-
-1.
-**Building_Status**: This column provides information on the condition of the building, but it is not directly related to the core objectives of the study, which focus on fire severity, casualties, and financial losses.
-With more than 20% missing data, it was deemed unnecessary.
-
-2.
-**Business_Impact**: While this column tracks the impact of the fire on businesses, it overlaps with the **Estimated_Dollar_Loss** column, which provides more comprehensive financial data.
-Due to the high missing rate, this column was removed.
-
-3.
-**Exposures**: This column captures the number of exposure fires but does not directly contribute to understanding fire severity or its impact on casualties and losses.
-With more than 20% missing data, it was excluded from the analysis.
-
-Kept variables: The remaining variables are essential for the completeness of the analysis. For categorical data entries, missing values are replaced with “unreported.”
-
--   **Fire_Alarm_System_Operation**: This variable is crucial because the study aims to understand fire infrastructure.
-
--   **Sprinkler_System_Presence**: This is also an important variable for understanding fire safety. Given that there are fewer variables under "Safety Infrastructure," it is essential to retain them.
 
@@
 ## Limitations and Biases in the Data {#sec-second-point}
+
 ### Potential Bias
@@
 
-#### Understand what is related to Severity ####
-
 ## Factor of Ignition Source ##
@@
 
-# Appendix {-}
-
-# Data and Variable Selection
-
-```{r}
-#| include: false
-#| warning: false
-#| message: false
+# Appendix {.unnumbered}
+
+# Variable Selection and Visualization
+
+```{r}
+#| inlcude: false
+#| warning: false
+#| message: false
+#| echo: false
 #### Workspace setup ####
@@
 
-#### Read data ####
-fire_data <- read_parquet(here::here("data/02-analysis_data/tfs_analysis_data.parquet"))
-#### Data Overview ####
-# Structure
-head(fire_data)
-str(fire_data)
-
+# Since I want to study possible sources leading to different severity of fire, no info about sources should be dropped
+clean_data <- clean_data |>
+  filter(ignition_source != "999 - Undetermined")
+
+### Ensure Datetime Consistency ###
+
+# Convert timestamps to POSIXct format (YYYY-MM-DD HH:MM:SS)
+clean_data$tfs_alarm_time <- ymd_hms(clean_data$tfs_alarm_time)  
+clean_data$tfs_arrival_time <- ymd_hms(clean_data$tfs_arrival_time) 
+clean_data <- clean_data[!is.na(clean_data$tfs_alarm_time), ]
+clean_data <- clean_data[!is.na(clean_data$tfs_arrival_time), ]
+
+```
+
+## Variables Selection Criterion
+
+From this comprehensive dataset which included 43 vairables, I first selected the variables that are most relevant to my topic of study, leaving a total of 17 variables.
+Variables that are not essential to the study's focus include:
+
+**Geographical Information of Incidence**: The study focuses on analyzing fire data for Toronto as a whole.
+Geographic details such as the specific station area, ward, or intersection are not critical to determining severity levels.
+Including these could even divert focus from more relevant variables.
+
+-   `Latitude`, `Longitude`, `Incident_Station_Area`, `Incident_Ward`, and `Intersection`.
+
+**Some Safety System-Related Variables**: The study aims to study some infrasture system suhc as alarm and sprinkler systems.
+Some variables providing redundant information about these systems were excluded.
+The dataset also included variables on smoke alarm systems; however, due to the excessive number of null values, these variables were excluded as they made meaningful analysis challenging.
+
+-   `Fire_Alarm_System_Presence`, `Fire_Alarm_System_Impact_on_Evacuation`, `Smoke_Alarm_at_Fire_Origin`, `Smoke_Alarm_at_Fire_Origin_Alarm_Failure`, `Smoke_Alarm_at_Fire_Origin_Alarm_Type`, and `Sprinkler_System_Operation`.
+
+**Repeated Information**: Certain variables contained information that overlapped with variables already chosen for inclusion. To simplify the analysis, these redundant features were removed.
+
+-   `Count_of_Persons_Rescued`, `Estimated_Number_Of_Persons_Displaced`, `Material_First_Ignited`, `Initial_CAD_Event_Type`, `Level_Of_Origin`, `Extent_Of_Fire`, `Status_of_Fire_On_Arrival`, and `Property_Use`.
+
+**Unnecessary Details**: There are detailes in the dataset that is not directly related to the purpose of this study.
+Incident-specific details, such as how long firefighters took to respond or clear the fire, are more relevant for evaluating operational efficiency.
+These do not directly contribute to understanding the severity levels.
+
+-   `Ext_agent_app_or_defer_time`, `Fire_Under_Control_Time`, `Last_TFS_Unit_Clear_Time`, `Method_Of_Fire_Control`, and `Incident_Number`.
+
+Upon reviewing the 17 initially selected variables of interest, I observed that some columns contain a significant proportion of missing values (NAs), see @tbl-missing-overview.
+As a result, certain columns (with more than 20% missing data) should be considered for removal if their relevance to the study topic is not as significant as other variables that should remain.
+
+```{r}
+#| label: tbl-missing-overview
+#| fig-cap: This table shows the percentage of missing values for each variable of interest.
+Variables such as 'area_of_origin,' 'building_status,' and 'business_impact' exhibit high missing percentages, indicating potential areas for further investigation or data imputation, while others like '_id' and 'final_incident_type' have negligible missing values.
+Understanding these missing values is crucial for the completeness and quality of the analysis..
+#| warning: false
+#| message: false
+#| echo: false
 # Missing Value
-missing_data <- colSums(is.na(fire_data))
-print(missing_data)
-
-```
-
-This appendix contains all plot/summary distribution of all variables in interested choose in analysis data set.
-
-1.  Are of Origin (see @fig-area-origin-distribution)
-2.  Civilian Casualties (see @tbl-civilian-casualties-summary)
-3.  Estimated Dollar Loss (see @tbl-estimatedloss)
+missing_values <- colSums(is.na(clean_data))
+missing_percentage <- round((missing_values / nrow(clean_data)) * 100, 2)
+missing_info_table <- data.frame(Missing_Percentage = missing_percentage)
+kable(missing_info_table)
+```
+
+The variables removed are:
+
+1.
+**Building_Status**: This column provides information on the condition of the building, but it is not directly related to the core objectives of the study, which focus on fire severity, casualties, and financial losses.
+With more than 20% missing data, it was deemed unnecessary.
+
+2.
+**Business_Impact**: While this column tracks the impact of the fire on businesses, it overlaps with the **Estimated_Dollar_Loss** column, which provides more comprehensive financial data.
+Due to the high missing rate, this column was removed.
+
+3.
+**Exposures**: This column captures the number of exposure fires but does not directly contribute to understanding fire severity or its impact on casualties and losses.
+With more than 20% missing data, it was excluded from the analysis.
+
+Kept variables: These variables are essential for the completeness of the analysis. Since all of them are categorical data entries, missing values are replaced with “unreported.”
+
+-   **Fire_Alarm_System_Operation**: This variable is crucial because the study aims to understand fire infrastructure.
+
+-   **Sprinkler_System_Presence**: This is also an important variable for understanding fire safety. Given that there are fewer variables under "Safety Infrastructure," it is essential to retain them.
+
+After dealing with columns with a big proportion of missing values, removing rows with a certain amount of missing values would still remain in a relatively representative dataset.
+Therefore, I decided to remove any columns with more than 30% missing values to the usability of the dataset fro analysis.
+
+## Distribution and Summary of Cleaned Variables
+
+Below are the plots or summary statistics showing the distribution of all variables of interest in the analysis dataset, after cleaning and removing columns.
+This includes the created variable 'Response Time.'.
+The description of each variable is sourced from the data portal in Open Data Toronto \[@TFSfireedata\].
+
+1.
+**Area of Origin**: OFM Area of Origin code and description, specifying where the fire started.
+This information will contribute to the understanding what are some origins of fire and how it may influence fire severity or outcomes.
+75 different recorded origins, grouped into 11 larger groups.
+(see @fig-area-origin-distr)
+2.  **Civilian Casualties**: Count of civilian casualties (injuries or fatalities). This is directly relates to the severity of fire incidents.(see @tbl-civilian-casualties-summary) 
+3.  **Estimated Dollar Loss**: Estimated financial loss caused by the fire. This is directly relates to the severity of fire incidents. (see @tbl-estimatedloss)
 
@@
 
-4.  Final Incidence Type (see @fig-final-incident-type-distribution)
-5.  Fire Alram Conditions (see @fig-firealarmsystem)
+4.
+**Final Incidence Type**: Final classification of the fire incident This can  give information about the nature of the incident, helpful to determine severity.
+(see @fig-final-incident-type-distribution)
+5.
+**Fire Alarm System Operation**: Describes the operation of fire alarm systems during the incident.
+This is important for assessing the role of alarms, this is a part of fire infrastructure.(see @fig-firealarmsystem)
 
@@
 
-6.  Ignition Source (see @fig-grouped-ignition-source-distribution)
-7.  Number of Responding Apparatus (See @fig-responding-apparatus)
+6.
+**Ignition Source**: OFM Ignition Source code and description, detailing how the fire started.
+This is for understanding factors that contribute to fire ignition (see @fig-grouped-ignition-source-distribution)
+7.
+**Number of Responding Apparatus**: Number of TFS apparatus that responded to the incident.
+This may correlate with the severity as it reflects the scale of response.
+(See @fig-responding-apparatus)
 
@@
 
-8.  Number of Responding Personnel (See @fig-responding-personnel-hist)
+8.
+**Number of Responding Personnel**: Number of TFS personnel that responded to the incident.
+This reflect the magnitude of resources deployed, could relate with fire’s severity similar as `Number_of_Responding_Apparatus`.
+(See @fig-responding-personnel-hist)
 
@@
 
-9.  Possible Cause (See @tbl-possible-cause-full)
+9.  **Possible Cause**: OFM Possible Cause code and description, explaining the potential cause of the fire. This tells the circumstances leading to the fire. (See @tbl-possible-cause-full)
 
@@
 
-11. Sprinkler System Presence (@tbl-sprinkler-system)
+11.
+**Sprinkler System Presence**: Describes whether a sprinkler system was present at the location.
+This provided information of the sprinkler system, which is also a part of fire infrastructure.
+(see @tbl-sprinkler-system)
 
@@
 
-12. tfs_alarm_time (Amount of Fire Incidents by Year, see @fig-fire-incidents-by-year) (Fire Incidents by Month and Year, See @fig-year-month-incident-count)
+12.
+**TFS Alarm Time**: Time stamp of when TFS was notified of the incident.
+This useful when analyzing response times and their impact on severity, as well as learning about the amount of incidents across times.(Amount of Fire Incidents by Year, see @fig-fire-incidents-by-year) (Fire Incidents by Month and Year, see @fig-year-month-incident-count)
 
@@
 
-14. Firefighter Casualties (See @fig-firefighter-casulties)
+14. **Firefighter Casualties**: Count of TFS firefighter casualties (injuries or fatalities). This is useful when determining the severity of fire. (See @fig-firefighter-casulties)
 
@@
 
-15. Response Time (See @fig-response-time-distribution)
+15.
+**Response Time**: This is created variable.
+It is calculated as the difference between the `TFS_Alarm_Time` (when the fire service is notified of the incident) and the `TFS_Arrival_Time`.
+(See @fig-response-time-distribution)
 

### 2024-12-08 paper/paper.qmd
fix some cross-reference problem

@@
 Among all categories, Cooking and Dinning area stands out to be the area which most fire originate from.
-Living space and Transportation area also has a great amount to incidence occured.
+Living space and Transportation area also has a great amount to incidence occurred.
 
@@
 #| label: tbl-missing-overview
-#| fig-cap: This table shows the percentage of missing values for each variable of interest.
+#| tbl-cap: This table shows the percentage of missing values for each variable of interest.
 Variables such as 'area_of_origin,' 'building_status,' and 'business_impact' exhibit high missing percentages, indicating potential areas for further investigation or data imputation, while others like '_id' and 'final_incident_type' have negligible missing values.
@@
 75 different recorded origins, grouped into 11 larger groups.
-(see @fig-area-origin-distr)
+(see @fig-area-origin-distribution)
 2.  **Civilian Casualties**: Count of civilian casualties (injuries or fatalities). This is directly relates to the severity of fire incidents.(see @tbl-civilian-casualties-summary) 
@@
 #| label: fig-fire-incidents-by-year
-#| fig-caption: This bar chart shows the number of fire incidents by year, from 2011 to 2023.
+#| fig-cap: This bar chart shows the number of fire incidents by year, from 2011 to 2023.
 ## 12. tfs_alarm_time ## 
@@
   theme_minimal() +
-  scale_x_continuous(breaks = seq(min(fire_data$year), max(fire_data$year), by = 1))
+  scale_x_continuous(breaks = seq(min(fire_data$year), max(fire_data$year), by = 1)) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
@@
 #| label: fig-year-month-incident-count
-#| tbl-cap: Monthly Incident Count by Year (Excluding 2022 and 2023 becasue from prvious plot, we can see they don't contribute much information)
+#| fig-cap: Monthly Incident Count by Year (Excluding 2022 and 2023 becasue from prvious plot, we can see they don't contribute much information)
 
@@
   scale_x_discrete(
-    breaks = 1:12,  # Show all months (1 to 12)
-    labels = month.name  # Use full month names
+    breaks = 1:12,
+    labels = month.name 
   ) +

### 2024-12-11 paper/paper.qmd
Change graphs, include more details

@@
 # Data {#sec-data}
-
 ## Overview
@@
 #| echo: false
+#| fig.width: 3
+#| fig.height: 2
 
@@
   theme(
-    axis.text.x = element_text(angle = 45, hjust = 1), 
+    axis.text.x = element_text(angle = 45, hjust = 1,size =6), 
     plot.margin = margin(10, 10, 10, 10)  # Add space around the plot
@@
   ylab("Frequency") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5)
+  geom_text(stat = "count", aes(label = ..count..), vjust = 0.5, size = 2)
 
@@
 #| warning: false
+#| fig.width: 3
+#| fig.height: 2
 
@@
   ) +
-  theme_minimal()
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 20, hjust = 0.5,size =6)
+  )
 
@@
 # Model {#sec-model}
-
 ## Define Severity
@@
 
-## Model Explnation
+## Model Explanation
 
@@
 
+## Model Limitation
+The logistic regression model has several limitations.
+First, it only considers predictors such as area of origin and ignition source, excluding potentially significant factors like weather conditions and building materials, which could affect fire severity.
+Additionally, the model assumes that the relationships between predictors and severity are linear, which may not fully capture the complexity of fire incidents.
+Most importantly, the categorization of severity is based on predetermined thresholds, which may not be universally applicable, as different contexts may require different severity definitions.
+Lastly, the model's predictive accuracy is influenced by the quality of the data.
+
+Notably, there is a slight class imbalance in the dataset, high-severity incidents are underrepresented.
+In the training dataset, only 74 out of 10,237 records are classified as high severity.
+This imbalance can affect the model’s accuracy and its ability to predict high-severity incidents effectively.
+
 ## Model Justification
@@
 # Results {#sec-results}
-
 ## Logistic Regression Model Results
@@
 The results provide insights into the relationship between 'Area of Origin' and the likelihood of high-severity fire incidents."
-
+#| fig.align: left
+#| fig.width: 6
+#| fig.height: 4
 library(rstanarm)
@@
 
+rownames(coefficients_table) <- gsub(
+  "area_of_origin_grouped", "Area of Origin: ", 
+  gsub("([a-z])([A-Z])", "\\1 \\2", rownames(coefficients_table)) # Add space before capital letters
+)
+
 # Print or work with the coefficients
-kable(coefficients_table)
+kable(coefficients_table, digits = 2)
+
 ```
@@
 #| tbl-cap: The table summarizes the distribution of estimated dollar losses from fire incidents.
+#| fig.width: 3
+#| fig.height: 2
+
 library(knitr)
@@
 # (No plot code provided in the original snippet)
-
 ```
@@
 The date is right skewed, closer to 1.
+#| fig.width: 3
+#| fig.height: 2
+
 ## 7. Number of Responding Apparatus (Numerical) ##
@@
   ylab("Frequency") +
-  stat_bin(aes(label = ..count..), geom = "text", vjust = -0.5, size = 3)
+  stat_bin(aes(label = ..count..), geom = "text", vjust = -0.5, size = 2)
 
@@
 Similar dsitribution as @fig-responding-apparatus.
+#| fig.width: 3
+#| fig.height: 2
+
 # 8. Number of Responding Personnel (Numerical) ##
@@
   ylab("Frequency") +
-  stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 3)
+  stat_bin(aes(label = ..count..), geom = "text", vjust = 0.5, size = 2)
 
@@
 #| label: tbl-possible-cause-full
-#| tbl-cap: Comprehensive table showing all possible fire casues.
+#| tbl-cap: A comprehensive table showing top 6 possible fire casues.
+#| fig.width: 3
+#| fig.height: 2
+
 ## 9. Possible Cause ##
@@
   arrange(desc(Freq))
-kable(cause_table_df, col.names = c("Possible Cause", "Frequency"))
+kable(head(cause_table_df), col.names = c("Possible Cause", "Frequency"))
 ```
@@
 #| fig-cap: This bar chart shows the number of fire incidents by year, from 2011 to 2023.
+#| fig.width: 3
+#| fig.height: 2
+
 ## 12. tfs_alarm_time ## 
@@
   geom_bar(fill = "lightblue", color = "black") +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) + 
+  geom_text(stat = "count", aes(label = ..count..), vjust = 0, size = 2) + 
   labs(title = "Fire Incidents by Year", x = "Year", y = "Incident Count") +
@@
 #| fig-cap: Monthly Incident Count by Year (Excluding 2022 and 2023 becasue from prvious plot, we can see they don't contribute much information)
+#| fig.width: 5
+#| fig.height: 4
 
@@
   theme(
-    strip.text.x = element_text(size = 10), 
-    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.5, size = 8), 
-    axis.title.x = element_text(size = 12), 
-    axis.title.y = element_text(size = 12), 
+    strip.text.x = element_text(size = 6), 
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1.5, size = 4), 
+    axis.text.y = element_text(angle = 45, hjust = 1, vjust = 1.5, size = 4),
+    axis.title.x = element_text(size = 10), 
+    axis.title.y = element_text(size = 10), 
     strip.background = element_rect(fill = "lightblue", color = "black"), 
@@
   ) 
-
 ```
@@
 #| fig-cap: The Pie chart showing the distribution of firefighter casualties, rnaging from 0-3. With being the maximum amount of occurence.
+#| fig.width: 3
+#| fig.height: 2
 
@@
   labs(fill = "Casualties")
+
 #table
@@
 
+# Idealized Survey Methodology and Data Collection for Fire Severity Study in Toronto
+This study investigates the severity of fire incidents in Toronto to enhance resource allocation and optimize the effectiveness of the Toronto Fire Services (TFS).
+I work to understand key factors contributing to fire severity, including fire origin areas and ignition sources.
+I will also explore the effectiveness of public fire response systems, such as fire alarms sprinkler, smoke detection systems, to see how see thier presenace when fire incidence came and their effectiveness on reducsing fire.
+The methodology focuses on using observational data from fire incident records, public reports, and survey, and with statistical modeling to identify critical patterns and guide recommendations for TFS operations with understaingd of current fire infratcure presense.
+
 # References

### 2024-12-11 paper/paper.qmd
Add idealized methodlogy appendix

@@
 # Data {#sec-data}
+
 ## Overview
@@
 # Model {#sec-model}
+
 ## Define Severity
@@
 ## Model Limitation
+
 The logistic regression model has several limitations.
@@
 # Results {#sec-results}
+
 ## Logistic Regression Model Results
@@
 Priority should be given to maintaining fire alarm systems and improving the implementation of sprinkler systems.
-These measures will be essential for mitigating the risks and impacts of severe fire incidents.
+These measures will be essential for mitigating the risks and impacts of severe fire incidents, incorporating updated existing regulations and laws, and actions following the Government of Ontario's guidelines [@ontario_fire_code].
+.
 
@@
 This includes the created variable 'Response Time.'.
-The description of each variable is sourced from the data portal in Open Data Toronto \[@TFSfireedata\].
+The description of each variable is sourced from the data portal in Open Data Toronto [@TFSfireedata].
 
@@
 
+# Idealized Research Methodology
+
+## Methodology Overview
+
+This study investigates the severity of fire incidents in Toronto to enhance resource allocation and optimize the effectiveness of the Toronto Fire Services (TFS).
+The primary focus is on understanding key factors contributing to fire severity, including fire origin areas and ignition sources.
+Additionally, this study explores the effectiveness of public fire response systems.
+Important components that need to be considered include fire alarms, sprinkler systems, and smoke detection systems, examining their presence during fire incidents and their effectiveness in reducing fire damage [@thomas2002effectiveness].
+The methodology integrates data from fire incident records and surveys, incorporating random forest modelling techniques to identify critical patterns and guide recommendations for TFS operations in the future.
+
+## **Target Population**
+
+The target population for this study involves fire incidents in the city of Toronto.
+The key focus is on predicting fire severity based on various factors that influence the outcome of fire incidents and understanding the effectiveness of Toronto's fire infrastructure systems during these events.
+The population will include all reported fire incidents within the city, classified into three levels of severity: Low, Medium, and High.
+The population will encompass different building types where fire incidents occur, including but not limited to residential buildings, industrial factories, business establishments, and others.
+
+## Sampling Approach
+
+In the part of predicting fire severity, this study’s findings will be based on observational data from real fire incidents that occurred in Toronto.
+Therefore, there will not necessarily be sampling as in traditional surveys to collect data.
+Data will be gathered from official fire department records over a time period of 10 years.
+This time frame is chosen to ensure both inclusivity and relevance to the current state.
+The dataset will be divided into strata, subgroups of the population that have similar characteristics \[\@bailey1991strata\], based on different building types, with each stratum containing corresponding information, including fire origins, operational states of fire infrastructure systems, severity-related variables, and actions taken by the department.
+
+A survey will be used to understand the current fire infrastructure system.
+I will employ the stratified sampling method.
+Stratified sampling involves dividing the target population into distinct strata based on shared characteristics \[\@neyman1992two\], in this case, it would be building type, as explained in the methodology overview paragraph.
+From each subgroup, a representative sample will be randomly selected to ensure that all relevant perspectives are included.
+All citizens involved in fire incidents (for example, people caught in the fire, building managers, and neighbours) will have a chance to fill out a form.
+The survey will inquire about the presence, timeliness, effectiveness, and ease of use of fire infrastructure and fire-reducing facilities, such as public fire extinguishers and fire water taps.
+
+This survey data, when combined with records maintained by the Toronto Fire Services (TFS), will provide valuable insights into the current state of fire infrastructure.
+It will help identify gaps and prioritize areas where time and resources should be allocated to enhance the system's effectiveness in the future.
+
+### **Fire Incident Records:**
+
+-   **Data Source**: This is the primary data source for determining fire severities and making predictions about fire outcomes.
+The data will be sourced from the Toronto Fire Services (TFS) fire incident reports and the Ontario Fire Marshal’s (OFM) system.
+These records provide detailed information on each fire incident, such as the origin, ignition source, severity, and actions taken by firefighters, as well as details about the fire infrastructure systems involved.
+
+<!-- -->
+
+-   **Variables to Collect**: The classification of fire severity will be based on several factors, including the level of damage, the difficulty of firefighting tasks, injuries, fatalities, and property destruction.
+To predict fire severity features such as the fire origin area, ignition source, fire response system operation, and the time of year will be investigated.
+Ideally, information on the presence and functionality of fire alarms, sprinkler systems, smoke detection systems, and other fire infrastructure systems will also be noted.
+
+-   **Privacy Protection**: Privacy of participant information and data is important to all observational data collection \[\@preibusch2013guide\].
+To protect the privacy of those involved, the data will be anonymized.
+Specific addresses will be removed to prevent the identification of individuals or locations.
+Additionally, rare fire incident data will only be included after obtaining the consent of the individuals involved, ensuring their participation is voluntary.
+
+### **Survey on Real Experience with Fire Infrastructure Systems:**
+
+-   Surveys will be distributed using a stratified sampling method to individuals involved in fire incidents, ensuring coverage across all levels of severity and building types.
+These surveys will gather information on the effectiveness of fire response systems, such as fire alarms, sprinkler systems, and smoke detection systems, in reducing damage and protecting properties.
+The responses will help guide future improvements in fire infrastructure.
+
+-   **Participant Recruitment:**
+
+    -   Individuals who were directly involved in or reported the fire incident will be asked to provide their contact information, such as phone numbers, to ensure can reach them for participation.
+In general, building managers will distribute the survey to all individuals within the building who may have experienced or been impacted by the fire incident.
+This ensures the study reaches a wide range of individuals with relevant insights.
+
+    -   Since fire incidents involve people’s safety, participants are likely to be motivated to share their experiences and provide valuable feedback on fire operations and infrastructure.
+As a way to express appreciation, participants who complete the survey will be given the opportunity to enter a raffle for a prize, which will be drawn at the end of each year.
+This incentive is designed to encourage engagement while ensuring that participation remains voluntary \[\@freeman2007standards\].
+
+    -   All Participants will also be directed to a Toronto Fire Prevention website (<https://www.toronto.ca/community-people/public-safety-alerts/fire-prevention-inspection-enforcement/>) [@toronto_fire_prevention_2024], where they can access additional information and resources related to fire safety.
+
+-   **Survey Design:**
+
+    -   The survey is designed to gather both qualitative and quantitative data regarding fire safety infrastructure and its effectiveness.
+The questions are phrased in a neutral, unbiased manner, and all multiple-choice questions will include a "does not apply" or “other” option to ensure inclusivity and clarity for all respondents.
+To minimize the risk of trauma triggerings \[\@alessi2023toward\], participants will be informed at the beginning of the survey that they may choose not to answer any questions that make them uncomfortable.
+A reminder of available support resources will be provided at the end of the survey.
+Participants will also be reminded of their right to quit the survey at any time without any consequences.
+
+    -   **Demographic Information**:  This section includes questions about the participant’s location, type of building and their knowledge of fire safety systems, providing context for the responses.
+Sensitive questions about personal can skipped if the participant feels uncomfortable.
+
+    -   **Presence and Accessibility**: This section assesses the presence of essential fire safety features, such as fire alarms, extinguishers, sprinklers, and smoke detectors, and evaluates how easily accessible these resources are during a fire incident.
+These questions are more focused on infrastructure objectively to avoid causing distress.
+
+-   **Suggestions for Improvement**: An open-ended section where participants can suggest areas for improvement or highlight fire safety issues that need attention.
+This allows respondents to provide insights into how fire safety infrastructure can be enhanced in the future.
+Since this section encourages participants to share personal stories and thoughts, it will be optional, and participants can choose whether or not to fill it out.
+
+## **Data Validation**
+
+**Consistency**: To ensure the consistency of the data, I will compare survey responses with the Toronto Fire Services (TFS) fire records for each incident.
+For instance, I will compare the reported fire origin, as experienced by the individuals involved, with the fire origin recorded by TFS.
+If discrepancies are found between the two sources, inconsistent data will be removed.
+
+**Missing Data**: If a record is missing important information, such as the area of the fire, building details, or TFS response personnel, the record will be removed to ensure the quality of the dataset.
+In cases where operational system information is missing, for example, if the TFS report doesn’t specify details on the operation of fire systems of an incident, I will cross-check with survey responses to find the missing data for that specific incident.
+If there is inconsistency in the fire system status, such as conflicting responses between the survey and TFS data, or if no survey response is available for a particular incident, the operational data will be classified as "unreported" \[\@stavseth2019handling\].
+
+**Category Validation**: To ensure all recorded information aligns with clearly defined and consistent categories.
+This includes verifying that the fire incident categories, including severity levels, fire origin, and response actions are accurately defined and organized according to standard protocols.
+I will review the definitions and ensure that terminology is used consistently across all data sources.
+Additionally, I will ensure that the severity of the fire is defined in a way that is consistent with fire severity classification standards specific to Toronto.
+For example, the Office of the Fire Marshal of Ontario has set standards for fire prevention and reporting \[\@ontario_fire_marshal\].
+
+## **Modeling - Severity Prediction and Recommendations for Agencies and the Public**
+
+### **Model - Random Forests**
+
+-   The Random Forests model will be used to identify key predictors of fire severity.
+These models are an ensemble of multiple decision trees, which together provide predictions by combining the results of various decision paths for critical variables under certain thresholds \[\@breiman2001random\].
+The model will identify important factors influencing fire severity, and the predictions will include the weighting of factors based on the frequency and severity of past incidents.
+For example: even though high-severity fires may be less frequent, understanding the factors leading to these incidents is crucial, so weight adjustments will ensure that less frequent but high-impact factors are appropriately considered in the predictions.
+
+### **Model validation**
+
+-   To ensure the models generalize well and prevent overfitting, cross-validation techniques will be applied, this the total data will be separated into subsets, and the model will be rained on some subsets, and tested on others \[\@berrar2019cross\].  This will increase the reliability of the model.
+The model performance will be assessed using evaluation metrics such as ROC curves \[\@hoo2017roc\] which is a visual demonstration of the model’s sensitivity an specificity level.
+Other evaluation criteria include accuracy scores will also be included to ensure that the model’s predictions are reliable and effective.
+
+### **Applications and Benefits**:
+
+-   The results from the Random Forest model will inform fire agencies in predicting potential fire incidents.
+By analyzing the time of year and other features, the model can help agencies prepare for periods when high-severity fires are more likely.
+Additionally, the model can provide guidance on resource allocation.
+For example, if in a fire incident call, people indicate that a fire originated in a factory near electrical machinery, and the model predicts a high-risk fire in that area, the fire department can allocate more resources to manage the situation effectively.
+On the other hand, if a fire occurs in a residential building with the origin in a bathroom, and the model predicts a lower severity, fewer resources can be allocated, leading to more efficient use of time and resources.
+
+-   Knowledge gained from the model can be used to educate the public about areas with higher risks of severe fires.
+By targeting fire prevention measures and raising awareness in these high-risk locations, the city’s overall fire safety can be improved, reducing damage and loss of life.
+For example, if fire origins like HVAC systems or electrical equipment, are discovered to be strongly linked to high-severity fires, the fire agency can prioritize fire prevention plans in these areas.
+
+## **Survey Analysis - Effectiveness of Fire Infrastructure and Future Improvements**
+
+The survey can gather feedback on the presence, timeliness, effectiveness, and ease of use of various fire infrastructure and fire-reducing facilities, such as fire extinguishers, fire hoses, sprinkler systems, and smoke detectors \[\@ \[\@thomas2002effectiveness\]\].
+The responses will help identify areas where improvements can be made and assist agencies in understanding the current state of fire safety infrastructure.
+
+-   A combination of survey responses and records maintained by Toronto Fire Services (TFS) will provide valuable insights into the effectiveness of existing fire systems.
+By analyzing the data through categorizing responses; identifying geographical patterns; correlating feedback with fire severity; and analyzing open-ended responses, I can identify which systems are working well and where additional resources or improvements may be needed.
+This could include ensuring buildings are equipped with adequate alarm systems, performing regular maintenance on sprinklers, and enhancing other fire safety features.
+
+### **Applications and Benefits**:
+
+-   The survey result will guide TFS in directing resources and efforts toward improving fire infrastructure to increase public safety and reduce the impact of fire incidents.
+For example, if the survey indicates that smoke detectors are frequently reported as ineffective or wrongly capturing situations, TFS may focus on ensuring that all buildings are equipped with properly functioning smoke-detecting systems and conducting regular maintenance checks.
+Also, if the survey highlights that certain districts have insufficient access to fire extinguishers or water hoses, TFS can prioritize the installation of these essential fire safety resources in those locations to better equip both residents and responders in case of an emergency.
+
+### Link to the Survey:
+
+Here is a link to constructed survey: [https://forms.gle/gQcRL3YTL52S9xgR6](#0)
+
+## **Budget Allocation Breakdown**
+
+**Data Collection (Fire Incident Records & Surveys) -  (30% - \$30,000)**: Obtaining fire incident records, anonymizing data, designing surveys, testing surveys, participant recruitment, and the end-of-year raffle.
+
+-   **Participant End-of-Year Raffle - (5% - \$5,000)**: Funded as part of the recruitment budget to encourage survey participation, with prizes distributed at the end of the year.
+
+**Data Processing and Cleaning (15% - \$15,000)**: Cleaning and validating survey responses, hiring professionals for data review, integrating fire records with survey data, and ensuring privacy protection.
+
+**Model Development and Testing (25% - \$25,000)**: Developing random forest models, training and testing models, and cross-validation to ensure accuracy.
+
+**Analysis and Reporting (20% - \$20,000)**: Analyzing survey data and model outputs, conducting research to generate actionable insights, and generating comprehensive reports.
+
+**Project Management (10% - \$10,000)**: Coordination of the project, hiring personnel, managing stakeholders, and overseeing timelines.
+
 \newpage
 
-# Idealized Survey Methodology and Data Collection for Fire Severity Study in Toronto
-This study investigates the severity of fire incidents in Toronto to enhance resource allocation and optimize the effectiveness of the Toronto Fire Services (TFS).
-I work to understand key factors contributing to fire severity, including fire origin areas and ignition sources.
-I will also explore the effectiveness of public fire response systems, such as fire alarms sprinkler, smoke detection systems, to see how see thier presenace when fire incidence came and their effectiveness on reducsing fire.
-The methodology focuses on using observational data from fire incident records, public reports, and survey, and with statistical modeling to identify critical patterns and guide recommendations for TFS operations with understaingd of current fire infratcure presense.
-
 # References

### 2024-12-11 paper/paper.qmd
Adjust reference to metholodgy appendix

@@
 
-## **Target Population**
+## Target Population
 
@@
 This time frame is chosen to ensure both inclusivity and relevance to the current state.
-The dataset will be divided into strata, subgroups of the population that have similar characteristics \[\@bailey1991strata\], based on different building types, with each stratum containing corresponding information, including fire origins, operational states of fire infrastructure systems, severity-related variables, and actions taken by the department.
+The dataset will be divided into strata, subgroups of the population that have similar characteristics [@bailey1991strata], based on different building types, with each stratum containing corresponding information, including fire origins, operational states of fire infrastructure systems, severity-related variables, and actions taken by the department.
 
@@
 I will employ the stratified sampling method.
-Stratified sampling involves dividing the target population into distinct strata based on shared characteristics \[\@neyman1992two\], in this case, it would be building type, as explained in the methodology overview paragraph.
+Stratified sampling involves dividing the target population into distinct strata based on shared characteristics [@neyman1992two], in this case, it would be building type, as explained in the methodology overview paragraph.
 From each subgroup, a representative sample will be randomly selected to ensure that all relevant perspectives are included.
@@
 
-### **Fire Incident Records:**
+### Fire Incident Records:
 
@@
 
-<!-- -->
-
 -   **Variables to Collect**: The classification of fire severity will be based on several factors, including the level of damage, the difficulty of firefighting tasks, injuries, fatalities, and property destruction.
@@
 
--   **Privacy Protection**: Privacy of participant information and data is important to all observational data collection \[\@preibusch2013guide\].
+-   **Privacy Protection**: Privacy of participant information and data is important to all observational data collection [@preibusch2013guide].
 To protect the privacy of those involved, the data will be anonymized.
@@
 
-### **Survey on Real Experience with Fire Infrastructure Systems:**
-
--   Surveys will be distributed using a stratified sampling method to individuals involved in fire incidents, ensuring coverage across all levels of severity and building types.
+### Survey on Real Experience with Fire Infrastructure Systems:
+
+Surveys will be distributed using a stratified sampling method to individuals involved in fire incidents, ensuring coverage across all levels of severity and building types.
 These surveys will gather information on the effectiveness of fire response systems, such as fire alarms, sprinkler systems, and smoke detection systems, in reducing damage and protecting properties.
@@
 
--   **Participant Recruitment:**
-
-    -   Individuals who were directly involved in or reported the fire incident will be asked to provide their contact information, such as phone numbers, to ensure can reach them for participation.
+#### Participant Recruitment:
+
+Individuals who were directly involved in or reported the fire incident will be asked to provide their contact information, such as phone numbers, to ensure can reach them for participation.
 In general, building managers will distribute the survey to all individuals within the building who may have experienced or been impacted by the fire incident.
@@
 
-    -   Since fire incidents involve people’s safety, participants are likely to be motivated to share their experiences and provide valuable feedback on fire operations and infrastructure.
+Since fire incidents involve people’s safety, participants are likely to be motivated to share their experiences and provide valuable feedback on fire operations and infrastructure.
 As a way to express appreciation, participants who complete the survey will be given the opportunity to enter a raffle for a prize, which will be drawn at the end of each year.
-This incentive is designed to encourage engagement while ensuring that participation remains voluntary \[\@freeman2007standards\].
-
-    -   All Participants will also be directed to a Toronto Fire Prevention website (<https://www.toronto.ca/community-people/public-safety-alerts/fire-prevention-inspection-enforcement/>) [@toronto_fire_prevention_2024], where they can access additional information and resources related to fire safety.
-
--   **Survey Design:**
-
-    -   The survey is designed to gather both qualitative and quantitative data regarding fire safety infrastructure and its effectiveness.
+This incentive is designed to encourage engagement while ensuring that participation remains voluntary [@freeman2007standards].
+
+All participants will also be directed to a Toronto Fire Prevention website (<https://www.toronto.ca/community-people/public-safety-alerts/fire-prevention-inspection-enforcement/>), where they can access additional information and resources related to fire safety [@toronto_fire_prevention_2024].
+
+#### Survey Design:
+
+The survey is designed to gather both qualitative and quantitative data regarding fire safety infrastructure and its effectiveness.
 The questions are phrased in a neutral, unbiased manner, and all multiple-choice questions will include a "does not apply" or “other” option to ensure inclusivity and clarity for all respondents.
-To minimize the risk of trauma triggerings \[\@alessi2023toward\], participants will be informed at the beginning of the survey that they may choose not to answer any questions that make them uncomfortable.
+To minimize the risk of trauma triggerings [@alessi2023toward], participants will be informed at the beginning of the survey that they may choose not to answer any questions that make them uncomfortable.
 A reminder of available support resources will be provided at the end of the survey.
@@
 
-## **Data Validation**
+## Data Validation
 
@@
 In cases where operational system information is missing, for example, if the TFS report doesn’t specify details on the operation of fire systems of an incident, I will cross-check with survey responses to find the missing data for that specific incident.
-If there is inconsistency in the fire system status, such as conflicting responses between the survey and TFS data, or if no survey response is available for a particular incident, the operational data will be classified as "unreported" \[\@stavseth2019handling\].
+If there is inconsistency in the fire system status, such as conflicting responses between the survey and TFS data, or if no survey response is available for a particular incident, the operational data will be classified as "unreported" [@stavseth2019handling].
 
@@
 Additionally, I will ensure that the severity of the fire is defined in a way that is consistent with fire severity classification standards specific to Toronto.
-For example, the Office of the Fire Marshal of Ontario has set standards for fire prevention and reporting \[\@ontario_fire_marshal\].
-
-## **Modeling - Severity Prediction and Recommendations for Agencies and the Public**
-
-### **Model - Random Forests**
-
--   The Random Forests model will be used to identify key predictors of fire severity.
-These models are an ensemble of multiple decision trees, which together provide predictions by combining the results of various decision paths for critical variables under certain thresholds \[\@breiman2001random\].
+For example, the Office of the Fire Marshal of Ontario has set standards for fire prevention and reporting [@ontario_fire_marshal].
+
+## Modeling - Severity Prediction and Recommendations for Agencies and the Public
+
+### Random Forests
+
+The Random Forests model will be used to identify key predictors of fire severity.
+These models are an ensemble of multiple decision trees, which together provide predictions by combining the results of various decision paths for critical variables under certain thresholds [@breiman2001random].
 The model will identify important factors influencing fire severity, and the predictions will include the weighting of factors based on the frequency and severity of past incidents.
@@
 
-### **Model validation**
-
--   To ensure the models generalize well and prevent overfitting, cross-validation techniques will be applied, this the total data will be separated into subsets, and the model will be rained on some subsets, and tested on others \[\@berrar2019cross\].  This will increase the reliability of the model.
-The model performance will be assessed using evaluation metrics such as ROC curves \[\@hoo2017roc\] which is a visual demonstration of the model’s sensitivity an specificity level.
+### Model Validation
+
+To ensure the models generalize well and prevent overfitting, cross-validation techniques will be applied, this the total data will be separated into subsets, and the model will be rained on some subsets, and tested on others [@berrar2019cross].
+This will increase the reliability of the model.
+The model performance will be assessed using evaluation metrics such as ROC curves [@hoo2017roc] which is a visual demonstration of the model’s sensitivity an specificity level.
 Other evaluation criteria include accuracy scores will also be included to ensure that the model’s predictions are reliable and effective.
 
-### **Applications and Benefits**:
-
--   The results from the Random Forest model will inform fire agencies in predicting potential fire incidents.
+### Applications and Benefits:
+
+The results from the Random Forest model will inform fire agencies in predicting potential fire incidents.
 By analyzing the time of year and other features, the model can help agencies prepare for periods when high-severity fires are more likely.
@@
 
--   Knowledge gained from the model can be used to educate the public about areas with higher risks of severe fires.
+Knowledge gained from the model can be used to educate the public about areas with higher risks of severe fires.
 By targeting fire prevention measures and raising awareness in these high-risk locations, the city’s overall fire safety can be improved, reducing damage and loss of life.
@@
 
-## **Survey Analysis - Effectiveness of Fire Infrastructure and Future Improvements**
-
-The survey can gather feedback on the presence, timeliness, effectiveness, and ease of use of various fire infrastructure and fire-reducing facilities, such as fire extinguishers, fire hoses, sprinkler systems, and smoke detectors \[\@ \[\@thomas2002effectiveness\]\].
+## Survey Analysis - Effectiveness of Fire Infrastructure and Future Improvements
+
+The survey can gather feedback on the presence, timeliness, effectiveness, and ease of use of various fire infrastructure and fire-reducing facilities, such as fire extinguishers, fire hoses, sprinkler systems, and smoke detectors [@thomas2002effectiveness].
 The responses will help identify areas where improvements can be made and assist agencies in understanding the current state of fire safety infrastructure.
 
--   A combination of survey responses and records maintained by Toronto Fire Services (TFS) will provide valuable insights into the effectiveness of existing fire systems.
+A combination of survey responses and records maintained by Toronto Fire Services (TFS) will provide valuable insights into the effectiveness of existing fire systems.
 By analyzing the data through categorizing responses; identifying geographical patterns; correlating feedback with fire severity; and analyzing open-ended responses, I can identify which systems are working well and where additional resources or improvements may be needed.
@@
 
-### **Applications and Benefits**:
+### Applications and Benefits:
 
@@
 
-### Link to the Survey:
-
-Here is a link to constructed survey: [https://forms.gle/gQcRL3YTL52S9xgR6](#0)
+## Link to the Survey:
+
+Here is a link to constructed survey: <https://forms.gle/gQcRL3YTL52S9xgR6/>
 

### 2024-12-11 paper/datasheet.qmd
Datasheet update

---
title: "Datasheet for the Toronto Fire Incidents Analysis Dataset"
subtitle: "Supplementary document to the paper \"Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response\""
author: "Maggie Zhang"
thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
date: today
abstract: "This supplementary file accompanies the paper, \"Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response\". Following the structure outlined by [@gebru2021datasheets], this datasheet describes the motivation, composition, collection process, preprocessing/cleaning/labeling, uses, distribution, and maintenance of the cleaned analysis dataset. The dataset is available in the Repository: [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident), under data folder, named: tfs_analysis_data. The original data was sourced from the Open Data Toronto portal [@OpenDataToronto] and is titled Fire Incidents [@FireIncidents]."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

# Acknowledgement
The original dataset, titled Fire Incidents [@FireIncidents], was obtained from the Open Data Toronto portal [@OpenDataToronto]. Maggie Zhang conducted the cleaning, analysis, and presentation as part of the Toronto Fire Incident Analysis Project. The processed dataset and related materials can be found in the repository [TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident). 
This datasheet follows the structure outlined by [@gebru2021datasheets], with the questions in this document derived from their published paper.

# Motivation
1. **Purpose of Dataset Creation**: *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
   - The dataset was created to analyze the main factors influencing the severity of fire incidents in Toronto, while providing more in-depth insights into the current fire infrastructure system. Results are intended to better inform Toronto Fire Services (TFS) future decision-making and raise public awareness of fire incidents.

2. **Creators**: *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
   - The original dataset was provided through the Open Data Toronto portal [@OpenDataToronto] by the Ontario Fire Marshal (OFM) and Toronto Fire Services (TFS). Data preprocessing and analysis were conducted by Maggie Zhang, a third-year undergraduate student at the University of Toronto.

3. **Funding**: *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
   - No external funding was used for this analysis.

4. **Additional Comments**: *Any other comments?*
   - The dataset reflects information available as of the latest update from the Ontario Fire Marshal (OFM). The original dataset will be updated annually, currently consisting of 32,929 rows and 43 columns of variables, providing information on fire incidents responded to by Toronto Fire Services (TFS). The analysis dataset for this study is based on the updated data from 2024.

# Composition
1. **Instances Represented**: *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*

   - Each instance represents a fire incident that Toronto Fire Services responded to.

2. **Number of Instances**: *How many instances are there in total (of each type, if appropriate)?*
   - The analysis dataset contains 14,623 unique fire incidents.

3. **Dataset Type**: *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
   - The dataset can be considered complete with respect to the reported fire incidents during the collection period from 2011 to 2023, aside from some incidents removed for privacy protection purposes [@FireIncidents]. Therefore, it is representative of the larger population of fire incidents.
  - Personal data have been removed, and exact addresses are aggregated, following the Municipal Freedom of Information and Protection of Privacy Act (MFIPPA).
  - There are incidents that are still under investigation, thus some data points are missing (NAs).

4. **Features**: *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
   - The raw dataset consists of 32,929 rows and 43 columns of variables, providing information on fire incidents responded to by Toronto Fire Services (TFS).
   - In the cleaned analysis dataset for this study, there are 14,623 rows and 15 features. Each feature's name and description are listed below, extracted from the Fire Incident Data Portal [@FireIncidents].
   - id	: Unique row identifier for Open Data database
   - Area_of_Origin: OFM Area of Origin code and description
   - Civilian_Casualties: Count of civilian casualties. A causality can be a fire related injury or fire related fatality.
   - Count_of_Persons_Rescued: Number of persons rescued
   - Estimated_Dollar_Loss: Estimated Dollar Loss
   - Fire_Alarm_System_Operation: OFM Fire Alarm System Operation code and description
   - Ignition_Source: OFM Ignition Source code and description
   - Initial_CAD_Event_Type: First event type in CAD system of this incident
   - Number_of_responding_apparatus: Number of TFS responding apparatus
   - Number_of_responding_personnel: Number of TFS responding personnel
   - Possible_Cause: OFM Possible Cause code and description
   - Sprinkler_System_Presence: OFM Sprinkler System Presence code and description
   - TFS_Alarm_Time: Timestamp of when TFS was notified of the incident
   - TFS_Arrival_Time: Timestamp of first arriving unit to incident
   - TFS_Firefighter_Casualties: Count of TFS casualties. A causality can be a fire related injury or fire related fatality.
   - Response_Time (Created Variables): time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time 
   - Visualizations of all features are contained in the paper, Appendix A section.

5. **Labels**: *Is there a label or target associated with each instance? If so, please provide a description.*
   - Fire severity is the target variable; it was categorized based on three key factors: the number of responding fire apparatus, the estimated dollar loss, and the total number of casualties.

6. **Missing Information**: *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*

   - There is no missing information in the analysis fire incident dataset.
   - The original dataset includes all fire incidents as defined by the Ontario Fire Marshal (OFM) from 2011 to December 31, 2023.

7. **Relationships**: *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
   - `number_of_responding_personnel` and `number_of_responding_apparatus` are highly correlated. This can be explained by the fact that personnel usually carry a fixed number of apparatus.
   - `response_time` is related with `tfs_arrival_time` and `tfs_alarm_time` by the way its was defined

8. **Data Splits**: *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

   - The dataset has been split into training (80%) and testing (20%) subsets for analysis.

9. **Errors or Redundancies**: *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
   -  There are no redundancies in the dataset; each instance corresponds to a unique fire incident.
   - All variables in the dataset are categorical. Errors may exist in measurements and data entry, such as locations or ignition causes.
   - Bias may arise based on the category settings, as discussed in the Paper Discussion Section.

10. **External Resources**:  *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
   - The dataset does not rely on external resources.

11. **Confidentiality**: *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the original dataset removes all data that may trigger privacy issues.

12. **Safty**: *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- Yes. There is a record of casualties and fire situations. People with trauma or fear of such incidents might feel anxious.

13. **Sub-Population**: *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No. There is no information linked to specific groups of people in the analysis data, as the focus of the study is to understand the city of Toronto as a whole.

14. **Identifying Individuals**: *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No

15. **Sensitive Data**: *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.

16. *Any other comments?*
	- None.

# Collection Process
1. **Data Acquisition**: *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*

   - The dataset was obtained directly from the Open Data Toronto portal. The original data are directly from TFS reports to the OFM.

2. **Collection Mechanisms**: *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
   - Data were collected through reports by fire departments and digitized for public use.

3. **Ethical Review**: *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*

   - No ethical concerns were identified since the dataset contains no personal information and no specific geographical information.

4. **Source of Data**: *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*

   - Data are from Open Data Toronto [@OpenDataToronto], a publicly available resource.

5. **Consent**: *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*

   - Not applicable, as the data are from reports.

# Preprocessing, Cleaning, Labelling
1. **Steps Taken**: *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*

   - Store raw dataset in its original csv format and later analysis data after cleaning and processing is saved into parquet file using `write_parquet` function in `arrow` package [@arrow2024].
   - Columns of interest were selected, as explained in the Data section and Appendix A of the Paper. The column names were then standardized. 
   - Rows with "Undetermined" ignition sources were removed, as they would not provide useful information for predicting severity. 
   - Time-related variables were converted to the POSIX format. 
   - All columns with minimal information were checked, and columns with more than 20% missing data were removed, considering their relevance to the study. More details on relevance decisions can be found in Appendix A of the Paper.
   - Rows with more than 30% missing data were removed.
   - Create new variable, Response_Time

2. **Raw Data**: *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*

   - The original raw dataset is archived for reference. It is stored in the [01 Raw Data] (https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident/tree/main/data/01-raw_data).

3. **Software Used**: *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*

   - Preprocess/clean/label was conducted using R (@citeR). Pakages include: The `tidyverse`[@wickham2023tidyverse], `dplyr`[@wickham2023dplyr], and `tidyr`[@wickham2023tidyr] were used for data manipulation. `ggplot2`[@wickham2016ggplot2] facilitated data visualization, while `lubridate`[@grolemund2011lubridate] was used for date-time handling. `knitr`[@xie2023knitr] helped integrate code into reports.

4. *Any other comments?*
   - No

# Uses
1. **Current Uses**: *Has the dataset been used for any tasks already? If so, please provide a description.*

   - The original dataset is stored on an open data source; therefore, other studies can be conducted accordingly.

2. **Limitations**: *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*

   - No. However, the current analysis dataset is derived from incomplete fire incident records, some of which are still under investigation..

3. **Prohibited Uses**: *Are there tasks for which the dataset should not be used? If so, please provide a description.*

   - No, except that the dataset should not be used for commercial purposes without proper attribution.

4. *Any other comments?*
	 - None.

# Distribution
1. **Third Party**: *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	 - No.

2. **Distribution Method**: *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
   - The dataset will be shared through a public GitHub repository [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident) in its `data` folder.

3. **Date**: *When will the dataset be distributed?*
   - December 15th, 2024.

4. **License**: *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset will be distributed under the MIT License.

5. **Restrictions**: *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
   - No

6. **Export Controls**: *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.

7. *Any other comments?*
	- None.

# Maintenance
1. **Responsible Party**: *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset will be hosted in Maggie Zhang's GitHub repository: [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident) repository. Maggie will be responsible for maintaining the dataset.

2. **Contact Information**: *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
   - To contact Maggie, email: maggie.zhang@mail.utoronto.ca.
   - To access the original dataset, visit the Open Data Toronto Portal: [@FireIncidents].

3. **Erratum**: *Is there an erratum? If so, please provide a link or other access point.*
	 - No.

4. **Updates**: *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
   - No further updates to the analysis dataset for this study are planned after the release date.
   - The original dataset on Open Data Toronto is refreshed annually.

5. **Collaborations**: *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
   - Collaboration inquiries are welcome via email to Maggie. If they want to build on or contribute to the dataset, they can fork the public GitHub repository: [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)
   - If they are interested in the dataset and would like to start from the original data, they can access it through the Open Data Toronto Portal at Fire Incident Data [@FireIncidents].

6. *Any other comments?*
	- None.

# References

### 2024-12-11 paper/datasheet.qmd
update datasheet with link to ODT

@@
 author: "Maggie Zhang"
-thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
+thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident); Original fire incidents dataset can be accessed from: [https://open.toronto.ca/dataset/fire-incidents/](https://open.toronto.ca/dataset/fire-incidents/)."
 date: today

### 2024-12-11 paper/paper.qmd
Add datasheet section to paper

@@
 thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
-date: "December 3, 2024"
+date: today
 date-format: long
@@
 [@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements.
-Finally, [@sec-appendix] includes supplementary information and additional data details.
+Finally, [@sec-appendixa] includes details about variable selection and cleaning, with the visualization of all variables in the analysis dataset; [@sec-appendixb] is a supplementary datasheet that explains the building of the analysis dataset; [@sec-appendixc] outlines the idealized methodology for conducting this study, given a budget of $100K.
 
@@
 However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
-A more detailed explanation will be provided in the @sec-appendix.
+A more detailed explanation will be provided in the @sec-appendixa.
 
@@
 The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
-Can check @sec-appendix for visualization/summary of the distribution of all variables.
+Can check @sec-appendixa for visualization/summary of the distribution of all variables.
 Below are a few important variables considered:
@@
 We can see that most of the time, there are no casualties.
-This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendix), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
+This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
 
@@
 
-# Variable Selection and Visualization
+# Variable Selection and Visualization {#sec-appendixa}
 
@@
 
-# Idealized Research Methodology
+# Supplemental Datasheet {#sec-appendixb}
+
+A datasheet for the dataset used in this analysis is included with this paper. Please
+refer to this link to the [Datasheet](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident/blob/main/paper/datasheet.pdf).
+
+# Idealized Research Methodology {#sec-appendixc}
 

### 2024-12-11 paper/datasheet.qmd
remove one overlapping reference

@@
 The dataset is available in the Repository: [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident), under data folder, named: tfs_analysis_data.
-The original data was sourced from the Open Data Toronto portal [@OpenDataToronto] and is titled Fire Incidents [@FireIncidents]."
+The original data was sourced from the Open Data Toronto portal [@OpenDataToronto] and is titled Fire Incidents [@TFSfireedata]."
 format: pdf
@@
 # Acknowledgement
-The original dataset, titled Fire Incidents [@FireIncidents], was obtained from the Open Data Toronto portal [@OpenDataToronto].
+The original dataset, titled Fire Incidents [@TFSfireedata], was obtained from the Open Data Toronto portal [@OpenDataToronto].
 Maggie Zhang conducted the cleaning, analysis, and presentation as part of the Toronto Fire Incident Analysis Project.
@@
 If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
-   - The dataset can be considered complete with respect to the reported fire incidents during the collection period from 2011 to 2023, aside from some incidents removed for privacy protection purposes [@FireIncidents].
+   - The dataset can be considered complete with respect to the reported fire incidents during the collection period from 2011 to 2023, aside from some incidents removed for privacy protection purposes [@TFSfireedata].
 Therefore, it is representative of the larger population of fire incidents.
@@
    - In the cleaned analysis dataset for this study, there are 14,623 rows and 15 features.
-Each feature's name and description are listed below, extracted from the Fire Incident Data Portal [@FireIncidents].
+Each feature's name and description are listed below, extracted from the Fire Incident Data Portal [@TFSfireedata].
    - id	: Unique row identifier for Open Data database
@@
    - To contact Maggie, email: maggie.zhang@mail.utoronto.ca.
-   - To access the original dataset, visit the Open Data Toronto Portal: [@FireIncidents].
+   - To access the original dataset, visit the Open Data Toronto Portal: [@TFSfireedata].
 
@@
 If they want to build on or contribute to the dataset, they can fork the public GitHub repository: [TFS_Toronto_Fire_Incident_data_folder](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)
-   - If they are interested in the dataset and would like to start from the original data, they can access it through the Open Data Toronto Portal at Fire Incident Data [@FireIncidents].
+   - If they are interested in the dataset and would like to start from the original data, they can access it through the Open Data Toronto Portal at Fire Incident Data [@TFSfireedata].
 

### 2024-12-11 paper/paper.qmd
Edit abtract

@@
 
-We use the statistical programming language R [@citeR]to investigate data about fire incidents[@TFSfireedata], downloaded from the Open Data Toronto Portal: https://open.toronto.ca/dataset/fire-incidents/ .
+We use the statistical programming language R [@citeR]to investigate data about fire incidents, downloaded from the Open Data Toronto Portal [@OpenDataToronto], named [Fire Incidents](https://open.toronto.ca/dataset/fire-incidents/) [@TFSfireedata].
 The dataset was last updated on December 31, 2023.

### 2024-12-14 paper/paper.qmd
adjust table size

--- 2024-12-11 paper/paper.qmd
@@
 abstract: "This study investigates fire severity levels in Toronto to improve resource allocation for Toronto Fire Services (TFS).
-The analysis identifies that higher-severity incidents are strongly correlated with incidents originating from mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires.
-The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment, are more likely to be severe.
-The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are successful in low and medium-severity fires, while sprinkler systems are often only partially operational.
+The analysis identifies that higher-severity incidents are strongly correlated with fires originating in mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires.
+The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment,  are more likely to cause severe fires.
+The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are effective in low and medium-severity fires, while sprinkler systems are often only partially operational.
 Results can guide TFS to enhance response times, prioritize resources, and optimize fire response operations, ultimately improving community safety in Toronto."
@@
 
-We use the statistical programming language R [@citeR]to investigate data about fire incidents[@TFSfireedata], downloaded from the Open Data Toronto Portal: https://open.toronto.ca/dataset/fire-incidents/ .
+I use the statistical programming language R [@citeR]to investigate data about fire incidents, downloaded from the Open Data Toronto Portal [@OpenDataToronto], named [Fire Incidents](https://open.toronto.ca/dataset/fire-incidents/) [@TFSfireedata].
 The dataset was last updated on December 31, 2023.
@@
 The dataset also provides detailed information on Toronto Fire Services' (TFS) responses to each fire incident.
-The data is mostly complete, though some incidents are still under investigation and categorized as "no loss outdoor fire." The dataset includes information on the origin of fires, impacts on civilians and buildings, as well as geographical data (see @tbl-preview-data).
+The data is mostly complete, though some incidents are still under investigation.
+Feature includes specification on the origin of fires, impacts on civilians and buildings, as well as geographical data (see @tbl-preview-data).
 Additionally, it contains details on TFS response times and actions taken.
@@
 #| label: tbl-preview-data
-#| tbl-cap: Preview of the Fire Incident data (https://open.toronto.ca/dataset/fire-incidents/) before cleaning, displaying the first six rows with columns decided first in interested for study.
+#| tbl-cap: Preview of the Fire Incident data (https://open.toronto.ca/dataset/fire-incidents/) before cleaning, displaying the first four incidents with columns decided first in interested for study.
 #| echo: false
@@
 
-#### Load original data ####
+## Load original data ##
 raw_data <- read_csv(file = here::here("data/01-raw_data/raw_fire_data.csv"))
@@
 
-### Preview data ###
+## Preview data ##
 head_table <- head(clean_data)
-kable(head(head_table)) |>
-  kable_styling(latex_options = c("scale_down"))
+kable(t((clean_data)[1:4,])) |>
+  kable_styling(latex_options = c("scale_down"), font_size = 7) |>
+  column_spec(1, width = "3.5cm") |>
+  column_spec(c(2, 3, 4, 5), width = "3cm")
 
@@
 However, the key question remains: how well do these fire safety infrastructures perform in real fire incidents?
-We come across result of question by looking at the result we get from analyzing the data set.
+I come across result of question by looking at the result get from analyzing the data set.
 In this analysis, I processed the data and visualized the distribution of fire alarm system operations and sprinkler system presence across different levels of fire severity (Low, Medium, High).
@@
 This stacked bar plot shows the relative frequencies of various fire alarm system operations within each severity group.
-We noctice that fire alarm was not always operated in high severity incidents, but not active in low severity fire.
+Can noctice that fire alarm was not always operated in high severity incidents, but not active in low severity fire.
 

### 2024-12-15 paper/paper.qmd
revise title, abstract, intro

@@
 title: "Investigating Fire Incidents Severity Levels in Toronto: Enhancing Resource Allocation for Fire Response"
-subtitle: "Identifying Key Factors for Fire Severity and Evaluating the Effectiveness of Public Fire Response Systems"
+subtitle: "Key Role of Origin of Fire and Opportunities to Improve the Effectiveness of Toronto's Public Fire Infrastructure"
 author: 
   - Maggie Zhang
-thanks: "Code and data are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
+thanks: "Code and data supporting this analysis are available at: [https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident](https://github.com/MaggieZ111119/TFS_Toronto_Fire_Incident)."
 date: today
@@
 abstract: "This study investigates fire severity levels in Toronto to improve resource allocation for Toronto Fire Services (TFS).
-The analysis identifies that higher-severity incidents are strongly correlated with fires originating in mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to lower severity fires.
-The study also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment,  are more likely to cause severe fires.
-The study also evaluates the effectiveness of public fire response systems, highlighting that fire alarms are effective in low and medium-severity fires, while sprinkler systems are often only partially operational.
-Results can guide TFS to enhance response times, prioritize resources, and optimize fire response operations, ultimately improving community safety in Toronto."
+The analysis identifies that higher-severity incidents are strongly correlated with fires originating in mechanical, HVAC, and electrical areas, while cooking and lighting areas tend to lead to more frequent but lower-severity fires.It also finds that fire incidents involving certain ignition sources, such as electrical units or lightning equipment, are more likely to cause severe fires.
+Additionally, the study evaluates the effectiveness of public fire response systems, highlighting that fire alarms are effective in low and medium-severity fires, while sprinkler systems are often only partially operational.
+These findings can assist TFS in enhancing response strategies, prioritizing resources, and optimizing fire response operations, ultimately improving community safety in Toronto."
 format: pdf
@@
 
-The study aims to estimate the severity of fires in the Toronto region.
-A fire incident, as defined by the Ontario Fire Marshal (OFM), refers to any event involving fire, explosion, or other situations suspected to lead to significant outcomes, such as injury or death [@ontario_fire_marshal].
-
-The severity of a fire can be assessed through various factors, including the intensity of the flames, the size of the affected area, the duration of the fire, its impact on human activities, and its potential threat to animal habitats.
+The study aims to estimate the severity of fires in the Toronto city A fire incident, as defined by the Ontario Fire Marshal (OFM), refers to any event involving fire, explosion, or other situations suspected to lead to significant outcomes, such as injury or death [@ontario_fire_marshal].
+
+The severity of a fire can be assessed through various factors, including the intensity of the flames, the size of the affected area, its impact on human activities, and its potential threat to animal habitats.
 There are numerous standards and considerations when discussing fire severity, but when it comes to prediction, it is unrealistic to consider all factors.
 Therefore, defining severity becomes the first task before any further analysis.
-In this paper, the severity of a fire is defined based on three aspects: casualty and financial loss, the type of incident, and the number of apparatuses dispatched by Toronto Fire Services (TFS) in response to the fire.
+In this study, the severity of a fire is defined based on three aspects: casualty and financial loss, the type of incident, and the number of apparatuses dispatched by Toronto Fire Services (TFS) in response to the fire.
 Further explanation of how severity is defined will be provided later in the paper.
 
-This research also identifies specific high-priority response criteria that can enhance TFS operations.
+This study also identifies high-priority response criteria that can assist in enhancing TFS' future operations.
 By examining the data, it has been found that the area of origin of a fire is a significant predictor of fire severity.
-Fires originating from mechanical, HVAC, and electrical areas are highly correlated with severe incidents, whereas fires originating from cooking and heating areas are more commonly associated with low-severity incidents.
-
-Additionally, beyond investigating potential predictors of severe fire incidents, this study also explores the effectiveness of current public fire response mechanisms, such as fire alarms and sprinkler systems, in the Toronto region.
-It was discovered that sprinkler systems, across all levels of fire severity, were largely only partially operational.
-On the other hand, fire alarm systems were successfully activated in low and medium-severity fire incidents.
+Fires originating from mechanical, HVAC, and electrical areas are highly correlated with severe incidents, whereas fires originating from cooking and heating areas are more commonly associated with low-severity incidents though they are more frequent.
+
+Additionally, this study explores the effectiveness of current public fire response mechanisms in Toronto, such as fire alarms and sprinkler systems.
+It was found that sprinkler systems, across all levels of fire severity, were mostly only partially operational.
+Fire alarm systems were successfully activated in low and medium-severity fire incidents but there is room for improvement in higher-severity incidents.
 
 The Office of the Fire Marshal (OFM) plays a crucial role in overseeing fire safety regulations in Ontario, ensuring fire protection, prevention, and public safety across the province.
-They are responsible for setting safety standards and improving public safety for all municipalities.
-Under the leadership of the OFM, municipal fire departments, such as Toronto Fire Services (TFS), allocate resources and develop plans to protect shared communities [@ontario_fire_marshal].
-Therefore, by analyzing data within their standards, identifying factors that contribute to high-risk fire incidents, and evaluating current operational mechanisms, TFS can gain valuable insights to develop strategic response protocols that prioritize these incidents.
-This knowledge will help both the public and fire departments enhance response times, allocate resources more effectively, and reduce the occurrence of lower-severity incidents.
+It is responsible for setting safety standards and improving public safety in all municipalities.
+Under the leadership of the OFM, municipal fire departments, such as Toronto Fire Services (TFS), allocate resources and develop plans to protect our communities [@ontario_fire_marshal].
+Therefore, by analyzing data within their standards, identifying factors that contribute to high-risk fire incidents, and evaluating current operational mechanisms, TFS can gain valuable insights to develop strategic protocols that prioritize certain incidents.
+These findings will help both the public and fire departments enhance response to incidents, allocate resources more effectively, and reduce the occurrence of controllable incidents.
 
 The remainder of this paper is structured as follows.
-[@sec-data] describes the data used in the analysis, including the sources and key variables and their visualizations.[@sec-model] outlines the models applied to estimate fire severity and identify key factors.
-[@sec-results] presents the results of the analysis.
-[@sec-discussion] provides an interpretation of the results in the context of current fire safety practices and suggests potential improvements.
-Finally, [@sec-appendixa] includes details about variable selection and cleaning, with the visualization of all variables in the analysis dataset; [@sec-appendixb] is a supplementary datasheet that explains the building of the analysis dataset; [@sec-appendixc] outlines the idealized methodology for conducting this study, given a budget of $100K.
-
-In this paper, I used several R packages for data analysis and visualization.
+Data [@sec-data] describes the data used in the analysis, including the sources and key variables along with their visualizations.
+Model [@sec-model] outlines the models applied to estimate fire severity and identify key factors.
+Result [@sec-results] presents the analysis results with supporting graphs and tables.
+Discussion [@sec-discussion] provides interpretations of the results in the context of current fire safety practices and suggests future improvements.
+Finally, Appendix [@sec-appendixa] includes details about variable selection criterion and data cleaning, with the visualization of all variables in the analysis dataset; Appendix [@sec-appendixb] is a supplementary datasheet that explains the building of the analysis dataset; Appendix [@sec-appendixc] outlines my idealized methodology for conducting this study, given a budget of $100K.
+
+In this paper, analyzing with R (@citeR), I used several packages throughout the analysis.
 The `tidyverse`[@wickham2023tidyverse], `dplyr`[@wickham2023dplyr], and `tidyr`[@wickham2023tidyr] were used for data manipulation.
@@
 It includes fire incidents following the Ontario Fire Marshal's (OFM) definition, with descriptions according to OFM standards.
-The dataset also provides detailed information on Toronto Fire Services' (TFS) responses to each fire incident.
+The dataset also provides detailed information on Toronto Fire Services (TFS) responses to each fire incident.
 The data is mostly complete, though some incidents are still under investigation.
@@
 The dataset is clearly structured, with the first column as ‘\_id,’ which is a unique identifier for each record in the database.
-There are 43 columns in total, containing information ranging from Incident Type to Infrastructure.
-These columns can be mainly divided into four categories based on my understanding:
+There are 43 columns in total in the original set, containing information ranging from incident type to infrastructure's presence.
+These columns can be divided into four categories based on my understanding:
 
 1.
-**Incident Type and Cause**: This category contains information directly related to each fire incident, including descriptive details about the incident, its nature, and origin.
+**Incident Type and Cause**: This category provides information directly related to each fire incident, including descriptive details about the incident, its nature, and origin.
 It covers aspects such as where the fire started, the materials involved, and potential causes.
@@
 This category is key for defining the severity of fires.
-Variables include: firefighter casualties, civilian casualties, building status after the fire, and the number of displaced people.
+Variables include firefighter casualties, civilian casualties, building status after the fire, and the number of displaced people.
 
@@
 It helps us understand the current system and supports decision-making for future improvements.
-Variables include: presence of fire alarm systems, sprinkler systems, and their operational status.
-
-This structured approach allows us to analyze and understand various aspects of fire incidents in Toronto.
+Variables include the presence of fire alarm systems, sprinkler systems, and their operational status.
+
+This dividing approach helps analyze and understand rounded aspects of fire incidents in Toronto.
 All variables in the dataset have the potential to provide valuable insights and perspectives for various analyses, contributing to fire prevention and safety improvements.
-However, for my analysis of fire incidents across Toronto, certain variables; such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
+However, for my focus on analyzing fire incidents across Toronto, certain variables, such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
 A more detailed explanation will be provided in the @sec-appendixa.
 
-I have selected 17 variables to begin my exploration of the dataset:
-
-1.  **\_id**: A unique identifier for each record, essential for distinguishing between incidents and for tracking and referencing.
-
-2.  **Area_of_Origin**: Represents the fire’s origin location, useful for analyzing geographical patterns and fire causes.
-
-3.  **Building_Status**: Indicates whether the building was occupied, under construction, or vacant, providing insights into fire severity and response needs.
-
-4.  **Business_Impact**: Classifies the economic impact of the fire on businesses, helping to understand disruptions and guide recovery efforts.
-
-5.  **Civilian_Casualties**: The count of civilian casualties, reflecting the human impact and assessing fire safety and evacuation effectiveness.
-
-6.  **Estimated_Dollar_Loss**: The monetary loss due to fire damage, key for assessing the economic burden and informing prevention efforts.
-
-7.  **Exposures**: The number of secondary fires caused by the primary incident, useful for evaluating the fire's spread and containment success.
-
-8.  **Final_Incident_Type**: The final classification of the fire (e.g., accidental, arson), critical for identifying causes and informing preventive strategies.
-
-9.  **Fire_Alarm_System_Operation**: Indicates whether the fire alarm system functioned properly, evaluating the system’s effectiveness in alerting and activating suppression.
-
-10. **Ignition_Source**: The source of ignition (e.g., electrical faults, arson), crucial for identifying fire causes and guiding safety measures.
-
-11. **Number_of_Responding_Appliances**: The number of fire trucks and equipment dispatched, reflecting the scale of the incident and resource allocation.
-
-12. **Number_of_Responding_Personnel**: The number of fire personnel involved, providing insight into the severity of the fire and the required response intensity.
-
-13. **Possible_Cause**: The potential cause of the fire, useful for identifying trends and informing fire prevention and safety campaigns.
-
-14. **Sprinkler_System_Presence**: Indicates if a sprinkler system was present, highlighting its impact on fire severity and potential casualties.
-
-15. **TFS_Alarm_Time**: The time when the fire service was notified, key for analyzing response time and minimizing damage.
-
-16. **TFS_Arrival_Time**: The time when the first fire service unit arrived, crucial for evaluating response speed and fire control effectiveness.
-
-17. **TFS_Firefighter_Casualties**: The number of firefighter casualties, useful for assessing the safety of firefighting operations and improving risk management.
+I selected 17 variables to begin my exploration of the dataset:
+
+```{r}
+# This is only a table help to structure layout of paper formatting, no statistical analysis information involve, thus no caption
+#| label: tbl_variables_description
+#| echo: false
+#| warning: false
+
+variable_table <- data.frame(
+  Variable = c(
+    "_id", 
+    "Area_of_Origin", 
+    "Building_Status", 
+    "Business_Impact", 
+    "Civilian_Casualties", 
+    "Estimated_Dollar_Loss", 
+    "Exposures", 
+    "Final_Incident_Type", 
+    "Fire_Alarm_System_Operation", 
+    "Ignition_Source", 
+    "Number_of_Responding_Appliances", 
+    "Number_of_Responding_Personnel", 
+    "Possible_Cause", 
+    "Sprinkler_System_Presence", 
+    "TFS_Alarm_Time", 
+    "TFS_Arrival_Time", 
+    "TFS_Firefighter_Casualties"
+  ),
+  Description = c(
+    "A unique identifier for each record, essential for distinguishing between incidents and for tracking and referencing.",
+    "Represents the fire’s origin location, useful for analyzing geographical patterns and fire causes.",
+    "Indicates whether the building was occupied, under construction, or vacant, providing insights into fire severity and response needs.",
+    "Classifies the economic impact of the fire on businesses, helping to understand disruptions and guide recovery efforts.",
+    "The count of civilian casualties, reflecting the human impact and assessing fire safety and evacuation effectiveness.",
+    "The monetary loss due to fire damage, key for assessing the economic burden and informing prevention efforts.",
+    "The number of secondary fires caused by the primary incident, useful for evaluating the fire's spread and containment success.",
+    "The final classification of the fire (e.g., accidental, arson), critical for identifying causes and informing preventive strategies.",
+    "Indicates whether the fire alarm system functioned properly, evaluating the system’s effectiveness in alerting and activating suppression.",
+    "The source of ignition (e.g., electrical faults, arson), crucial for identifying fire causes and guiding safety measures.",
+    "The number of fire trucks and equipment dispatched, reflecting the scale of the incident and resource allocation.",
+    "The number of fire personnel involved, providing insight into the severity of the fire and the required response intensity.",
+    "The potential cause of the fire, useful for identifying trends and informing fire prevention and safety campaigns.",
+    "Indicates if a sprinkler system was present, highlighting its impact on fire severity and potential casualties.",
+    "The time when the fire service was notified, key for analyzing response time and minimizing damage.",
+    "The time when the first fire service unit arrived, crucial for evaluating response speed and fire control effectiveness.",
+    "The number of firefighter casualties, useful for assessing the safety of firefighting operations and improving risk management."
+  )
+)
+
+kable(variable_table) |> 
+  kable_styling(font_size = 6) |>
+  column_spec(2, width = "9cm")
+
+```
 
@@
 
-The raw dataset was stored in its original csv format to ensure data integrity and reproducibility, and later analysis data after cleaning and processing is saved into parquet file using `write_parquet` function in `arrow` package [@arrow2024].
-
-After loading the original raw data, I selected the columns of interest as explained in the previous section.
-These columns were then renamed to lowercase, and spaces were replaced with underscores to ensure consistency.
-After reviewing the data, I kept in mind that this study focuses on the relationship between ignition sources and fire severity.
-As a result, I removed rows with "Undetermined" ignition sources.
-
-Since there are time-related variables in the dataset, I converted them to the POSIX format.
-Additionally, I noticed the presence of several missing (NA) and null values.
-Given that the dataset contains numerous categorical variables, I considered how to handle missing data differently for these types.
+The raw dataset was stored in its original CSV format to ensure data integrity and reproducibility, and later analysis data after cleaning and processing was saved into a parquet file using the `write_parquet` function in the `arrow` package [@arrow2024].
+After loading and cleaning the data, I retained only the relevant columns and renamed them for consistency.
+I removed rows with 'Undetermined' ignition sources, ensuring that the dataset focused on the relationship between ignition sources and fire severity following the study topic.
+
+I converted time-related variables to the POSIX format.
+Moreover, I noticed the presence of quite an amount of missing values (NAs) and NULL values.
+Given that the dataset contains numerous categorical variables, I considered how to handle them differently for these types.
 I first examined whether any columns contained too little information to provide valuable insights into fire severity or its predictors.
@@
 To maintain data quality, rows with more than 30% missing data are removed.
-After ensuring that there are no missing values in the dataset, a new variable, Response_Time, is created to calculate the time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time (see @fig-response-time-distribution).
+With complete data, a new variable named `Response_Time` is created to calculate the time difference (in minutes) between TFS_Alarm_Time and TFS_Arrival_Time (see @fig-response-time-distribution).
 This new variable will help analyze fire response efficiency and is a key indicator for the study.
+As shown in the plot, TFS response times center around 5 minutes, with less variation towards longer times.
 
@@
   theme_minimal() +
-  ggtitle("Response Time Distribution") +
   xlab("Response Time (minutes)") +

### 2024-12-15 paper/paper.qmd
update data, model, result

@@
 However, for my focus on analyzing fire incidents across Toronto, certain variables, such as latitude and longitude, incident numbers in the CAD system, Incident_Ward, and road intersections—are being excluded from the study.
-A more detailed explanation will be provided in the @sec-appendixa.
+A more detailed explanation will be provided in the Appendix @sec-appendixa.
 
@@
 ```{r}
-# This is only a table help to structure layout of paper formatting, no statistical analysis information involve, thus no caption
 #| label: tbl_variables_description
@@
 
+# This is only a table help to structure the layout of paper formatting, no statistical analysis information involve, thus no caption
 variable_table <- data.frame(
@@
 kable(variable_table) |> 
-  kable_styling(font_size = 6) |>
+  kable_styling(font_size = 6.5) |>
   column_spec(2, width = "9cm")
@@
 #| warning: false
+#| fig.width: 5
+#| fig.height: 3
+
 library(ggplot2)
@@
 **Understanding Variables of Interest**:\
-To understand the distribution of various categorical and numerical variables in the fire incident dataset, I visualized each variable along with its distribution and reviewed its summary statistics.
-The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modeling.
-Can check @sec-appendixa for visualization/summary of the distribution of all variables.
+To understand the distribution of various categorical and numerical variables in the fire incident dataset, I analyzed each variable along with its distribution and reviewed its summary statistics.
+The goal is to gain insights into the key characteristics of the data, which will support the later stages of analysis and modelling.
+For a comprehensive overview of all chosen variables, including their visualizations and summaries, refer to Appendix @sec-appendixa.
 Below are a few important variables considered:
@@
 These groups were then visualized to better understand the most common fire origins within the dataset (see @fig-area-origin-distribution).
-Among all categories, Cooking and Dinning area stands out to be the area which most fire originate from.
-Living space and Transportation area also has a great amount to incidence occurred.
+Among all categories, the Cooking and Dining area stands out to be the area which most fires originate from.
+Living space and Transportation areas also have a great amount of incidence occurred.
 
@@
     This is a numerical variable that provides insight into the severity of the incidents.
-Basic descriptive statistics (e.g., mean, median, and range) were calculated using the `summary()` function.
+Basic descriptive statistics were calculated using the `summary()` function.
 Additionally, a scatter plot was created to explore the relationship between the number of civilian casualties and the incident ID (see @tbl-civilian-casualties-summary).
 We can see that most of the time, there are no casualties.
-This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fire places.
+This data set ranges across over 10 years (see @fig-fire-incidents-by-year in @sec-appendixa), looking at all levels of casualties, and taking its average over 6 years, we can see that this indicates the city has done a good job in protecting civilians from fireplaces.
 
@@
 A bar plot was created to visualize the distribution of different incident types.(see @fig-final-incident-type-distribution).
-It's clear that there are more fire than explotion.
+It's clear that there is more fire than explosion.
 
@@
 The distribution of ignition sources was visualized with a bar plot (see @fig-grouped-ignition-source-distribution).
-We can tell that fire-related and open flame is one of the biggest source of fire incidence; cooking and heating equipment are second place as sources, which align with @fig-area-origin-distribution result, where Ketchen is the place most fire incidence starts off.
+We can tell that fire-related and open flame is one of the biggest sources of fire incidence; cooking and heating equipment are second place as sources, which aligns with @fig-area-origin-distribution result, where Ketchen is the place most fire incidence starts off.
 
@@
 First, the total number of responding apparatus from TFS was calculated and divided into bins after plotting and observing its distribution.
-Each incident was assigned to a bin based on how many apparatus responded, helping categorize incidents in terms of the scale of response.
+Each incident was assigned to a bin based on how many apparatuses responded, helping categorize incidents in terms of the scale of response.
 
@@
 The transformed data was stored in a new variable called `log_estimated_dollar_loss`.
-Quantiles (25th, 50th, 75th percentiles) were used to divide the dollar loss into categories to make severity thresholds more meaningful.
+Quantiles (25th, 50th, and 75th percentiles) were used to divide the dollar loss into categories to make severity thresholds more meaningful.
 
@@
 
-A logistic regression model was chosen because the outcome variable, Severity_high, is binary (0/1), and logistic regression is well-suited for binary classification tasks, especially when working with categorical predictor variables.
+A logistic regression model was chosen because the outcome variable, Severity_high, is binary (0/1), and logistic regression is appropriate for binary classification tasks.
+Especially here, we are working with categorical predictor variables.
 The logistic regression model used the following predictors.
@@
 
-Notably, there is a slight class imbalance in the dataset, high-severity incidents are underrepresented.
+Notably, there is a slight class imbalance in the dataset, and high-severity incidents are underrepresented.
 In the training dataset, only 74 out of 10,237 records are classified as high severity.
@@
 
-The initial logistic regression model was evaluated for the significance of the predictors using the p-values from the model summary.
+The initial logistic regression model was tested for the significance of the predictors using the p-values from the model summary.
 Since a high p-value suggests that the variable does not significantly contribute to the model in prediction, it was removed to simplify the model and improve interpretability.
-This decision follows the principle outlined by Occam's Razor [@blumer1987occam], which proposes that people prefer simplicity.
+This decision also follows the principle outlined in Occam's Razor [@blumer1987occam], which proposes that simplicity is preferred.
 
 To check for multicollinearity among the predictors, the Variance Inflation Factor (VIF) was calculated.
-According to standard practice, if any predictors had a high VIF (greater than 5), they were considered for removal to avoid multicollinearity, which could distort the model's estimates [@alin2010multicollinearity].Based on the results, the variable `ignition_source_grouped` was removed from the model due to its non-significance (high p-value).
+According to standard practice, if any predictors had a high VIF (greater than 5), they were considered for removal to avoid multicollinearity, which could distort the model's estimates [@alin2010multicollinearity].
+Based on the results, the variable `ignition_source_grouped` was removed from the model due to its non-significance (high p-value).
 The updated model retained only `area_of_origin_grouped`, which showed a significant relationship with the likelihood of high-severity fires.
@@
 
-The logistic regression model was fitted to predict the probability of high severity fire incidents based on the categorical variables **area of origin**.
-The final model keep only significant predictors based on p-values and col-linearity diagnostics.
-below is a summary ( see @fig-logisticresults-summary) of the model's estimates.
+The logistic regression model was fitted to predict the probability of high-severity fire incidents based on the categorical variables `Area of Origin`.
+The final model keeps only significant predictors based on p-values and colinearity diagnostics.
+Below is a summary @fig-logisticresults-summary of the model's estimates.
 The coefficients of each predictor.
-And Also included other information about model evaluation.
+It also included other information about model evaluation.
 It's noticeable that RMSE is not ideal, but this is already an improvement from the initial model.
@@
 #| fig.align: left
-#| fig.width: 6
-#| fig.height: 4
+
 library(rstanarm)
@@
 We can tell most categories are relative to the severity of fire incidence.
-For example, Fire from mechanical, HVAC, and electrical Areas are highly contributing to the occurrence of ore sever fire incidence.
+For example, Fire from mechanical, HVAC, and electrical Areas are highly contributing to the occurrence of more severe fire incidents.
 

### 2024-12-15 paper/paper.qmd
fix discussion

@@
 While false positives may lead to overreaction, failing to detect a severe fire could escalate into a major disaster, resulting in significant financial losses and potentially causing more casualties.
-Therefore, a model that is tuned to minimize false negatives is preferable, as it prioritizes catching severe fires—even at the expense of some false positives.
+Therefore, a model that is tuned to minimize false negatives is preferable, as it prioritizes catching severe fires even at the expense of some false positives.
 However, I will continue to fit additional models to strike a better balance between specificity and sensitivity.
@@
 1.
-**Classification Bias**: Classification bias occurs when events are mis-classified into categories that do not accurately represent the event itself [@brogden1950theory].
+**Classification Bias**: Classification bias occurs when events are misclassified into categories that do not accurately represent the event itself [@brogden1950theory].
 This leads to incorrect or misleading conclusions.
